---
title: "Simulated_Data_Analysis"
author: "Ryley Dorney"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyr)
library(naniar)
library(readr)
library(dplyr)
library(ggplot2)
library(ggh4x)
library(biomaRt)
library(AnnotationHub)
library(HGNChelper)
library(UpSetR)
library(ComplexUpset)
library(patchwork)
library(ggforce)
library(ggpubr)
```

Identify resource usage of each tool :)

```{bash}
#bash ~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/ResourceUsage.sh
```

```{r}
Analysis_output_folder <- "/bioinformatics/ryley/Algorithm_Benchmark/benchmark_data_analysis/Analysis/benchmark_fusion_results/"

standard_chrs <- paste0("chr", c(1:22, "X", "Y", "M"))
#load relevant gtf file
gencodev44gtf <- import("/bioinformatics/ryley/Gencode44/reference_v44/gencode.v44.annotation.gtf")

########################################
# Make function to find antisense genes
########################################
genes <- gencodev44gtf %>%
  as.data.frame() %>%
  filter(type == "gene") %>%
  makeGRangesFromDataFrame(
    seqnames.field = "seqnames",
    start.field = "start",
    end.field = "end",
    strand.field = "strand",
    keep.extra.columns = TRUE
  )

################################################################
# Prepare dictionary
################################################################
ah <- AnnotationHub()
query(ah, c("EnsDb", "Hsapiens", "110"))#Load Ensembl the version that bests correlate to gencode version 44
ensembldbv110 <- ah[["AH113665"]]
query(ah, c("EnsDb", "Hsapiens", "109"))#Load Ensembl the version that bests correlate to gencode version 43
ensembldbv109 <- ah[["AH109606"]]
```

The 'biomarRt' package is used to help identify corresponding gene ids to gene names and vice versa.

```{r}
ensembl_current <-  useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl")
#the below ensembl versions best correlate to gencode version 43 and 44
ensemblv109 <- useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl", version = 109)
ensemblv110 <- useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl", version = 110)
```

Here we read in the data table containing information about the fusions generated with Fusim to be spiked in. We then modify the table to have separate columns for partner genes. We then use biomart to find the corresponding gene names/ids. This helps find potential alternative names/ids for gene fusions recalled.

```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/simulated_data_info/SimulatedDataInfo_format.R")
```

Here we perform a sanity check to ensure the simulated fusions are within our simulated fastq files. The fusion names will be within the headers of the fastq, all prefixed with "ref\|", therefore we extract fusion names using this. BadRead also simulates false chimeras, so I also need to remove fusions which were turned into false chimeras.

```{bash}
#Identify the fusions spiked in to the fastq files
fastq_directory=/bioinformatics/ryley/Algorithm_Benchmark/Adapter_porechop_trimmed/fastq_files/
cd ${fastq_directory}
zgrep -H '@.*ref' *Spiked*fastq.gz |
awk -F '[ ,:=]+' '{
    for (i=1; i<=NF; i++) {
        printf "%s%s", $i, (i==NF ? ORS : ",")
    }
}' > Spiked_Fusions.csv

grep "chimera" Spiked_Fusions.csv > Spiked_Fusions_with_chimeras.csv
grep -v "chimera" Spiked_Fusions.csv > Spiked_Fusions_no_chimeras.csv
```

Read the file generated containing fusions, and count how many fusions are in each file. We can also see how many supporting reads there are for each fusion

```{r}
Simulateddatainfo_directory <- "~/LongReadFusionCallerBenchmark/SimulatedData/simulated_data_info/"
Spiked_fusion_file <-  paste0(Simulateddatainfo_directory, "Spiked_Fusions_no_chimeras.csv")
Spiked_Fusions <- read.csv(Spiked_fusion_file, header = FALSE)
colnames(Spiked_Fusions) <- c("simulation", "read_id","gene_info","strand","range","lfdgp","length","ergrg","error_free_length","gr","read_identity")

Spiked_Fusions <- Spiked_Fusions %>% dplyr::select(c(simulation, read_id, gene_info, strand, range, length, error_free_length, read_identity )) %>%
  mutate(gene_info = gsub("ref\\|", "", gene_info)) %>%
  separate(gene_info, into = c("name1","name2", "name3"), sep = "-", remove = FALSE)%>%
  mutate(depth = case_when(
    grepl("100GB", simulation) ~ "100GB",
    grepl("10GB", simulation) ~ "10GB",
    grepl("1GB", simulation) ~ "1GB" # Keep original value if no match
  ), Sequence_Identity = case_when(
    grepl("95", simulation) ~ "95%",
    grepl("90", simulation) ~ "90%",
    grepl("85", simulation) ~ "85%" # Keep original value if no match
  ))
Spiked_Counts <- unique(Spiked_Fusions[c(1,3,12,13)]) |>
  dplyr::count(simulation, depth, Sequence_Identity, name = "Spiked_Number")
Spiked_ReadCounts <- Spiked_Fusions[c(1,3,12,13)] |>
  dplyr::count(simulation, gene_info, depth, Sequence_Identity, name = "Spiked_ReadSupp")
write_csv(Spiked_ReadCounts, file = paste0(Simulateddatainfo_directory, "Fusion_ReadCounts.csv"))

#the numbers above include our self-fusions which were excluded from analysis, so we will filter them out here and recount
Spiked_Counts_wo_SF <- unique(filter(Spiked_Fusions, name1 != name2)[c(1,3,12,13)]) |>
  dplyr::count(simulation, depth, Sequence_Identity, name = "Spiked_Number")
Spiked_ReadCounts_wo_SF <- filter(Spiked_Fusions, name1 != name2) |> #filter out self fusions
  dplyr::count(simulation, gene_info, depth, Sequence_Identity, name = "Spiked_ReadSupp")

```

LongGF suggests to extract the called fusions form the output log using 'grep SumGF'.

```{bash}
LONGGF_DIR=/bioinformatics/ryley/Gencode44/Algorithms/LongGF
#RPROJ_DIR=/home/ryleyd/LongReadFusionCallerBenchmark

#grep the results from the calls and create a tsv in the home directory
grep SumGF ${LONGGF_DIR}/LongGF*Spiked*.log > ${LONGGF_DIR}/LongGF_Spiked_summary.tsv
grep SumGF ${LONGGF_DIR}/*Adapter_Simulated*.log > ${LONGGF_DIR}/LongGF_Negative_Control_summary.tsv

```

Read in the LongGF fusions into RStudio and separate components of output by partner genes.

```{r}
#LongGF data wrangling
RPROJ_DIR<-"/home/ryleyd/LongReadFusionCallerBenchmark"
LONGGF_DIR<-"/bioinformatics/ryley/Gencode44/Algorithms/LongGF"
#list files
my_files <- list.files(path = LONGGF_DIR, pattern = "tsv", full.names=TRUE)
all_tsv <-lapply(my_files, read.table)
names <- gsub(".tsv","", list.files(path = LONGGF_DIR, pattern = "tsv",full.names = FALSE),fixed = TRUE)
#assign each longGF file
for(i in names){
  assign(i, read.table(paste0(LONGGF_DIR,"/", i, ".tsv")))
}
LongGF_Sim <- do.call(rbind, lapply(my_files, function(filename) {
  read.table(filename) %>%
    mutate(Source = V1)
})) %>% 
  mutate(control = ifelse(grepl("Spiked", Source), "positive", "negative"))%>% 
  separate(V2, into = c("Gene1", "Gene2"), sep = ":", remove = FALSE) %>% 
  separate(V4, into = c("chromosome1", "breakpoint1"), sep = ":", remove = FALSE) %>% 
  separate(V5, into = c("chromosome2", "breakpoint2"), sep = ":", remove = FALSE) %>% #V2 = fusion gene name :
    filter(chromosome1 %in% standard_chrs)%>%
  filter(chromosome2 %in% standard_chrs)

ogLongGF <- LongGF_Sim
```

BiomaRt is then used to find the corresponding gene names/ids for fusions called by LongGF.

```{r}
#gets ensembl gene ID and check it matches the putative chromosome 
Gene_Name<-rbind (getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name"),
                             filters = "external_gene_name",
                             values = unique(c(LongGF_Sim$Gene1, LongGF_Sim$Gene2)),
                             mart = ensemblv110)) %>% unique()
Gene_Name$chromosome_name <- paste0("chr", Gene_Name$chromosome_name)

  
  LongGF_Sim_ensembl <- LongGF_Sim %>%
  left_join(Gene_Name, by = c('Gene1' = 'external_gene_name')) %>%
  left_join(Gene_Name, by = c('Gene2' = 'external_gene_name')) %>%
  mutate(
    ensembl_gene1 = ifelse(is.na(ensembl_gene_id.x) | nchar(ensembl_gene_id.x) == 0, Gene1, ensembl_gene_id.x),
    ensembl_gene2 = ifelse(is.na(ensembl_gene_id.y) | nchar(ensembl_gene_id.y) == 0, Gene2, ensembl_gene_id.y)
  ) %>%
  unite("fusion.gene.id", c(ensembl_gene1, ensembl_gene2), sep = ":", remove = FALSE, na.rm = TRUE) %>%
  # Create a filtering condition using case_when
  filter(case_when(
      is.na(chromosome_name.x) ~ TRUE, # Keep if chromosome_name.x is NA
      chromosome_name.x == chromosome1 ~ TRUE, # Keep if chromosome names match
      chromosome_name.x == "chrMT" & chromosome1 == "chrM" ~ TRUE, # Allow chrMT/chrM equivalence
      grepl("PATCH|HSCHR", chromosome_name.x) & !grepl("^chr([1-9]|1[0-9]|2[0-2]|X|Y|M)$", chromosome1) ~ TRUE, 
      TRUE ~ FALSE # Remove rows that don't meet the above conditions
    )) %>%
  filter(case_when(
      is.na(chromosome_name.y) ~ TRUE,
      chromosome_name.y == chromosome2 ~ TRUE,
      chromosome_name.y == "chrMT" & chromosome2 == "chrM" ~ TRUE,
      grepl("PATCH|HSCHR", chromosome_name.y) & !grepl("^chr([1-9]|1[0-9]|2[0-2]|X|Y|M)$", chromosome2) ~ TRUE,
      TRUE ~ FALSE # Remove rows that don't meet the above conditions
    )) 
```

Check if LongGF called any fusions that were spiked in

```{r}
Annot_LongGF_Sim <- LongGF_Sim_ensembl %>% 
  left_join(Simulated_Fusion_Info_2, by = c("fusion.gene.id"), relationship = "many-to-many")
```

Assign label to false fusions or partially recalled fusions by LongGF

```{r}
###############################
# Check antisense sense fusions
###############################
#Load necessary library
library(rtracklayer)
library(GenomicRanges)
library(dplyr)
#load relevant gtf file
gencodev44gtf <- import("/bioinformatics/ryley/Gencode44/reference_v44/gencode.v44.annotation.gtf")

########################################
# Make function to find antisense genes
########################################
genes <- gencodev44gtf %>%
  as.data.frame() %>%
  filter(type == "gene") %>%
  makeGRangesFromDataFrame(
    seqnames.field = "seqnames",
    start.field = "start",
    end.field = "end",
    strand.field = "strand",
    keep.extra.columns = TRUE
  )

get_antisense_overlaps <- function(gene, genes_gr, id_type = c("gene_name", "gene_id")) {
  
  id_type <- match.arg(id_type)
  
  if (id_type == "gene_id") {
    # strip version if present (ENSG00000xxxx.y)
    gene <- sub("\\..*$", "", gene)
    gene_ids <- sub("\\..*$", "", genes_gr$gene_id)
    target <- genes_gr[gene_ids == gene]
  } else {
    target <- genes_gr[genes_gr$gene_name == gene]
  }
  
  if (length(target) == 0)
    stop("Gene not found in annotation")
  
  hits <- findOverlaps(target, genes_gr, ignore.strand = TRUE)
  
  query_idx   <- queryHits(hits)
  subject_idx <- subjectHits(hits)
  
  hit_df <- data.frame(
    target_gene   = as.character(target$gene_name[query_idx]),
    target_gene_id = as.character(target$gene_id[query_idx]),
    target_strand = as.character(strand(target)[query_idx]),
    gene_name     = as.character(genes_gr$gene_name[subject_idx]),
    gene_id       = as.character(genes_gr$gene_id[subject_idx]),
    gene_type     = genes_gr$gene_type[subject_idx],
    seqnames      = as.character(seqnames(genes_gr)[subject_idx]),
    start         = start(genes_gr)[subject_idx],
    end           = end(genes_gr)[subject_idx],
    strand        = as.character(strand(genes_gr)[subject_idx]),
    stringsAsFactors = FALSE
  )
  
  hit_df %>%
    dplyr::filter(
      gene_id != target_gene_id,   # remove self
      strand != target_strand      # opposite strand
    ) %>%
    dplyr::distinct(
      gene_name, gene_id, gene_type,
      seqnames, start, end, strand
    ) 
}

LongGF_sensecheck <- LongGF_Sim
# get unique genes from both columns
genes_to_check <- unique(c(LongGF_sensecheck$Gene1, LongGF_sensecheck$Gene2))

# run get_antisense_overlaps on each gene
LongGF_antisense_list <- lapply(genes_to_check, 
                                function(g) get_antisense_overlaps(g, genes, id_type = "gene_name"))

# combine into a single data frame
LongGF_antisense_df <- bind_rows(LongGF_antisense_list, .id = "query_gene_index")

# add the original gene names
LongGF_antisense_df$query_gene <- genes_to_check[as.integer(LongGF_antisense_df$query_gene_index)]
LongGF_antisense_df$query_gene_index <- NULL

antisense_pairs <- LongGF_antisense_df %>%
  dplyr::select(query_gene, gene_name) %>%
  distinct()

Annot_LongGF_Sim <- Annot_LongGF_Sim %>%
  group_by(V2.x) %>%
  mutate(fusionType = case_when(
    is.na(fusionType) ~ paste0("chromosomal_misalignment:", dplyr::first(na.omit(fusionType))),  # Only update NA values
    .default = fusionType  # Keep existing values unchanged
  ))%>% ungroup()%>%
  replace_with_na(replace = list(fusionType = "chromosomal_misalignment:NA")) 

Annot_LongGF_Sim <- LongGF_Sim_ensembl %>% left_join(Simulated_Fusion_Info_2, by = c("fusion.gene.id"), relationship = "many-to-many") %>%
  group_by(V2.x) %>%
  mutate(fusionType = case_when(
    is.na(fusionType) ~ paste0("chromosomal_misalignment:", dplyr::first(na.omit(fusionType))),  # Only update NA values
    .default = fusionType  # Keep existing values unchanged
  ))%>% ungroup()%>%
  replace_with_na(replace = list(fusionType = "chromosomal_misalignment:NA"))  
Annot_LongGF_Sim$fusionType <- mapply(function(g1, g2, current_type, chr1, chr2, gene_name1, gene_name2) {
   # Check if the current fusionType is empty
  if (current_type == "" || is.na(current_type)) {
    #check for truncated_tri_fusions
    if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g1, ":", g2)) |
            str_detect(paste(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V1, subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V3, sep = ":"), paste0(g1, ":", g2)))) {
      return("truncated_tri_fusion")
      #check for reverse order fusions
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1))|
                  str_detect(paste(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V1, subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V3, sep = ":"), paste0(g2, ":", g1)))){
      return("reverse_order:truncated_tri_fusion") 
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1)))){
       matching_row <- subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusionType[which(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1)))]
      return(paste("reverse_order:", matching_row[1])) 
       #check for chromosomal misalignment
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$original.fusion.gene.id, paste0(g1, ":", g2)))){
       matching_row <- subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusionType[which(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$original.fusion.gene.id, paste0(g1, ":", g2)))]
      return(paste("chromosomal_misalignment:", matching_row[1])) 
       #check for false fusions
    } else if ((grepl("chrM:", chr1, ignore.case = TRUE) & !grepl("chrM:", chr2, ignore.case = TRUE)) | (!grepl("chrM:", chr1, ignore.case = TRUE) & grepl("chrM:", chr2, ignore.case = TRUE))){
      return("false_fusion:mitochondrial_genomic") 
    }else if (grepl("chrM:", chr1, ignore.case = TRUE) & grepl("chrM:", chr2, ignore.case = TRUE)){
      return("false_fusion:mitochondrial") 
    }else if ((gene_name1 == gene_name2) & (chr1 != chr2)){
      return("false_fusion:self_misalignment") 
    }else if (paste(gene_name1, gene_name2) %in% paste(antisense_pairs$query_gene, antisense_pairs$gene_name)
      | 
        paste(gene_name2, gene_name1) %in% paste(antisense_pairs$query_gene, antisense_pairs$gene_name)){
      return("false_fusion:Sense-Antisense") 
    }else {
      return("false_fusion")
    }
  }
  return(current_type)
}, 
Annot_LongGF_Sim$ensembl_gene1, Annot_LongGF_Sim$ensembl_gene2, Annot_LongGF_Sim$fusionType, 
Annot_LongGF_Sim$chromosome1, Annot_LongGF_Sim$chromosome2,
Annot_LongGF_Sim$Gene1, Annot_LongGF_Sim$Gene2)
```

Repeat the same above pre-processing steps for FusionSeeker

```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/PreprocessFusionSeeker.R")
```

Repeat pre-processing steps for JAFFAL

```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/PreprocessJAFFAL.R")
```

Repeat pre-processing steps for Genion

```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/PreprocessGenion.R") 
```

Repeat pre-processing steps for CTAT-LR

```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/PreprocessCTAT-LR.R")
```

Append corresponding depth and sequencing identity labels

```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/Append_DepSeqIDALg.R")
```

Label fusions as False Calls, Partial Real and True Recall in Genion and JAFFAL datasets. Also write Genion_Fffscores.tsv and JAFFAL_scoring.tsv

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/AlgorithmRecallCategory.R")

supp_Figure15
```

Consolidate fusion caller outputs into one dataframe to more easily compare

write combined_data.tsv and combined_data_annotation.tsv

```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/Consolidate_df.R")
```

Check number of false, partial and true calls. Generate Supplementary Figure 2

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/ColouredGridStats.R")
# Display the combined plot
Supplementary_Figure2
ggsave("~/LongReadFusionCallerBenchmark/Figures/Supplementary_Figure2.png", Supplementary_Figure2, width = 210, height = 210, units = "mm")
```

Calculate Recall vs Precision, and F1 score.

Also write stat_summary.tsv and readsupp_filtering.tsv

```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/Precision_Recall_ReadSupp_stats.R") 

```

Generate Supplementary Figures 8-10

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Readsupport_vs_Precision-F1.R")
```

Make Figure 3A

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/F1_RecPrec_subset.R")
```

Check the effect of removing obvious false chimeras from the fusion calls to see if they greatly influence precision. There visually appears to be minimal/negligible influence. also export manual_readsupp_filtering.tsv

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Manual_Filtering.R")
```

Create a summary of the number of false, partial and true calls in every simulation. Also write fusion_profile_stat_summary.tsv

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/FusionProfile_StatSummary.R")
```

Make Supplementary Figure 1 and 4

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Fusion_Profile_Grid.R")
```

Make Figure 1 part D-G

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Figure1.R")
```

Make Supplementary Figure 5

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Make_supFig5.R")
```

Create Supplementary Figure 6

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Make_supFig6.R")
```

Make Supplementary Figure 7

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Make_supFig7.R")
```

Make Figure 2

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Figure2.R")
```

Fusionseeker and some of fusion callers will produce duplicate calls the same fusion multiple times with different read support because of breakpoint differences. We need to consolidate these for ease in comparing how read support for the same fusion will differ by fusion caller tool. Export fusions_in_all_algorithms.tsv

```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/ReadSupport_Consolidation.R")
```

Make Figure 3B: comparing read support for the same fusions as called by different algorithms

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/ReadSuppComp.R")
```

Make Compound Figure 3

```{r}
Figure3<- (Figure3A / Figure3B) +
          plot_annotation(tag_levels = list(c("A", "", "", "B")))
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure3_paper.png", plot = Figure3, width = 210, height = 297, units = "mm")  
```

identify fusions in more than one algorithm, and export fusions_in_n_algorithms.tsv

```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/FusionMultiAlg.R")
```

Supplementary Figure 11

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Make_SupFig11.R")
```

Generate upset plots to see the total number of overlapping fusions between fusion callers Make Figure 1H: True Calls Overlap

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/TrueCalls_Overlap.R")
```

Create Compound Figure 1:

```{r}
Figure1_paper <- (Figure1D | Figure1E)/
                  (Figure1F | Figure1G)/
                  (True_overlap) +   
  plot_annotation(tag_levels = list(c("D", "E", "F", "G", "H")), 
                  theme = theme(plot.title = element_text(size = 14, face = "bold")                                ))
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure1_paper.pdf", plot = Figure1_paper, width = 300, height = 430, units = "mm", limitsize = FALSE)
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure1_paper.png", plot = Figure1_paper, width = 300, height = 430, units = "mm", limitsize = FALSE)
```

Make Supplementary Figure 3

```{r}
#Supplementary Figure 3A
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/FusionTypesOverlap_Upset.R")
supp_Figure3
```

Repeat analysis with removed obvious false chimeras to see if they have any influence. Any differences observed are negligible

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Overlap_ManualFiltering.R")
```

Calculate the breakpoint accuracy. Write absolute_Breakpoint_Accuracy.tsv

```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/Breakpoint_Accuracy.R")
```

Make Figure 4 & Supplementary Figure 13

```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Breakpoint_stat_plots.R")
```
