---
title: "simulated_data_analysis"
author: "Ryley Dorney"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(naniar)
library(readr)
library(dplyr)
library(ggplot2)
library(ggh4x)
library(biomaRt)
library(UpSetR)
library(ComplexUpset)
library(patchwork)
```

```{r}
Analysis_output_folder <- "/bioinformatics/ryley/Algorithm_Benchmark/benchmark_data_analysis/Analysis/benchmark_fusion_results/"
```

The 'biomarRt' package is used to help identify corresponding gene ids to gene names and vice versa. Whilst gencode version 43 was used for this study, we also check the current ensembl version for where ensembl IDs may have been assigned a gene name in the newer version for comparing with literature.

```{r}
ensembl_current <-  useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl")
#the below ensembl versions best correlate to gencode version 43
ensemblv109 <- useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl", version = 109)
ensemblv110 <- useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl", version = 110)
ensembljul2023 <- useMart("ensembl", dataset = "hsapiens_gene_ensembl", host = "https://jul2023.archive.ensembl.org")
```

Here we read in the data table containing information about the fusions generated with Fusim to be spiked in. We then modify the table to have separate columns for partner genes. We then use biomart to find the corresponding gene names/ids. This helps find potential alternative names/ids for gene fusions recalled.
```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/simulated_data_info/SimulatedDataInfo_format.R")
```

LongGF suggests to extract the called fusions form the output log using 'grep SumGF'.
```{bash}
pwd
#change to folder contain LongGF fusion calls
cd /bioinformatics/ryley/Algorithm_Benchmark/Adapter_porechop_trimmed/LongGF/
#grep the results from the calls and create a tsv in the home directory
grep SumGF LongGF*Spiked*.log > /home/ryleyd/LongReadFusionCallerBenchmark/LongGF_Spiked.tsv
grep SumGF *Adapter_Simulated*.log > /home/ryleyd/LongReadFusionCallerBenchmark/LongGF_Negative_Control.tsv
#return to home directory
cd ~
```

Read in the LongGF fusions into RStudio and separate components of output by partner genes. 
```{r}
#LongGF data wrangling
#list files
my_files <- list.files(path = "/home/ryleyd/LongReadFusionCallerBenchmark", pattern = "LongGF", full.names=TRUE)
all_tsv <-lapply(my_files, read.table)
names <- gsub(".tsv","", list.files(path = "/home/ryleyd/LongReadFusionCallerBenchmark", pattern = "LongGF",full.names = FALSE),fixed = TRUE)
#assign each longGF file
for(i in names){
  assign(i, read.table(paste0("/home/ryleyd/LongReadFusionCallerBenchmark/", i, ".tsv")))
}
#seperate columns to more easily identify information correlating to each gene partner
LongGF_Sim <- rbind(LongGF_Negative_Control, LongGF_Spiked)%>%
      mutate(Source = basename(V1)) %>% mutate(control = ifelse(grepl("Spiked", Source), "positive", "negative"))%>% 
  separate(V2, into = c("Gene1", "Gene2"), sep = ":", remove = FALSE) %>% 
  separate(V4, into = c("chromosome1", "breakpoint1"), sep = ":", remove = FALSE) %>% 
  separate(V5, into = c("chromosome2", "breakpoint2"), sep = ":", remove = FALSE) #V2 = fusion gene name :
ogLongGF <- LongGF_Sim
```

BiomaRt is then used to find the corresponding gene names/ids for fusions called by LongGF.
```{r}
#gets ensembl gene ID and check it matches the putative chromosome 
Gene_Name<-rbind (getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name"),
                             filters = "external_gene_name",
                             values = unique(c(LongGF_Sim$Gene1, LongGF_Sim$Gene2)),
                             mart = ensembljul2023),
                  getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name"),
                             filters = "external_gene_name",
                             values = unique(c(LongGF_Sim$Gene1, LongGF_Sim$Gene2)),
                             mart = ensemblv109)) %>% unique()
Gene_Name$chromosome_name <- paste0("chr", Gene_Name$chromosome_name)

  
  LongGF_Sim_ensembl <- LongGF_Sim %>%
  left_join(Gene_Name, by = c('Gene1' = 'external_gene_name')) %>%
  left_join(Gene_Name, by = c('Gene2' = 'external_gene_name')) %>%
  mutate(
    ensembl_gene1 = ifelse(is.na(ensembl_gene_id.x) | nchar(ensembl_gene_id.x) == 0, Gene1, ensembl_gene_id.x),
    ensembl_gene2 = ifelse(is.na(ensembl_gene_id.y) | nchar(ensembl_gene_id.y) == 0, Gene2, ensembl_gene_id.y)
  ) %>%
  unite("fusion.gene.id", c(ensembl_gene1, ensembl_gene2), sep = ":", remove = FALSE, na.rm = TRUE) %>%
  # Create a filtering condition using case_when
  filter(case_when(
      is.na(chromosome_name.x) ~ TRUE, # Keep if chromosome_name.x is NA
      chromosome_name.x == chromosome1 ~ TRUE, # Keep if chromosome names match
      chromosome_name.x == "chrMT" & chromosome1 == "chrM" ~ TRUE, # Allow chrMT/chrM equivalence
      grepl("PATCH|HSCHR", chromosome_name.x) & !grepl("^chr([1-9]|1[0-9]|2[0-2]|X|Y|M)$", chromosome1) ~ TRUE, 
      TRUE ~ FALSE # Remove rows that don't meet the above conditions
    )) %>%
  filter(case_when(
      is.na(chromosome_name.y) ~ TRUE,
      chromosome_name.y == chromosome2 ~ TRUE,
      chromosome_name.y == "chrMT" & chromosome2 == "chrM" ~ TRUE,
      grepl("PATCH|HSCHR", chromosome_name.y) & !grepl("^chr([1-9]|1[0-9]|2[0-2]|X|Y|M)$", chromosome2) ~ TRUE,
      TRUE ~ FALSE # Remove rows that don't meet the above conditions
    )) 
```

Check if LongGF called any fusions that were spiked in
```{r}
Annot_LongGF_Sim <- LongGF_Sim_ensembl %>% 
  left_join(Simulated_Fusion_Info_2, by = c("fusion.gene.id"), relationship = "many-to-many")
```

Assign label to false fusions or partially recalled fusions by LongGF
```{r}
Annot_LongGF_Sim <- Annot_LongGF_Sim %>%
  group_by(V2.x) %>%
  mutate(fusionType = case_when(
    is.na(fusionType) ~ paste0("chromosomal_misalignment:",first(na.omit(fusionType))),  # Only update NA values
    .default = fusionType  # Keep existing values unchanged
  ))%>% ungroup()%>%
  replace_with_na(replace = list(fusionType = "chromosomal_misalignment:NA")) 

Annot_LongGF_Sim <- LongGF_Sim_ensembl %>% left_join(Simulated_Fusion_Info_2, by = c("fusion.gene.id"), relationship = "many-to-many") %>%
  group_by(V2.x) %>%
  mutate(fusionType = case_when(
    is.na(fusionType) ~ paste0("chromosomal_misalignment:",first(na.omit(fusionType))),  # Only update NA values
    .default = fusionType  # Keep existing values unchanged
  ))%>% ungroup()%>%
  replace_with_na(replace = list(fusionType = "chromosomal_misalignment:NA"))  
Annot_LongGF_Sim$fusionType <- mapply(function(g1, g2, current_type, chr1, chr2, gene_name1, gene_name2) {
   # Check if the current fusionType is empty
  if (current_type == "" || is.na(current_type)) {
    #check for truncated_tri_fusions
    if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g1, ":", g2)) |
            str_detect(paste(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V1, subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V3, sep = ":"), paste0(g1, ":", g2)))) {
      return("truncated_tri_fusion")
      #check for reverse order fusions
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1))|
                  str_detect(paste(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V1, subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V3, sep = ":"), paste0(g2, ":", g1)))){
      return("reverse_order:truncated_tri_fusion") 
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1)))){
       matching_row <- subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusionType[which(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1)))]
      return(paste("reverse_order:", matching_row[1])) 
       #check for chromosomal misalignment
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$original.fusion.gene.id, paste0(g1, ":", g2)))){
       matching_row <- subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusionType[which(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$original.fusion.gene.id, paste0(g1, ":", g2)))]
      return(paste("chromosomal_misalignment:", matching_row[1])) 
       #check for false fusions
    } else if ((grepl("chrM:", chr1, ignore.case = TRUE) & !grepl("chrM:", chr2, ignore.case = TRUE)) | (!grepl("chrM:", chr1, ignore.case = TRUE) & grepl("chrM:", chr2, ignore.case = TRUE))){
      return("false_fusion:mitochondrial_genomic") 
    }else if (grepl("chrM:", chr1, ignore.case = TRUE) & grepl("chrM:", chr2, ignore.case = TRUE)){
      return("false_fusion:mitochondrial") 
    }else if ((gene_name1 == gene_name2) & (chr1 != chr2)){
      return("false_fusion:self_misalignment") 
    }else {
      return("false_fusion")
    }
  }
  return(current_type)
}, Annot_LongGF_Sim$ensembl_gene1, Annot_LongGF_Sim$ensembl_gene2, Annot_LongGF_Sim$fusionType, Annot_LongGF_Sim$chromosome1, Annot_LongGF_Sim$chromosome2, Annot_LongGF_Sim$Gene1, Annot_LongGF_Sim$Gene2)
```

Repeat the same above pre-processing steps for FusionSeeker
```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/PreprocessFusionSeeker.R")
```

Repeat pre-processing steps for JAFFAL
```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/PreprocessJAFFAL.R")
```

Repeat pre-processing steps for Genion
```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/PreprocessGenion.R") 
```

Repeat pre-processing steps for CTAT-LR 
```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/PreprocessCTAT-LR.R")
```

Append corresponding depth and sequencing identity labels 
```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/Append_DepSeqIDALg.R")
```

Label fusions as False Calls, Partial Real and True Recall in Genion and JAFFAL datasets
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/AlgorithmRecallCategory.R")
```

Consolidate fusion caller outputs into one dataframe to more easily compare
```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/Consolidate_df.R")
```

Check number of false, partial and true calls. Generate Supplementary Figure 2
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/ColouredGridStats.R")
# Display the combined plot
Supplementary_Figure2
ggsave("~/LongReadFusionCallerBenchmark/Figures/Supplementary_Figure2.png", Supplementary_Figure2, width = 210, height = 210, units = "mm")
```

Calculate Recall vs Precision, and F1 score
```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/Precision_Recall_ReadSupp_stats.R") 
```

Generate Supplementary Figures 8-10
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Readsupport_vs_Precision-F1.R")
```

Make Figure 3A 
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/F1_RecPrec_subset.R")
```

Check the effect of removing obvious false chimeras from the fusion calls to see if they greatly influence precision. There visually appears to be minimal/negligible influence.
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Manual_Filtering.R")
```

Create a summary of the number of false, partial and true calls in every simulation
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/FusionProfile_StatSummary.R")
```

Make Supplementary Figure 1 and 4
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Fusion_Profile_Grid.R")
```

Make Figure 1 part D-G
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Figure1.R")
```

Make Supplementary Figure 5
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Make_supFig5.R")
```

Create Supplementary Figure 6
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Make_supFig6.R")
```

Make Supplementary Figure 7
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Make_supFig7.R")
```

Make Figure 2
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Figure2.R")
```

Fusionseeker and some of fusion callers will produce duplicate calls the same fusion multiple times with different read support because of breakpoint differences. We need to consolidate these for ease in comparing how read support for the same fusion will differ by fusion caller tool
```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/ReadSupport_Consolidation.R")
```

Make Figure 3B: comparing read support for the same fusions as called by different algorithms
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/ReadSuppComp.R")
```
Make Compound Figure 3
```{r}
Figure3<- (Figure3A / Figure3B) +
          plot_annotation(tag_levels = list(c("A", "", "", "B")))
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure3_paper.png", plot = Figure3, width = 210, height = 297, units = "mm")  
```

identify fusions in more than one algorithm
```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/FusionMultiAlg.R")
```

Supplementary Figure 11
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Make_SupFig11.R")
```

Generate upset plots to see the total number of overlapping fusions between fusion callers
```{r}
source("~/LongReadFusionCallerBenchmark/SimulatedData/Data_Analysis_Scripts/Overlap_sets.R")
```

Make Figure 1H: True Calls Overlap
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/TrueCalls_Overlap.R")
```
Create Compound Figure 1:
```{r}
Figure1_paper <- (Figure1D | Figure1E)/
                  (Figure1F | Figure1G)/
                  (True_overlap) +   
  plot_annotation(tag_levels = list(c("D", "E", "F", "G", "H")), 
                  theme = theme(plot.title = element_text(size = 14, face = "bold")                                ))
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure1_paper.pdf", plot = Figure1_paper, width = 300, height = 430, units = "mm", limitsize = FALSE)
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure1_paper.png", plot = Figure1_paper, width = 300, height = 430, units = "mm", limitsize = FALSE)
```

Make Supplementary Figure 3 
```{r}
#Supplementary Figure 3A
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/FusionTypesOverlap_Upset.R")
supp_Figure3
```

Repeat analysis with removed obvious false chimeras to see if they have any influence. Any differences observed are negligible
```{r}
source("~/LongReadFusionCallerBenchmark/Figures/SCRIPTS/Overlap_ManualFiltering.R")
```

Calculate the breakpoint accuracy
```{r}
Simulated_BreakPoint <- Simulated_Fusion_Info_3 
JAFFAL_BreakPoint <- Annot_JAFFAL_Sim[c(1:25, 43, 50:52)] %>% distinct() %>% filter(control == "positive", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType), spanning.reads >= 2) 
  #chrom1, base1, chrom2, base2
SB <- subset(Simulated_BreakPoint, fusion.gene.id == "ENSG00000009830:ENSG00000087258")
J <- subset(JAFFAL_BreakPoint, fusion.gene.id == "ENSG00000009830:ENSG00000087258" & depth == "100GB" & Sequence_Identity == "95%")
gene_names <- str_split("ENSG00000009830:ENSG00000087258", pattern = ":")[[1]] 
      gene1 <- gene_names[1]
      gene2 <- gene_names[2]
     
      SB1 <- as.integer(tail(str_split(subset(SB, alternative_ID == gene1)$exonEnds, ",")[[1]], 1))
      SB2 <- as.integer(head(str_split(subset(SB, alternative_ID == gene2)$exonStarts, ",")[[1]],1))

      JB1 <- as.numeric(J$base1) - SB1
      JB2 <- as.numeric(J$base2) - SB2
 
 list(c(gene1, gene2) , c(SB1, SB2) , c(JB1, JB2))
 
 nrow(distinct(Breakpoint_Accuracy))

Genion_BreakPoint <- Annot_Genion_Sim[c(1:27, 37, 44:46)] %>% distinct() %>% filter(control == "positive", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType)) %>% separate(V8, into = c("Gene1", "Gene2", "Gene3"), sep = ";") %>% 
  separate(Gene1, into = c("Gene1.exonStarts", "Gene1.exonEnds"), sep = "-") %>% separate(Gene2, into = c("Gene2.exonStarts", "Gene2.exonEnds"), sep = "-") %>% separate(Gene3, into = c("Gene3.exonStarts", "Gene3.exonEnds"), sep = "-") %>%
  separate(Gene1.exonStarts, into=c("C1", "Gene1.exonStarts")) %>% separate(Gene2.exonStarts, into=c("C2", "Gene2.exonStarts")) %>% separate(Gene3.exonStarts, into=c("C3", "Gene3.exonStarts")) %>% mutate_at(vars(C1, C2, C3), funs(paste0("chr", .)))# "C1",  "Gene1.2", "C2", "Gene2.1", "Gene2.2", "C3", "Gene3.1" 

JAFFAL_BreakPoint <- Annot_JAFFAL_Sim[c(1:25, 43, 50:52)] %>% distinct() %>% filter(control == "positive", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType), spanning.reads >= 2) #chrom1, base1, chrom2, base2
LongGF_BreakPoint <- Annot_LongGF_Sim[c(1:24, 34, 41:43)] %>% distinct() %>% filter(control == "positive", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType)) #chrom1, breakpoint1, chrom2, base2
FusionSeeker_BreakPoint <- Annot_FusionSeeker_Sim[c(1:28, 38, 45:47)] %>% distinct() %>% filter(control == "positive", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType)) #Chrom1, Breakpoint1, Chrom2, Base2

JAFFAL_BreakPoint_Accuracy <-  data.frame(matrix(nrow = 0, ncol = length(c('seq_depth', 'seq_id', 'gene_fusion', 'JB1', 'JB2'))))
colnames(JAFFAL_BreakPoint_Accuracy) <- c('seq_depth', 'seq_id', 'gene_fusion','JB1', 'JB2')
for (seq_depth in c('1GB' , '10GB' , '100GB')) {
  for (seq_id in c('85%' , '90%' , '95%')) {   
    for(gene_fusion in unique(c(JAFFAL_BreakPoint$fusion.gene.id))){
      SB <- subset(Simulated_BreakPoint, fusion.gene.id == gene_fusion)
      
      gene_names <- str_split(gene_fusion, pattern = ":")[[1]] 
      gene1 <- gene_names[1]
      gene2 <- gene_names[2]
     
      SB1 <- as.integer(tail(str_split(subset(SB, alternative_ID == gene1)$exonEnds, ",")[[1]], 1))
      SB2 <- as.integer(head(str_split(subset(SB, alternative_ID == gene2)$exonStarts, ",")[[1]],1))
    
      gene3 <- gene_names[3]
      SB3 <- NA
      SB4 <- NA
      if(is.na(gene3) == FALSE){
        SB3 <- as.integer(str_split(subset(SB, alternative_ID == gene2)$exonEnds, ",")[[1]])
        SB4 <- as.integer(str_split(subset(SB, alternative_ID == gene3)$exonStarts, ",")[[1]])
      }
      
      J  <- subset(JAFFAL_BreakPoint, fusion.gene.id == gene_fusion & depth == seq_depth & Sequence_Identity == seq_id) 
      if(nrow(J) > 0){
          JB1 <- as.numeric(J$base1) - SB1
          JB2 <- as.numeric(J$base2) - SB2
          }else{
            JB1 <- "NA"
            JB2 <- "NA"
            }
      new_row <-  data.frame(seq_depth, seq_id, gene_fusion, JB1, JB2)
        colnames(new_row) <- colnames(JAFFAL_BreakPoint_Accuracy)
         JAFFAL_BreakPoint_Accuracy <- rbind(JAFFAL_BreakPoint_Accuracy, new_row)
    }
  }
}

Genion_Breakpoint_Accuracy <- data.frame(matrix(nrow = 0, ncol = length(c('seq_depth', 'seq_id', 'gene_fusion', 'GB1', 'GB2', 'GB3', 'GB4'))))
colnames(Genion_Breakpoint_Accuracy) <- c('seq_depth', 'seq_id', 'gene_fusion', 'GB1', 'GB2', 'GB3', 'GB4')
for (seq_depth in c('1GB' , '10GB' , '100GB')) {
  for (seq_id in c('85%' , '90%' , '95%')) {   
    for(gene_fusion in unique(c(Genion_BreakPoint$V1))){
      SB <- subset(Simulated_BreakPoint, fusion.gene.id == gene_fusion)
      gene_names <- str_split(gene_fusion, pattern = ":")[[1]] 
      gene1 <- gene_names[1]
      gene2 <- gene_names[2]
     
      SB1 <- as.integer(tail(str_split(subset(SB, alternative_ID == gene1)$exonEnds, ",")[[1]], 1))
      SB2 <- as.integer(head(str_split(subset(SB, alternative_ID == gene2)$exonStarts, ",")[[1]],1))
      
      gene3 <- gene_names[3]
      SB3 <- NA
      SB4 <- NA
      if(is.na(gene3) == FALSE){
        SB3 <- as.integer(str_split(subset(SB, alternative_ID == gene2)$exonEnds, ",")[[1]])
        SB4 <- as.integer(str_split(subset(SB, alternative_ID == gene3)$exonStarts, ",")[[1]])
      } 
      G <- subset(Genion_BreakPoint,  V1 == gene_fusion & depth == seq_depth & Sequence_Identity == seq_id)
        if(nrow(G) > 0){
    
          GB1 <- as.numeric(G$Gene1.exonEnds) - SB1
          GB2 <- as.numeric(G$Gene2.exonStarts) - SB2
          GB3 <- "NA"
          GB4 <- "NA"
          if(is.na(gene3) == FALSE){
                GB3 <- toString(apply(outer(as.numeric(G$Gene2.exonEnds) , SB3, "-"), 1, function(x) x[which.min(abs(x))]))
                GB4 <- toString(apply(outer(as.numeric(G$Gene3.exonStarts) , SB4, "-"), 1, function(x) x[which.min(abs(x))]))
              }
          }else{
            GB1 <- "NA"
            GB2 <- "NA"
            GB3 <- "NA"
            GB4 <- "NA"
          }
     new_row <-  data.frame(seq_depth, seq_id, gene_fusion, GB1, GB2, GB3, GB4)
     colnames(new_row) <- colnames(Genion_Breakpoint_Accuracy)
     Genion_Breakpoint_Accuracy <- rbind(Genion_Breakpoint_Accuracy, new_row)
    }
   } 
  }
      
LongGF_Breakpoint_Accuracy <- data.frame(matrix(nrow = 0, ncol = length(c('seq_depth', 'seq_id', 'gene_fusion', 'LB1', 'LB2'))))
colnames(LongGF_Breakpoint_Accuracy) <- c('seq_depth', 'seq_id', 'gene_fusion', 'LB1', 'LB2')
for (seq_depth in c('1GB' , '10GB' , '100GB')) {
  for (seq_id in c('85%' , '90%' , '95%')) {   
    for(gene_fusion in unique(c(LongGF_BreakPoint$fusion.gene.id))){
      SB <- subset(Simulated_BreakPoint, fusion.gene.id == gene_fusion)
      gene_names <- str_split(gene_fusion, pattern = ":")[[1]] 
      gene1 <- gene_names[1]
      gene2 <- gene_names[2]
     
      SB1 <- as.integer(tail(str_split(subset(SB, alternative_ID == gene1)$exonEnds, ",")[[1]], 1))
      SB2 <- as.integer(head(str_split(subset(SB, alternative_ID == gene2)$exonStarts, ",")[[1]],1))
    
      gene3 <- gene_names[3]
      SB3 <- NA
      SB4 <- NA
      if(is.na(gene3) == FALSE){
        SB3 <- as.integer(str_split(subset(SB, alternative_ID == gene2)$exonEnds, ",")[[1]])
        SB4 <- as.integer(str_split(subset(SB, alternative_ID == gene3)$exonStarts, ",")[[1]])
      } 
      L <- subset(LongGF_BreakPoint, fusion.gene.id== gene_fusion & depth == seq_depth & Sequence_Identity == seq_id)
      if(nrow(L) > 0){
      LB1 <- as.numeric(L$breakpoint1) - SB1
      LB2 <- as.numeric(L$breakpoint2) - SB2
      }else{
        LB1 <- "NA"
        LB2 <- "NA"
        }
      new_row <-  data.frame(seq_depth, seq_id, gene_fusion, LB1, LB2)
      colnames(new_row) <- colnames(LongGF_Breakpoint_Accuracy)
      LongGF_Breakpoint_Accuracy <- rbind(LongGF_Breakpoint_Accuracy, new_row)
  }
 } 
}

FusionSeeker_Breakpoint_Accuracy <- data.frame(matrix(nrow = 0, ncol = length(c('seq_depth', 'seq_id', 'gene_fusion', 'FSB1', 'FSB2'))))
colnames(FusionSeeker_Breakpoint_Accuracy) <- c('seq_depth', 'seq_id', 'gene_fusion', 'FSB1', 'FSB2')
for (seq_depth in c('1GB' , '10GB' , '100GB')) {
  for (seq_id in c('85%' , '90%' , '95%')) {   
    for(gene_fusion in unique(c(FusionSeeker_BreakPoint$fusion.gene.id))){
      SB <- subset(Simulated_BreakPoint, fusion.gene.id == gene_fusion)
      FS <- subset(FusionSeeker_BreakPoint, fusion.gene.id == gene_fusion & depth == seq_depth & Sequence_Identity == seq_id)
      gene_names <- str_split(gene_fusion, pattern = ":")[[1]] 
      gene1 <- gene_names[1]
      gene2 <- gene_names[2]
     
      SB1 <- as.integer(tail(str_split(subset(SB, alternative_ID == gene1)$exonEnds, ",")[[1]], 1))
      SB2 <- as.integer(head(str_split(subset(SB, alternative_ID == gene2)$exonStarts, ",")[[1]],1))
    
      gene3 <- gene_names[3]
      SB3 <- NA
      SB4 <- NA
      if(is.na(gene3) == FALSE){
        SB3 <- as.integer(str_split(subset(SB, alternative_ID == gene2)$exonEnds, ",")[[1]])
        SB4 <- as.integer(str_split(subset(SB, alternative_ID == gene3)$exonStarts, ",")[[1]])
      }
      if(nrow(FS) > 0){
      FSB1 <- as.numeric(FS$Breakpoint1) - SB1
      FSB2 <- as.numeric(FS$Breakpoint2) - SB2
      }else{
        FSB1 <- "NA"
        FSB2 <- "NA"
      }
      new_row <-  data.frame(seq_depth, seq_id, gene_fusion, FSB1, FSB2)
      colnames(new_row) <- colnames(FusionSeeker_Breakpoint_Accuracy)
       FusionSeeker_Breakpoint_Accuracy <- rbind(FusionSeeker_Breakpoint_Accuracy, new_row)
      }
     } 
    }

Breakpoint_Accuracy <- full_join(FusionSeeker_Breakpoint_Accuracy, LongGF_Breakpoint_Accuracy) %>% full_join(Genion_Breakpoint_Accuracy) %>% full_join(JAFFAL_BreakPoint_Accuracy )

Breakpoint_Accuracy[is.na(Breakpoint_Accuracy)] <- "NA"

Breakpoint_Accuracy <- Breakpoint_Accuracy %>% filter(!if_all(FSB1:JB2, ~ .x == "NA")) %>%
  pivot_longer(cols = FSB1:JB2, names_to =  "Tool_Breakpoint", values_to = "Distance_from_Simulated_Breakpoint") %>%
  mutate(Algorithm = case_when(
    grepl("^GB", Tool_Breakpoint) ~ "Genion",
    grepl("^JB", Tool_Breakpoint) ~ "JAFFAL",
    grepl("^LB", Tool_Breakpoint) ~ "LongGF",
    grepl("^FSB", Tool_Breakpoint) ~ "FusionSeeker"))%>%
  mutate(values_vector = strsplit(as.character(Distance_from_Simulated_Breakpoint), ","))%>%
  unnest(values_vector) %>%
 mutate(values_vector = as.numeric(values_vector)) %>%
  mutate(Breakpoint_number = case_when(
    grepl("B1", Tool_Breakpoint) ~ "Breakpoint_1",
    grepl("B2", Tool_Breakpoint) ~ "Breakpoint_2",
    grepl("B3", Tool_Breakpoint) ~ "Breakpoint_3",
    grepl("B4", Tool_Breakpoint) ~ "Breakpoint_4"))

Breakpoint_Accuracy <-filter(Breakpoint_Accuracy, Distance_from_Simulated_Breakpoint != "NA") %>% distinct()
#some breakpoints are before the value and have a negative value, this we transform the values to absolute distances
absolute_Breakpoint_Accuracy <- Breakpoint_Accuracy %>%
  mutate(values_vector = abs(values_vector))
```

Make Figure 4 & Supplementary Figure 13
```{r}
library(ggforce)
library(ggpubr)
absolute_Breakpoint_Accuracy$Distance_from_Simulated_Breakpoint <-
  as.numeric(as.character(absolute_Breakpoint_Accuracy$Distance_from_Simulated_Breakpoint))

#Statistical test: Check if data is normally distributed
breakpoint_stat_data <- filter(absolute_Breakpoint_Accuracy, seq_depth == "100GB", seq_id == "95%")
breakpoint_stat_data$Log1p_values <- log1p(breakpoint_stat_data$values_vector)
by(breakpoint_stat_data$Log1p_values,
   breakpoint_stat_data$Algorithm,
    shapiro.test)
breakpoint1_stat_data <- filter(absolute_Breakpoint_Accuracy, seq_depth == "100GB", seq_id == "95%", Breakpoint_number == "Breakpoint_1")
breakpoint1_stat_data$Log1p_values <- log1p(breakpoint1_stat_data$values_vector)
by(breakpoint1_stat_data$Log1p_values,
   breakpoint1_stat_data$Algorithm,
    shapiro.test)
breakpoint2_stat_data <- filter(absolute_Breakpoint_Accuracy, seq_depth == "100GB", seq_id == "95%", Breakpoint_number == "Breakpoint_2")
breakpoint2_stat_data$Log1p_values <- log1p(breakpoint2_stat_data$values_vector)
by(breakpoint2_stat_data$Log1p_values,
   breakpoint2_stat_data$Algorithm,
    shapiro.test)

ggplot(breakpoint_stat_data,
        aes(sample = Log1p_values)) +
     stat_qq() +
     stat_qq_line(color = "red") +
     facet_grid(Breakpoint_number ~ Algorithm) +
     theme_minimal() +
     labs(title = "Q–Q Plots by Algorithm and Breakpoint",
          x = "Theoretical Quantiles",
          y = "Sample Quantiles")
#Data is not normally distributed, so we will do a Kruskal Wallis Test followed by a pairwise test.
kruskal.test(Log1p_values ~ Algorithm, data = breakpoint_stat_data)
kruskal.test(Log1p_values ~ Algorithm, data = breakpoint1_stat_data)
kruskal.test(Log1p_values ~ Algorithm, data = breakpoint2_stat_data)

#Figure 4 ####
Figure4 <- ggplot(filter(absolute_Breakpoint_Accuracy, seq_depth == "100GB", seq_id == "95%"), 
       aes(y = Distance_from_Simulated_Breakpoint, x = Breakpoint_number, color = Algorithm)) +
  geom_boxplot(outlier.shape = NA) +
  geom_sina(maxwidth = 0.8)+
  theme_minimal()+
  theme(legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        panel.grid.minor = element_blank(),   # remove minor log lines
        panel.grid.major.x = element_blank(),  # optional: remove vertical major lines
        plot.title = element_text(size = 12),
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 12))+ 
  scale_y_continuous(transform = scales::pseudo_log_trans(base=10) , 
                     breaks = c(-1000000000, -100000000, -10000000, -1000000, -100000, -10000, -1000, -100, 
                                -10, -1, 
                                0, 
                                1, 10, 100, 1000, 10000, 100000, 1000000, 10000000, 100000000, 1000000000),
                     labels = c(expression(-10^9), expression(-10^8), expression(-10^7), expression(-10^6), expression(-10^5), expression(-10^4), expression(-10^3), expression(-10^2),
                                expression(-10^1), "-1", 
                                "0", 
                                "1", expression(10^1), 
                                expression(10^2), expression(10^3), expression(10^4), expression(10^5), expression(10^6), expression(10^7), expression(10^8), expression(10^9))
  )+   
  scale_x_discrete(labels = c("Breakpoint 1", "Breakpoint 2"))+
  labs(y = "Distance from simulated breakpoint",
       x= "")+
  stat_compare_means(method = "kruskal.test", aes(group = Algorithm), label.y = 8.5 , show.legend = FALSE)+
  geom_pwc(method = "dunn_test", p.adjust.method = "bonferroni", hide.ns = TRUE, label = "p.adj.signif",
           y.position = c(6.5, 7.5 , 5.5, 6.5 ,7.5 ), show.legend = FALSE)
 
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure4_paper.pdf", plot = Figure4, width = 297, height = 210, units = "mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure4_paper.png", plot = Figure4)
           
supp_Figure13 <-ggplot(filter(absolute_Breakpoint_Accuracy), aes(y = Distance_from_Simulated_Breakpoint, x = Breakpoint_number, color = Algorithm)) +
  geom_boxplot(outlier.shape = NA) +
  geom_sina()+
  theme_minimal()+ 
  scale_y_continuous(transform = scales::pseudo_log_trans(base=10) , 
                     breaks = c(-1000000000, -100000000, -10000000, -1000000, -100000, -10000, -1000, -100, 
                                -10, -1, 
                                0, 
                                1, 10, 100, 1000, 10000, 100000, 1000000, 10000000, 100000000, 1000000000),
                     labels = c(expression(-10^9), expression(-10^8), expression(-10^7), expression(-10^6), expression(-10^5), expression(-10^4), expression(-10^3), expression(-10^2),
                                expression(-10^1), "-1", 
                                "0", 
                                "1", expression(10^1), 
                                expression(10^2), expression(10^3), expression(10^4), expression(10^5), expression(10^6), expression(10^7), expression(10^8), expression(10^9))
  )+   
  scale_x_discrete(labels = c("Breakpoint 1", "Breakpoint 2"))+
  labs(y = "Absolute distance from simulated breakpoint", x= "")+
  stat_compare_means(method = "kruskal.test", aes(group = Algorithm), label.y = 8.5 , show.legend = FALSE, label = "p.format",)+
  geom_pwc(method = "dunn_test", p.adjust.method = "bonferroni", hide.ns = TRUE, label = "p.adj.signif",
           y.position = c(6.5, 7.5 , 5.5, 6.5 ,7.5 ), show.legend = FALSE)+
  facet_grid(seq_depth ~ seq_id)+ 
  theme(legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        panel.grid.minor = element_blank(),   # remove minor log lines
        panel.grid.major.x = element_blank(),  # optional: remove vertical major linesplot.title = element_text(size = 12),
        strip.text.x = element_text(size = 12),
        strip.text.y = element_text(size = 12),
        axis.text = element_text(size = 9.5),
        axis.title = element_text(size = 12))
ggsave("~/LongReadFusionCallerBenchmark/Figures/supp_Figure13.png", plot = supp_Figure13,  width = 300, height = 420 , unit ="mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/supp_Figure13.pdf", plot = supp_Figure13,  width = 297, height = 420 , unit ="mm")
```

