---
title: "Preparing data and read lengths"
author: "Ryley Dorney"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```



Define colours for each platform
```{r}
################################################################
# Define colors for other platforms
################################################################
platform_colours <- c(
  "Illumina.PCR_cDNA.B1" = "#FFA500",
  "Illumina.PCR_cDNA.B2" = "#CC8400",
  "PacBio.PCR_cDNA.B1"   = "#C80972",
  "PacBio.PCR_cDNA.B2"   = "#8B064F",
  "ONT.direct_cDNA.B1" = "#009ACD",
  "ONT.direct_cDNA.B2" = "#00688B",
  "ONT.direct_RNA.B1"  = "#00CDCD",
  "ONT.direct_RNA.B2"  = "#008B8B",
  "ONT.PCR_cDNA.B1" = "#1f77b4",
  "ONT.PCR_cDNA.B2" = "#15496e"
)
# Merge colors
color_map <- c(platform_colours)
```

Trimming Nanopore reads of their adaptors
```{bash}
~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Preprocessing/DoradoTrimAdaptor.sh
```

Filtering Nanopore Reads for lengths greater than 100bp, Quality is 8 or greater
```{bash}
bash ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Preprocessing/Filter_NanoporeReads_Chopper.sh
```

Trim adapter sequences and poly_g, and quality filter by Q15 or greater using fastp on Illumina Sequencing Data
```{bash}
bash ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Preprocessing/Fastp_trim_and_filter.sh
```

Using JAFFAL to find fusions in the long read Huh-7 Data, and JAFFA_direct to find fusions in the short-read Huh-7 Data

```{bash}
bash ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Fusions/RunJAFFAL.sh
bash ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Fusions/Run_JAFFAdirect.sh
```

Use Genion, FusionSeeker & LongGF to find more fusions
```{bash}
~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Fusions/Run_GFSLGF.sh
```

```{bash}
for input_file in /bioinformatics/ryley/Library_Benchmark/*.sorted.bam; do
  file_prefix=$(basename "$input_file" .bam)
samtools stats -@ 8 $input_file > ${file_prefix}.stats
done
multiqc /bioinformatics/ryley/Library_Benchmark/

```

```{bash}
# Create an read lengths data file
echo -e "file\tread_length\tcount" > ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/library_read_length_summary.tsv

# Loop through each *.stats file and extract RL lines
for file in /bioinformatics/ryley/Library_Benchmark/bam_and_sam_files/*stats*; do
  grep "^RL" "$file" | cut -f 2- | while read -r length count; do
    file_prefix=$(basename "$file" .stats )
    echo -e "${file_prefix}\t${length}\t${count}" >> ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/library_read_length_summary.tsv
  done
done
```

```{r}
Longlibrary_read_lengths <- read.delim("library_read_length_summary.tsv", header = TRUE)%>% 
  mutate(library_type = case_when(grepl('FBA22517|FBA22660', file)  ~ "direct_RNA", 
                                  grepl('FBA62655|FBA43334', file)  ~ "direct_cDNA", 
                                  grepl('FAZ|FLNCcDNA', file)  ~ "PCR_cDNA"), 
         RNA_sample = case_when(grepl('FAZ80247|FBA62655|FBA22660|PB1_FLNCcDNA', file)  ~ "Huh7_p9_8_7_B1",
                                grepl('FBA22517|FAZ83542|FBA43334|PB2_FLNCcDNA', file) ~ "Huh7_p9_8_7_B2"), 
         Platform = case_when(grepl('FBA22517|FBA22660|FAZ83542|FBA43334|FBA62655|FAZ80247', file)  ~  "ONT",
                              grepl('PB', file)  ~  "PacBio"),
         Alignment = case_when(grepl('transcriptome', file)  ~  "transcriptome", 
                               grepl('stranded', file)  ~  "stranded_genome",
                               TRUE ~ "genome"))
```

