---
title: "Preparing_data"
author: "Ryley Dorney"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Trimming Nanopore reads of their adaptors
```{bash}
#direct cDNA nanopore seq use the  LSK-114 kit
direct_cDNA_Directory=$(/bioinformatics/ryley/Library_Benchmark/raw_files/Huh7_Library_Comparison_direct_cDNA/)

for reads in $(${direct_cDNA_Directory}*/*/F*fastq.gz); do
output_prefix=$(basename ${reads} .fastq.gz)
/opt/dorado-0.9.0-linux-x64/bin/dorado trim --sequencing-kit SQK_LSK_114 --emit-fastq ${reads} > ${direct_cDNA_Directory}_dorado_trimmed_${output_prefix}.fastq
gzip ${direct_cDNA_Directory}_dorado_trimmed_${output_prefix}.fastq &
done

#PCR cDNA nanopore seq use the LSK-114 kit
PCR_cDNA_Directory=$(/bioinformatics/ryley/Library_Benchmark/raw_files/Huh-7_Library_Comparision_PCR-cDNA/)

for reads in $(${PCR_cDNA_Directory}*/*/F*fastq.gz); do
output_prefix=$(basename ${reads} .fastq.gz)
/opt/dorado-0.9.0-linux-x64/bin/dorado trim --sequencing-kit SQK_PCS114 --emit-fastq ${reads} > ${PCR_cDNA_Directory}_dorado_trimmed_${output_prefix}.fastq
gzip ${PCR_cDNA_Directory}_dorado_trimmed_${output_prefix}.fastq &
done

#direct RNA seq trims reads and primers while it is sequencing, so we do not need to trim here
```

Filtering Nanopore Reads for lengths greater than 100bp, Quality is 8 or greater
```{bash}
bash ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Preprocessing/Filter_NanoporeReads_Chopper.sh
```
Trim adapter sequences and poly_g, and quality filter by Q15 or greater using fastp on Illumina Sequencing Data
```{bash}
bash ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Preprocessing/Fastp_trim_and_filter.sh
```

Using JAFFAL to find fusions in the long read Huh-7 Data, and JAFFA_direct to find fusions in the short-read Huh-7 Data
```{bash}
bash ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Fusions/RunJAFFAL.sh
bash ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Fusions/Run_JAFFAdirect.sh
```
Use Genion, FusionSeeker & LongGF to find more fusions??
```{bash}

```


```{bash}
#conda create --name STAR_env bioconda::star
##Need to choose a different aligner for aligning short reads
#/opt/JAFFA-version-2.3/tools/bin/bowtie2 --very-fast -p 10 -x /opt/JAFFA-version-2.3/hg38_genCode47 -1 /bioinformatics/ryley/Library_Benchmark/AGRF_Sequencing_Data/Illumina_NovaSeq/fastp_trimmed/Huh7_B1_22T7FFLT3_CATTAGTGCG-TGACTACATA_R1.fastq.gz -2 /bioinformatics/ryley/Library_Benchmark/AGRF_Sequencing_Data/Illumina_NovaSeq/fastp_trimmed/Huh7_B1_22T7FFLT3_CATTAGTGCG-TGACTACATA_R2.fastq.gz -S /bioinformatics/ryley/Library_Benchmark/Huh7_B1_Illumina.sam
#samtools view -@ 4 -bS /bioinformatics/ryley/Library_Benchmark/Huh7_B1_Illumina.sam | samtools sort -@ 4 -o /bioinformatics/ryley/Library_Benchmark/Huh7_B1_Illumina.sorted.bam &

#/opt/JAFFA-version-2.3/tools/bin/bowtie2 --very-fast -p 10 -x /opt/JAFFA-version-2.3/hg38_genCode47 -1 /bioinformatics/ryley/Library_Benchmark/AGRF_Sequencing_Data/Illumina_NovaSeq/fastp_trimmed/Huh7_B2_22T7FFLT3_ACGGTCAGGA-CGGCCTCGTT_R1.fastq.gz -2 /bioinformatics/ryley/Library_Benchmark/AGRF_Sequencing_Data/Illumina_NovaSeq/fastp_trimmed/Huh7_B2_22T7FFLT3_ACGGTCAGGA-CGGCCTCGTT_R2.fastq.gz -S /bioinformatics/ryley/Library_Benchmark/Huh7_B2_Illumina.sam
#samtools view -@ 4 -bS /bioinformatics/ryley/Library_Benchmark/Huh7_B2_Illumina.sam | samtools sort -@ 4 -o /bioinformatics/ryley/Library_Benchmark/Huh7_B2_Illumina.sorted.bam 

#wget https://ftp.ebi.ac.uk/pub/databases/gencode/Gencode_human/release_47/gencode.v47.annotation.gtf.gz -O ~/reference_files/gencode.v47.annotation.gtf.gz
#gunzip ~/reference_files/gencode.v47.annotation.gtf.gz
```

```{bash Aligning Reads}
#need to rerun
#align long reads to genome
for input_file in /bioinformatics/ryley/Library_Benchmark/Huh-7_Library_Comparison/*/*/*.fastq.gz /bioinformatics/ryley/Library_Benchmark/Huh-7_Library_Comparision_PCR-cDNA/*/*/*.fastq.gz /bioinformatics/ryley/Library_Benchmark/Huh7_Library_Comparison_direct_cDNA/*/*/*.fastq.gz /bioinformatics/ryley/Library_Benchmark/AGRF_Sequencing_Data/PacBio_Revio_full_length_cDNA/*.fastq.gz; do
  file_prefix=$(basename "$input_file" .fastq.gz); /opt/JAFFA-version-2.3/tools/bin/minimap2 -t 4 -ax splice /opt/JAFFA-version-2.3/hg38.fa $input_file | samtools view -@ 8 -b | samtools sort -@ 8 -o /bioinformatics/ryley/Library_Benchmark/${file_prefix}.sorted.bam 
done

#wget https://ftp.ebi.ac.uk/pub/databases/gencode/Gencode_human/release_47/gencode.v47.transcripts.fa.gz 
#align stranded reads against genome
for input_file in /bioinformatics/ryley/Library_Benchmark/Huh-7_Library_Comparison/*/*/*.fastq.gz; do
  file_prefix=$(basename "$input_file" .fastq.gz)
  /opt/JAFFA-version-2.3/tools/bin/minimap2 -t 4 -ax splice -uf -k14 /opt/JAFFA-version-2.3/hg38.fa $input_file | samtools view -@ 8 -b | samtools sort -@ 8 -o /bioinformatics/ryley/Library_Benchmark/${file_prefix}.stranded.sorted.bam 
done

wait
#align  reads against transcriptome
for input_file in /bioinformatics/ryley/Library_Benchmark/Huh-7_Library_Comparision_PCR-cDNA/*/*/*.fastq.gz /bioinformatics/ryley/Library_Benchmark/Huh7_Library_Comparison_direct_cDNA/*/*/*.fastq.gz /bioinformatics/ryley/Library_Benchmark/AGRF_Sequencing_Data/PacBio_Revio_full_length_cDNA/*.fastq.gz /bioinformatics/ryley/Library_Benchmark/Huh-7_Library_Comparison/*/*/*.fastq.gz; do
  file_prefix=$(basename "$input_file" .fastq.gz)
  /opt/JAFFA-version-2.3/tools/bin/minimap2 -t 4 -ax lr:hq ~/reference_files/gencode.v47.transcripts.fa.gz $input_file | samtools view -@ 8 -b | samtools sort -@ 8 -o /bioinformatics/ryley/Library_Benchmark/${file_prefix}.transcriptome.sorted.bam 
done

samtools index -@ 8 -M -b /bioinformatics/ryley/Library_Benchmark/*.sorted.bam 

```


```{bash}
for input_file in /bioinformatics/ryley/Library_Benchmark/*.sorted.bam; do
  file_prefix=$(basename "$input_file" .bam)
samtools stats -@ 8 $input_file > ${file_prefix}.stats
done
multiqc /bioinformatics/ryley/Library_Benchmark/

```
```{bash}
# Create an read lengths data file
echo -e "file\tread_length\tcount" > ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/library_read_length_summary.tsv

# Loop through each *.stats file and extract RL lines
for file in /bioinformatics/ryley/Library_Benchmark/bam_and_sam_files/*stats*; do
  grep "^RL" "$file" | cut -f 2- | while read -r length count; do
    file_prefix=$(basename "$file" .stats )
    echo -e "${file_prefix}\t${length}\t${count}" >> ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/library_read_length_summary.tsv
  done
done
```

```{r}
Longlibrary_read_lengths <- read.delim("library_read_length_summary.tsv", header = TRUE)%>% 
  mutate(library_type = case_when(grepl('FBA22517|FBA22660', file)  ~ "direct_RNA", 
                                  grepl('FBA62655|FBA43334', file)  ~ "direct_cDNA", 
                                  grepl('FAZ|FLNCcDNA', file)  ~ "PCR_cDNA"), 
         RNA_sample = case_when(grepl('FAZ80247|FBA62655|FBA22660|PB1_FLNCcDNA', file)  ~ "Huh7_p9_8_7_B1",
                                grepl('FBA22517|FAZ83542|FBA43334|PB2_FLNCcDNA', file) ~ "Huh7_p9_8_7_B2"), 
         Platform = case_when(grepl('FBA22517|FBA22660|FAZ83542|FBA43334|FBA62655|FAZ80247', file)  ~  "ONT",
                              grepl('PB', file)  ~  "PacBio"),
         Alignment = case_when(grepl('transcriptome', file)  ~  "transcriptome", 
                               grepl('stranded', file)  ~  "stranded_genome",
                               TRUE ~ "genome"))
```

