---
title: "Preparing data and read lengths"
author: "Ryley Dorney"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Define colours for each platform and algorithms
```{r}
################################################################
# Define colors for other platforms
################################################################
platformlibsamp_colours <- c(
  "Illumina.PCR_cDNA.B1" = "#FFA500",
  "Illumina.PCR_cDNA.B2" = "#CC8400",
  "PacBio.PCR_cDNA.B1"   = "#C80972",
  "PacBio.PCR_cDNA.B2"   = "#8B064F",
  "ONT.PCR_cDNA.B1" = "#1f77b4",
  "ONT.PCR_cDNA.B2" = "#15496e",
  "ONT.direct_cDNA.B1" = "#009ACD",
  "ONT.direct_cDNA.B2" = "#00688B",
  "ONT.direct_RNA.B1"  = "#00CDCD",
  "ONT.direct_RNA.B2"  = "#008B8B"
)

# Merge colors
platformlibsamp_colourmap <- c(platformlibsamp_colours)

###########
# template to order labels
#df$full_label <- factor(df$full_label, levels = c("Illumina.PCR_cDNA.B1", "Illumina.PCR_cDNA.B2", "PacBio.PCR_cDNA.B1", "PacBio.PCR_cDNA.B2", "ONT.PCR_cDNA.B1","ONT.PCR_cDNA.B2", "ONT.direct_cDNA.B1", "ONT.direct_cDNA.B2", "ONT.direct_RNA.B1", "ONT.direct_RNA.B2")) 
#df$Platform <- factor(df$Platform, levels = c("Illumina", "PacBio", "ONT")) 
#df$library_type <- factor(df$library_type, levels = c("PCR_cDNA", "direct_cDNA", "direct_RNA")) 
###########

scales::show_col(pal_npg("nrc")(7))
Algorithm_Colours <- c(
  "CTAT-LR-Fusion"="#E64B35FF",
  "FusionSeeker"="#F39B7FFF",
  "Genion" ="#4DBBD5FF",
  "GFSeeker" ="#3C5488FF",
  "JAFFAL" ="#00A087FF",
  "JAFFA-direct" ="#91D1C2FF",
  "LongGF" ="#8491B4FF"
  )
Alg_colour_map <- c(Algorithm_Colours)
```

Trimming Nanopore reads of their adaptors
```{bash}
bash ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/PreprocessingReads/DoradoTrimAdaptor.sh
```

Filtering Nanopore Reads for lengths greater than 100bp, Quality is 8 or greater
```{bash}
bash ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/PreprocessingReads/Filter_NanoporeReads_Chopper.sh
```

Trim adapter sequences and poly_g, and quality filter by Q15 or greater using fastp on Illumina Sequencing Data
```{bash}
bash ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/PreprocessingReads/Fastp_trim_and_filter.sh
```

Quality control
```{bash}
#Read Lengths and Quality
#Nanoplot
pip install nanoplot

NanoPlot -t 10 --bam /bioinformatics/ryley/Gencode44/Huh7_Library/FusionSeeker/*sorted.bam --minlength 100 -o Huh7_bam_summary_plots

NanoPlot -t 10 --summary/bioinformatics/ryley/Library_Benchmark/raw_files/Huh*/*/*/sequencing*txt -o ONT_Only_summary-plots


#LongReadSum
```


```{bash}
for input_file in /bioinformatics/ryley/Library_Benchmark/*.sorted.bam; do
  file_prefix=$(basename "$input_file" .bam)
samtools stats -@ 8 $input_file > ${file_prefix}.stats
done
multiqc /bioinformatics/ryley/Library_Benchmark/

```

```{bash}
# Create an read lengths data file
echo -e "file\tread_length\tcount" > ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/library_read_length_summary.tsv

# Loop through each *.stats file and extract RL lines
for file in /bioinformatics/ryley/Library_Benchmark/bam_and_sam_files/*stats*; do
  grep "^RL" "$file" | cut -f 2- | while read -r length count; do
    file_prefix=$(basename "$file" .stats )
    echo -e "${file_prefix}\t${length}\t${count}" >> ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/library_read_length_summary.tsv
  done
done
```

```{r}
Longlibrary_read_lengths <- read.delim("library_read_length_summary.tsv", header = TRUE)%>% 
  mutate(library_type = case_when(grepl('FBA22517|FBA22660', file)  ~ "direct_RNA", 
                                  grepl('FBA62655|FBA43334', file)  ~ "direct_cDNA", 
                                  grepl('FAZ|FLNCcDNA', file)  ~ "PCR_cDNA"), 
         RNA_sample = case_when(grepl('FAZ80247|FBA62655|FBA22660|PB1_FLNCcDNA', file)  ~ "Huh7_p9_8_7_B1",
                                grepl('FBA22517|FAZ83542|FBA43334|PB2_FLNCcDNA', file) ~ "Huh7_p9_8_7_B2"), 
         Platform = case_when(grepl('FBA22517|FBA22660|FAZ83542|FBA43334|FBA62655|FAZ80247', file)  ~  "ONT",
                              grepl('PB', file)  ~  "PacBio"),
         Alignment = case_when(grepl('transcriptome', file)  ~  "transcriptome", 
                               grepl('stranded', file)  ~  "stranded_genome",
                               TRUE ~ "genome"))
```

Identify Fusions
Find fusions in the long read Huh-7 Data, and use JAFFA_direct to find fusions in the short-read Huh-7 Data
```{bash}
bash ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Fusions/runtools/RunJAFFAL.sh
bash ~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Fusions/runtools/Run_JAFFAdirect.sh

~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Fusions/runtools/Run_GFSLGF.sh
```

Pre-process Fusion Data
```{r}
source("~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Fusions/FormattingandPreprocessFusions.R")
source("~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Fusions/label_fusions/AssignEnsemblID.R")
```

Check FusionTypes
```{r}
source("~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Fusions/label_fusions/CheckFalseFusions.R")
source("~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Fusions/label_fusions/LabelSenseAntisense.R")
source("~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Fusions/label_fusions/LabelFusionTypes.R")
```
Labelled Fusions can be found in /bioinformatics/ryley/Gencode44/Huh7_Library/
Check Known Fusion
```{r}
#make space by clearing the heavier objects
rm(c(JAFFAL_mitocheck_annotated, JAFFAL_mitocheck, JAFFAL_sensecheck_annotated, Huh7_JAFFAL_ensemblID, Huh7_JAFFAL))
gc()

source("~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Fusions/label_fusions/CheckKnownFusions.R")
```

