---
title: "simulated_data_analysis"
author: "Ryley Dorney"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggh4x)
library(biomaRt)
library(UpSetR)
library(ComplexUpset)
library(patchwork)
```
```{r}
Analysis_output_folder <- "/bioinformatics/ryley/Algorithm_Benchmark/benchmark_data_analysis/Analysis/benchmark_fusion_results/"

```

The 'biomarRt' package is used to help identify corresponding gene ids to gene names and vice versa. Whilst gencode version 43 was used for this study, we also check the current ensembl version for where ensembl IDs may have been assigned a gene name in the newer version for comparing with literature.
```{r}
ensembl_current <-  useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl")
#the below ensembl versions best correlate to gencode version 43
ensemblv109 <- useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl", version = 109)
ensemblv110 <- useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl", version = 110)
ensembljul2023 <- useMart("ensembl", dataset = "hsapiens_gene_ensembl", host = "https://jul2023.archive.ensembl.org")
```

Here we read in the data table containing information about the fusions generated with Fusim to be spiked in. We then modify the table to have separate columns for partner genes. I then use biomart to find the corresponding gene names/ids. This helps find potential alternative names/ids for gene fusions recalled.
```{r}
simulated_fusions <- read.table(file = "~/benchmark_data_analysis/FUSIM_benchmark_fusions.txt", header=TRUE, sep= "\t")
#convert from txt to tsv and save file
write_tsv(simulated_fusions, file="~/simulated_fusion_info.tsv")

#seperate the fusion gene IDs into individual gene IDs for downstream annotation steps
Simulated_Fusion_Info <- simulated_fusions %>% separate(fusionGene, into= c('V1','V2','V3'), sep = "-", remove = FALSE) %>% mutate(across(fusionGene, ~ gsub("-", "::", .))) %>% mutate(geneName_ID = geneName) %>% mutate(across(c(V1, V2, V3, geneName), ~ gsub("\\..*","", .)))%>% separate(fusionGene, into= c('Gene1_vID','Gene2_vID','Gene3_vID'), sep = "::", remove = FALSE)
# Retrieve gene names for the corresponding Ensembl IDs. These mappings are essential for accurate comparison between simulated and experimentally detected fusions across different annotation systems.
Gene1_Name<-rbind(getBM(attributes = c("external_gene_name", "ensembl_gene_id"),
                             filters = "ensembl_gene_id",
                             values = (Simulated_Fusion_Info$geneName),
                             mart = ensembljul2023),
                      getBM(attributes = c("external_gene_name", "ensembl_gene_id"),
                             filters = "ensembl_gene_id",
                             values = (Simulated_Fusion_Info$geneName),
                             mart = ensemblv109),
                      getBM(attributes = c("external_gene_name", "ensembl_gene_id"),
                             filters = "ensembl_gene_id",
                             values = (Simulated_Fusion_Info$geneName),
                             mart = ensemblv110)) %>% unique()
External_Gene_Name<-rbind(getBM(attributes = c("external_gene_name", "ensembl_gene_id", "ensembl_gene_id_version", "chromosome_name"),
                             filters = "external_gene_name",
                             values = unique(Gene1_Name$external_gene_name),
                             mart = ensembljul2023), 
                          getBM(attributes = c("external_gene_name", "ensembl_gene_id", "ensembl_gene_id_version", "chromosome_name"),
                             filters = "external_gene_name",
                             values = unique(Gene1_Name$external_gene_name),
                             mart = ensemblv109), 
                          getBM(attributes = c("external_gene_name", "ensembl_gene_id", "ensembl_gene_id_version", "chromosome_name"),
                             filters = "external_gene_name",
                             values = unique(Gene1_Name$external_gene_name),
                             mart = ensemblv110)) %>% unique()
External_Gene_Name$chromosome_name <- paste0("chr",External_Gene_Name$chromosome_name)

Alternative_Names_Gene_Fusion <- rbind(getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name"),
                             filters = "external_gene_name",
                             values = (External_Gene_Name$external_gene_name),
                             mart = ensembljul2023),
                      getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name"),
                             filters = "external_gene_name",
                             values = (External_Gene_Name$external_gene_name),
                             mart = ensemblv109),
                      getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name"),
                             filters = "external_gene_name",
                             values = (External_Gene_Name$external_gene_name),
                             mart = ensemblv110)) %>% unique()
Alternative_Names_Gene_Fusion$chromosome_name <- paste0("chr", Alternative_Names_Gene_Fusion$chromosome_name)

Sim_Info_Names <- left_join(Simulated_Fusion_Info, Alternative_Names_Gene_Fusion[c(1,2)], by = c("geneName"="ensembl_gene_id"))%>% 
  left_join(Alternative_Names_Gene_Fusion, by = c("external_gene_name", "chrom" = "chromosome_name")) %>% mutate(
    alternative_ID = coalesce(ensembl_gene_id, geneName)) %>% select(-c(ensembl_gene_id))

#create a dataframe inclusive of alternative fusion gene names as gene names can be variable.  
Simulated_Fusion_Info_2<-Sim_Info_Names %>%
  mutate(Gene1_alternative_ID = map(Gene1_vID, ~ alternative_ID[which(geneName_ID == .x)]),
    Gene2_alternative_ID = map(Gene2_vID, ~ alternative_ID[which(geneName_ID == .x)]),
    Gene3_alternative_ID = map(Gene3_vID, ~ if (!is.na(.x)) alternative_ID[which(geneName_ID == .x)] else NA_character_)) %>%
  unnest_longer(Gene1_alternative_ID) %>%
  unnest_longer(Gene2_alternative_ID) %>%
  unnest_longer(Gene3_alternative_ID, keep_empty = TRUE)%>%
  unite("fusion.gene.id" , c(Gene1_alternative_ID, Gene2_alternative_ID, Gene3_alternative_ID), sep = ":", remove= FALSE, na.rm = TRUE) %>%
   select(-c(alternative_ID)) %>% unique() %>%
  unite("original.fusion.gene.id", c(V1, V2), sep=":", remove= FALSE, na.rm = TRUE)

Simulated_Fusion_Info_3<-Sim_Info_Names %>%
  mutate(Gene1_alternative_ID = map(Gene1_vID, ~ alternative_ID[which(geneName_ID == .x)]),
    Gene2_alternative_ID = map(Gene2_vID, ~ alternative_ID[which(geneName_ID == .x)]),
    Gene3_alternative_ID = map(Gene3_vID, ~ if (!is.na(.x)) alternative_ID[which(geneName_ID == .x)] else NA_character_)) %>%
  unnest_longer(Gene1_alternative_ID) %>%
  unnest_longer(Gene2_alternative_ID) %>%
  unnest_longer(Gene3_alternative_ID, keep_empty = TRUE)%>%
  unite("fusion.gene.id" , c(Gene1_alternative_ID, Gene2_alternative_ID, Gene3_alternative_ID), sep = ":", remove= FALSE, na.rm = TRUE) %>%
  unique() %>%
  unite("original.fusion.gene.id", c(V1, V2), sep=":", remove= FALSE, na.rm = TRUE)
```

LongGF suggests to extract the called fusions form the output log using 'grep SumGF'.
```{bash}
pwd
#change to folder contain LongGF fusion calls
cd /bioinformatics/ryley/Algorithm_Benchmark/Adapter_porechop_trimmed/LongGF/
#grep the results from the calls and create a tsv in the home directory
grep SumGF LongGF*Spiked*.log > /home/ryleyd/LongReadFusionCallerBenchmark/LongGF_Spiked.tsv
grep SumGF *Adapter_Simulated*.log > /home/ryleyd/LongReadFusionCallerBenchmark/LongGF_Negative_Control.tsv
#return to home directory
cd ~
```
Read in the LongGF fusions into RStudio and separate components of output by partner genes. 
```{r}
#LongGF data wrangling
#list files
my_files <- list.files(path = "/home/ryleyd/LongReadFusionCallerBenchmark", pattern = "LongGF", full.names=TRUE)
all_tsv <-lapply(my_files, read.table)
names <- gsub(".tsv","", list.files(path = "/home/ryleyd/LongReadFusionCallerBenchmark", pattern = "LongGF",full.names = FALSE),fixed = TRUE)
#assign each longGF file
for(i in names){
  assign(i, read.table(paste0("/home/ryleyd/LongReadFusionCallerBenchmark/", i, ".tsv")))
}
#seperate columns to more easily identify information correlating to each gene partner
LongGF_Sim <- rbind(LongGF_Negative_Control, LongGF_Spiked)%>%
      mutate(Source = basename(V1)) %>% mutate(control = ifelse(grepl("Spiked", Source), "positive", "negative"))%>% 
  separate(V2, into = c("Gene1", "Gene2"), sep = ":", remove = FALSE) %>% 
  separate(V4, into = c("chromosome1", "breakpoint1"), sep = ":", remove = FALSE) %>% 
  separate(V5, into = c("chromosome2", "breakpoint2"), sep = ":", remove = FALSE) #V2 = fusion gene name :
ogLongGF <- LongGF_Sim
```
BiomaRt is then used to find the corresponding gene names/ids for fusions called by LongGF.
```{r}
#gets ensembl gene ID and check it matches the putative chromosome 
Gene_Name<-rbind (getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name"),
                             filters = "external_gene_name",
                             values = unique(c(LongGF_Sim$Gene1, LongGF_Sim$Gene2)),
                             mart = ensembljul2023),
                  getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name"),
                             filters = "external_gene_name",
                             values = unique(c(LongGF_Sim$Gene1, LongGF_Sim$Gene2)),
                             mart = ensemblv109)) %>% unique()
Gene_Name$chromosome_name <- paste0("chr", Gene_Name$chromosome_name)

  
  LongGF_Sim_ensembl <- LongGF_Sim %>%
  left_join(Gene_Name, by = c('Gene1' = 'external_gene_name')) %>%
  left_join(Gene_Name, by = c('Gene2' = 'external_gene_name')) %>%
  mutate(
    ensembl_gene1 = ifelse(is.na(ensembl_gene_id.x) | nchar(ensembl_gene_id.x) == 0, Gene1, ensembl_gene_id.x),
    ensembl_gene2 = ifelse(is.na(ensembl_gene_id.y) | nchar(ensembl_gene_id.y) == 0, Gene2, ensembl_gene_id.y)
  ) %>%
  unite("fusion.gene.id", c(ensembl_gene1, ensembl_gene2), sep = ":", remove = FALSE, na.rm = TRUE) %>%
  # Create a filtering condition using case_when
  filter(case_when(
      is.na(chromosome_name.x) ~ TRUE, # Keep if chromosome_name.x is NA
      chromosome_name.x == chromosome1 ~ TRUE, # Keep if chromosome names match
      chromosome_name.x == "chrMT" & chromosome1 == "chrM" ~ TRUE, # Allow chrMT/chrM equivalence
      grepl("PATCH|HSCHR", chromosome_name.x) & !grepl("^chr([1-9]|1[0-9]|2[0-2]|X|Y|M)$", chromosome1) ~ TRUE, 
      TRUE ~ FALSE # Remove rows that don't meet the above conditions
    )) %>%
  filter(case_when(
      is.na(chromosome_name.y) ~ TRUE,
      chromosome_name.y == chromosome2 ~ TRUE,
      chromosome_name.y == "chrMT" & chromosome2 == "chrM" ~ TRUE,
      grepl("PATCH|HSCHR", chromosome_name.y) & !grepl("^chr([1-9]|1[0-9]|2[0-2]|X|Y|M)$", chromosome2) ~ TRUE,
      TRUE ~ FALSE # Remove rows that don't meet the above conditions
    )) 
```
check if LongGF called any fusions that were spiked in
```{r}
Annot_LongGF_Sim <- LongGF_Sim_ensembl %>% 
  left_join(Simulated_Fusion_Info_2, by = c("fusion.gene.id"), relationship = "many-to-many")
```
Assign label to false fusions or partially recalled fusions by LongGF
```{r}
Annot_LongGF_Sim <- Annot_LongGF_Sim %>%
  group_by(V2.x) %>%
  mutate(fusionType = case_when(
    is.na(fusionType) ~ paste0("chromosomal_misalignment:",first(na.omit(fusionType))),  # Only update NA values
    .default = fusionType  # Keep existing values unchanged
  ))%>% ungroup()%>%
  replace_with_na(replace = list(fusionType = "chromosomal_misalignment:NA")) 

Annot_LongGF_Sim <- LongGF_Sim_ensembl %>% left_join(Simulated_Fusion_Info_2, by = c("fusion.gene.id"), relationship = "many-to-many") %>%
  group_by(V2.x) %>%
  mutate(fusionType = case_when(
    is.na(fusionType) ~ paste0("chromosomal_misalignment:",first(na.omit(fusionType))),  # Only update NA values
    .default = fusionType  # Keep existing values unchanged
  ))%>% ungroup()%>%
  replace_with_na(replace = list(fusionType = "chromosomal_misalignment:NA"))  
Annot_LongGF_Sim$fusionType <- mapply(function(g1, g2, current_type, chr1, chr2, gene_name1, gene_name2) {
   # Check if the current fusionType is empty
  if (current_type == "" || is.na(current_type)) {
    #check for truncated_tri_fusions
    if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g1, ":", g2)) |
            str_detect(paste(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V1, subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V3, sep = ":"), paste0(g1, ":", g2)))) {
      return("truncated_tri_fusion")
      #check for reverse order fusions
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1))|
                  str_detect(paste(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V1, subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V3, sep = ":"), paste0(g2, ":", g1)))){
      return("reverse_order:truncated_tri_fusion") 
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1)))){
       matching_row <- subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusionType[which(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1)))]
      return(paste("reverse_order:", matching_row[1])) 
       #check for chromosomal misalignment
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$original.fusion.gene.id, paste0(g1, ":", g2)))){
       matching_row <- subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusionType[which(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$original.fusion.gene.id, paste0(g1, ":", g2)))]
      return(paste("chromosomal_misalignment:", matching_row[1])) 
       #check for false fusions
    } else if ((grepl("chrM:", chr1, ignore.case = TRUE) & !grepl("chrM:", chr2, ignore.case = TRUE)) | (!grepl("chrM:", chr1, ignore.case = TRUE) & grepl("chrM:", chr2, ignore.case = TRUE))){
      return("false_fusion:mitochondrial_genomic") 
    }else if (grepl("chrM:", chr1, ignore.case = TRUE) & grepl("chrM:", chr2, ignore.case = TRUE)){
      return("false_fusion:mitochondrial") 
    }else if ((gene_name1 == gene_name2) & (chr1 != chr2)){
      return("false_fusion:self_misalignment") 
    }else {
      return("false_fusion")
    }
  }
  return(current_type)
}, Annot_LongGF_Sim$ensembl_gene1, Annot_LongGF_Sim$ensembl_gene2, Annot_LongGF_Sim$fusionType, Annot_LongGF_Sim$chromosome1, Annot_LongGF_Sim$chromosome2, Annot_LongGF_Sim$Gene1, Annot_LongGF_Sim$Gene2)
```
Repeat the same above pre-processing steps for FusionSeeker
```{r}
#FusionSeeker_Sim # Gene1 = Gene ID 1, Gene2 = Gene ID 2

myfiles<-list.files(path = "/bioinformatics/ryley/Algorithm_Benchmark/Adapter_porechop_trimmed/FusionSeeker", pattern = "confident_genefusion.txt", full.names = TRUE, recursive = TRUE)
FusionSeeker_Sim <-  do.call(rbind, lapply(myfiles, function(filename) {
  read.table(filename, header = TRUE) %>%
    mutate(Source = basename(dirname(filename)))%>% mutate(control = ifelse(grepl("Spiked", Source), "positive", "negative"))
}))
ogFusionSeeker <- do.call(rbind, lapply(myfiles, function(filename) {
  read.table(filename, header = TRUE) %>%
    mutate(Source = basename(dirname(filename)))%>% mutate(control = ifelse(grepl("Spiked", Source), "positive", "negative"))
}))

FusionSeeker_Sim <- FusionSeeker_Sim %>%
  unite("fusionGene", c(Gene1, Gene2), sep="::", remove= FALSE, na.rm = TRUE)%>%
  mutate(across(c(Gene1, Gene2), ~ gsub("\\..*","", .)))%>%
  separate(fusionGene, into= c('Gene1_vID','Gene2_vID'), sep = "::", remove = FALSE) %>%
  unite("original.fusion.gene.id", c(Gene1, Gene2), sep=":", remove= FALSE, na.rm = TRUE)
PAR_Y_problem1 <- FusionSeeker_Sim %>%
  filter(grepl("PAR_Y", Gene1_vID, ignore.case = TRUE)) %>% select(Gene1) 
PAR_Y_problem2 <-  FusionSeeker_Sim %>%
  filter(grepl("PAR_Y", Gene2_vID, ignore.case = TRUE)) %>% select(Gene2)

Gene_Name<-rbind(getBM(attributes = c("external_gene_name", "ensembl_gene_id"),
                             filters = "ensembl_gene_id",
                             values = (unique(c(PAR_Y_problem1$Gene1, PAR_Y_problem2$Gene2))),
                             mart = ensembljul2023),
                      getBM(attributes = c("external_gene_name", "ensembl_gene_id"),
                             filters = "ensembl_gene_id",
                             values = (unique(c(PAR_Y_problem1$Gene1, PAR_Y_problem2$Gene2))),
                             mart = ensemblv109),
                      getBM(attributes = c("external_gene_name", "ensembl_gene_id"),
                             filters = "ensembl_gene_id",
                             values = (unique(c(PAR_Y_problem1$Gene1, PAR_Y_problem2$Gene2))),
                             mart = ensemblv110)) %>% unique()
External_Gene_Name<-rbind(getBM(attributes = c("external_gene_name", "ensembl_gene_id", "ensembl_gene_id_version", "chromosome_name"),
                             filters = c("chromosome_name", "external_gene_name"),
                             values = list("Y", unique(Gene_Name$external_gene_name)),
                             mart = ensembljul2023), 
                          getBM(attributes = c("external_gene_name", "ensembl_gene_id", "ensembl_gene_id_version", "chromosome_name"),
                             filters = c("chromosome_name", "external_gene_name"),
                             values = list("Y", unique(Gene_Name$external_gene_name)),
                             mart = ensemblv109), 
                          getBM(attributes = c("external_gene_name", "ensembl_gene_id", "ensembl_gene_id_version",  "chromosome_name"),
                             filters = c("chromosome_name", "external_gene_name"),
                             values = list("Y", unique(Gene_Name$external_gene_name)),
                             mart = ensemblv110)) %>% unique()
External_Gene_Name$chromosome_name <- paste0("chr", External_Gene_Name$chromosome_name)
FusionSeeker_Sim_ensembl <- FusionSeeker_Sim %>% left_join(select((right_join(Gene_Name, External_Gene_Name, by= "external_gene_name")), c(ensembl_gene_id.x, ensembl_gene_id.y, ensembl_gene_id_version, chromosome_name)), by= c('Gene1'='ensembl_gene_id.x', 'Chrom1'='chromosome_name')) %>% 
  mutate(Gene1_alternative_ID= coalesce(ensembl_gene_id.y, Gene1), Gene1_alternative_vID= coalesce(ensembl_gene_id_version, Gene1_vID)) %>%
  select(-c(ensembl_gene_id.y, ensembl_gene_id_version))%>% 
  left_join(select((right_join(Gene_Name, External_Gene_Name, by= "external_gene_name")), c(ensembl_gene_id.x, ensembl_gene_id.y, ensembl_gene_id_version, chromosome_name)), by= c('Gene2'='ensembl_gene_id.x', 'Chrom2'='chromosome_name')) %>% 
  mutate(Gene2_alternative_ID= coalesce(ensembl_gene_id.y, Gene2), Gene2_alternative_vID= coalesce(ensembl_gene_id_version, Gene2_vID)) %>%
  select(-c(ensembl_gene_id.y, ensembl_gene_id_version)) %>%
  unite("fusion.gene.id" , c(Gene1_alternative_ID, Gene2_alternative_ID), sep = ":", remove= FALSE, na.rm = TRUE) %>% unique() 

#Assign label to false fusions or partially recalled fusions
Annot_FusionSeeker_Sim <- FusionSeeker_Sim_ensembl%>%
  left_join(Simulated_Fusion_Info_2, by = 'fusion.gene.id')%>%
  group_by(fusionGene.x) %>%
  mutate(fusionType = case_when(
    is.na(fusionType) ~ paste0("chromosomal_misalignment:",first(na.omit(fusionType))),  # Only update NA values
    .default = fusionType  # Keep existing values unchanged
  ))%>% ungroup()%>%
  replace_with_na(replace = list(fusionType = "chromosomal_misalignment:NA"))

Annot_FusionSeeker_Sim$fusionType <- mapply(function(g1, g2, Gen1, Gen2, current_type, chr1, chr2) {
   # Check if the current fusionType is empty
  if (current_type == "" || is.na(current_type)) {
    if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g1, ":", g2)) |
            str_detect(paste(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V1, subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V3, sep = ":"), paste0(g1, ":", g2)))) {
      return("truncated_tri_fusion")
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1)) |
            str_detect(paste(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V1, subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V3, sep = ":"), paste0(g2, ":", g1)))){
      return("reverse_order:truncated_tri_fusion") 
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1)))){
       matching_row <- subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusionType[which(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1)))]
      return(paste("reverse_order:", matching_row[1])) 
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$original.fusion.gene.id, paste0(g1, ":", g2)))){
       matching_row <- subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusionType[which(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$original.fusion.gene.id, paste0(g1, ":", g2)))]
      return(paste0("chromosomal_misalignment:", matching_row[1])) 
    } else if ((grepl("chrM", chr1, ignore.case = TRUE) & !grepl("chrM", chr2, ignore.case = TRUE)) | (!grepl("chrM", chr1, ignore.case = TRUE) & grepl("chrM", chr2, ignore.case = TRUE))){
      return("false_fusion:mitochondrial_genomic") 
    }else if (grepl("chrM", chr1, ignore.case = TRUE) & grepl("chrM", chr2, ignore.case = TRUE)){
      return("false_fusion:mitochondrial") 
    }else if (((g1 == g2)|(Gen1 == Gen2)) & (chr1 != chr2)){
      return("false_fusion:self_misalignment") 
    }else {
      return("false_fusion")
    }
  }
  return(current_type)
}, Annot_FusionSeeker_Sim$Gene1_alternative_ID.x, Annot_FusionSeeker_Sim$Gene2_alternative_ID.x,
 Annot_FusionSeeker_Sim$Gene1, Annot_FusionSeeker_Sim$Gene2, 
Annot_FusionSeeker_Sim$fusionType, Annot_FusionSeeker_Sim$Chrom1, Annot_FusionSeeker_Sim$Chrom2)

setdiff(unique(c(ogFusionSeeker$ID, ogFusionSeeker$Source)), unique(c(Annot_FusionSeeker_Sim$ID, Annot_FusionSeeker_Sim$Source)))

```
Repeat pre-processing steps for JAFFAL
```{r}
#JAFFAL
myfiles<-list.files(path = "/bioinformatics/ryley/Algorithm_Benchmark/Adapter_porechop_trimmed/JAFFAL/", pattern = "jaffa_results.csv", full.names = TRUE, recursive = TRUE)
JAFFAL_Sim <-  do.call(rbind, lapply(myfiles, function(filename) {
  read.table(filename, header = TRUE, sep = ",") %>%
    mutate(Source = basename(dirname(filename))) %>% mutate(control = ifelse(grepl("Spiked", sample), "positive", "negative"))
})) %>% separate(fusion.genes, into = c("Gene1", "Gene2"), sep = ":", remove = FALSE)
ogJAFFAL <-  do.call(rbind, lapply(myfiles, function(filename) {
  read.table(filename, header = TRUE, sep = ",") %>%
    mutate(Source = basename(dirname(filename))) %>% mutate(control = ifelse(grepl("Spiked", sample), "positive", "negative"))
}))

 getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name", "hgnc_symbol"),
       filters = "external_gene_name",
       values = "CBY3",
       mart = ensemblv109) 
 getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name", "hgnc_symbol"),
       filters = "external_gene_name",
       values = "CBY3",
       mart = ensembljul2023) 
# fusion.genes = fusion gene name : 
Gene_Name<-rbind(getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name"),
                             filters = "external_gene_name",
                             values = unique(c(JAFFAL_Sim$Gene1, JAFFAL_Sim$Gene2)),
                             mart = ensembljul2023) , getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name"),
                             filters = "external_gene_name",
                             values = unique(c(JAFFAL_Sim$Gene1, JAFFAL_Sim$Gene2)),
                             mart = ensemblv109)) %>% unique()
Gene_Name$chromosome_name <- paste0("chr", Gene_Name$chromosome_name)

JAFFAL_Sim_ensembl <- left_join(JAFFAL_Sim, Gene_Name, by= c('Gene1'='external_gene_name')) %>% 
  left_join(Gene_Name, by= c('Gene2'='external_gene_name'))  %>%
          mutate(ensembl_gene1 = ifelse(is.na(ensembl_gene_id.x) |nchar(ensembl_gene_id.x) == 0, Gene1, ensembl_gene_id.x), 
                 ensembl_gene2 = ifelse(is.na(ensembl_gene_id.y) |nchar(ensembl_gene_id.y) == 0, Gene2, ensembl_gene_id.y)) %>%
  unite("fusion.gene.id", c(ensembl_gene1, ensembl_gene2), sep=":", remove= FALSE, na.rm = TRUE) %>%
  # Create a filtering condition using case_when
  filter(case_when(
      is.na(chromosome_name.x) ~ TRUE, # Keep if chromosome_name.x is NA
      chromosome_name.x == chrom1 ~ TRUE, # Keep if chromosome names match
      chromosome_name.x == "chrMT" & chrom1 == "chrM" ~ TRUE, # Allow chrMT/chrM equivalence
      grepl("PATCH|HSCHR", chromosome_name.x) & !grepl("^chr([1-9]|1[0-9]|2[0-2]|X|Y|M)$", chrom1) ~ TRUE, 
      TRUE ~ FALSE # Remove rows that don't meet the above conditions
    )) %>%
  filter(case_when(
      is.na(chromosome_name.y) ~ TRUE,
      chromosome_name.y == chrom2 ~ TRUE,
      chromosome_name.y == "chrMT" & chrom2 == "chrM" ~ TRUE,
      grepl("PATCH|HSCHR", chromosome_name.y) & !grepl("^chr([1-9]|1[0-9]|2[0-2]|X|Y|M)$", chrom2) ~ TRUE,
      TRUE ~ FALSE # Remove rows that don't meet the above conditions
    )) %>%  
    select(-c("ensembl_gene_id.x", "chromosome_name.x", "ensembl_gene_id.y", "chromosome_name.y")) %>% unique()

setdiff(unique(JAFFAL_Sim_ensembl$fusion.genes), unique(ogJAFFAL$fusion.genes))

#Annotate as known or false positive
Annot_JAFFAL_Sim <- JAFFAL_Sim_ensembl %>% 
  left_join(Simulated_Fusion_Info_2, by = c('fusion.gene.id')) %>% unique()%>%
  group_by(fusion.genes) %>%
  mutate(fusionType = case_when(
    is.na(fusionType) ~ paste0("chromosomal_misalignment:",first(na.omit(fusionType))),  # Only update NA values
    .default = fusionType  # Keep existing values unchanged
  ))%>% ungroup()%>%
  replace_with_na(replace = list(fusionType = "chromosomal_misalignment:NA")) 
# Check if ensembl_gene1 and ensembl_gene2 are contained in fusion.gene.id
Annot_JAFFAL_Sim$fusionType <- mapply(function(g1, g2, Gen1, Gen2, current_type, chr1, chr2) {
   # Check if the current fusionType is empty
  if (current_type == "" || is.na(current_type)) {
    if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g1, ":", g2))|
            str_detect(paste(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V1, subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V3, sep = ":"), paste0(g1, ":", g2)))) {
      return("truncated_tri_fusion")
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1)))){
       matching_row <- subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")[which(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1)))]
      return(paste("reverse_order:", matching_row[1])) 
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1))|
            str_detect(paste(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V1, subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V3, sep = ":"), paste0(g2, ":", g1)))){
      return("reverse_order:truncated_tri_fusion") 
    } else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$original.fusion.gene.id, paste0(g1, ":", g2)))){
       matching_row <- subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusionType[which(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$original.fusion.gene.id, paste0(g1, ":", g2)))]
      return(paste("chromosomal_misalignment:", matching_row[1])) 
    }else if ((grepl("chrM:", chr1, ignore.case = TRUE) & !grepl("chrM:", chr2, ignore.case = TRUE)) | (!grepl("chrM:", chr1, ignore.case = TRUE) & grepl("chrM:", chr2, ignore.case = TRUE))){
      return("false_fusion:mitochondrial_genomic") 
    }else if (grepl("chrM:", chr1, ignore.case = TRUE) & grepl("chrM:", chr2, ignore.case = TRUE)){
      return("false_fusion:mitochondrial") 
    }else if (((g1 == g2)|(Gen1 == Gen2)) & (chr1 != chr2)){
      return("false_fusion:self_misalignment") 
    }else {
      return("false_fusion")
    }
  }
  return(current_type)
}, Annot_JAFFAL_Sim$ensembl_gene1, Annot_JAFFAL_Sim$ensembl_gene2, Annot_JAFFAL_Sim$Gene1, Annot_JAFFAL_Sim$Gene2, Annot_JAFFAL_Sim$fusionType, Annot_JAFFAL_Sim$chrom1, Annot_JAFFAL_Sim$chrom2)

#JAFFAL_3Gene_Sim 
myfiles<-list.files(path = "/bioinformatics/ryley/Algorithm_Benchmark/Adapter_porechop_trimmed/JAFFAL/", pattern = ".3gene_summary", full.names = TRUE, recursive = TRUE)
JAFFAL_3Gene_Sim <-  do.call(rbind, lapply(myfiles, function(filename) {
  read.table(filename, header = TRUE, sep = "\t") %>%
    mutate(Source = basename(dirname(filename))) %>% mutate(control = ifelse(grepl("Spiked", Source), "positive", "negative"))
})) %>% separate(Fusion, into = c("Gene1", "Gene2", "Gene3"), sep = ":", remove = FALSE)
Gene_Name<-rbind(getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name"),
                             filters = "external_gene_name",
                             values = unique(c(JAFFAL_3Gene_Sim$Gene1, JAFFAL_3Gene_Sim$Gene2, JAFFAL_3Gene_Sim$Gene3)),
                             mart = ensembljul2023) ,
                 getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name"),
                             filters = "external_gene_name",
                             values = unique(c(JAFFAL_3Gene_Sim$Gene1, JAFFAL_3Gene_Sim$Gene2, JAFFAL_3Gene_Sim$Gene3)),
                             mart = ensemblv109))

JAFFAL_3Gene_Sim <- left_join(JAFFAL_3Gene_Sim, Gene_Name, by= c('Gene1'='external_gene_name')) %>%
  left_join(Gene_Name, by= c('Gene2'='external_gene_name')) %>%
  left_join(Gene_Name, by= c('Gene3'='external_gene_name')) %>% unique() %>%
          mutate(ensembl_gene1 = ifelse(is.na(ensembl_gene_id.x) |nchar(ensembl_gene_id.x) == 0, Gene1, ensembl_gene_id.x), 
                 ensembl_gene2 = ifelse(is.na(ensembl_gene_id.y) |nchar(ensembl_gene_id.y) == 0, Gene2, ensembl_gene_id.y), 
                 ensembl_gene3 = ifelse(is.na(ensembl_gene_id) |nchar(ensembl_gene_id) == 0, Gene3, ensembl_gene_id)) %>%
  unite("fusion.gene.id", c(ensembl_gene1, ensembl_gene2, ensembl_gene3), sep=":", remove= FALSE, na.rm = TRUE)

Annot_JAFFAL_3Gene_Sim <- JAFFAL_3Gene_Sim[c(1:10,17:20)] %>% left_join(Simulated_Fusion_Info_2, by = c('fusion.gene.id'))
Annot_JAFFAL_3Gene_Sim$fusionType <- mapply(function(g1, g2, g3, current_type) {
   if (current_type == "" || is.na(current_type)) {
        
     #check if reverse order tri-fusion
        if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g3, ":", g2, ":", g1)))) {
            return("reverse_order:tri_fusion")}
     
     #check if its a butchered tri-fusion
        if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g3, ":", g1, ":", g2)) |
                str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g2, ":", g3, ":", g1))| 
                str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g1, ":", g3, ":", g2))| 
                str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1, ":", g3)))) {
            return("wrong_order:tri_fusion")
          }else {
          return("false_fusion")
        }
   }
  return(current_type)}, Annot_JAFFAL_3Gene_Sim$Gene1, Annot_JAFFAL_3Gene_Sim$Gene2, Annot_JAFFAL_3Gene_Sim$Gene3, Annot_JAFFAL_3Gene_Sim$fusionType)
```
Repeat pre-processing steps for Genion
```{r}
myfiles<- list.files(path = "/bioinformatics/ryley/Algorithm_Benchmark/Adapter_porechop_trimmed/Genion", pattern = "*_genion$", full.names = TRUE)
Genion_Sim<- do.call(rbind, lapply(myfiles, function(filename) {
  if (file.info(filename)$size > 0) {
    read.table(filename) %>%
      mutate(Source = basename(filename)) %>% mutate(control = ifelse(grepl("Spiked", Source), "positive", "negative"))
  } else {
    NULL
  }
}))
og_Genion <-Genion_Sim

# Annotate
Annot_Genion_Sim <- Genion_Sim %>% separate(V1, into = c("V1.1", "V1.2", "V1.3"), "::", remove=FALSE) %>% separate(V2, into = c("V2.1", "V2.2", "V2.3"), "::", remove=FALSE) %>% separate(V8, into = c("chr1", "chr2", "chr3"), ";", remove=FALSE)
Annot_Genion_Sim[c("V1", "V2")] <- lapply(Genion_Sim[c("V1", "V2")], function(x) gsub("::", ":", x))
Annot_Genion_Sim <- Annot_Genion_Sim %>% left_join(Simulated_Fusion_Info_2, by = c('V1'='fusion.gene.id')) 
Annot_Genion_Sim$fusionType <- mapply(function(g1, g2, g3, current_type, chr1, chr2, chr3, gene_name1, gene_name2) {
  # Check if the current fusionType is empty
  if (current_type == "" || is.na(current_type)) {
    
        #check if this is two part or tri part fusion
        if(g3 == ""|| is.na(g3)){
          
           #check if part of tri fusion
            if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g1, ":", g2)) |
                str_detect(paste(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V1, subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V3, sep = ":"), paste0(g1, ":", g2)))) {
              return("truncated_tri_fusion")} 
            else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1)) |
            str_detect(paste(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V1, subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$V3, sep = ":"), paste0(g2, ":", g1)))){
              return("reverse_order:truncated_tri_fusion")} 
          
          #check for reverse order two part genes
            else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1)))){
              matching_row <- subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusionType[which(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1)))]
              return(paste("reverse_order:", matching_row[1]))}
        } 
          #check for chromosomal misalignment
          else if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$original.fusion.gene.id, paste0(g1, ":", g2)))){
               matching_row <- subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$fusionType[which(str_detect(subset(Simulated_Fusion_Info_2, fusionType != "tri_fusion")$original.fusion.gene.id, paste0(g1, ":", g2)))]
              return(paste("chromosomal_misalignment:", matching_row[1])) 
            }
      
      #check if this fusion contains mitochondrial genes 
        if ((grepl("chrM:", chr1, ignore.case = TRUE) & (!grepl("chrM:", chr2, ignore.case = TRUE) | !grepl("chrM:", chr3, ignore.case = TRUE))) |
                 (grepl("chrM:", chr2, ignore.case = TRUE) & (!grepl("chrM:", chr1, ignore.case = TRUE)| !grepl("chrM:", chr3, ignore.case = TRUE)))|
                 (grepl("chrM:", chr3, ignore.case = TRUE) & (!grepl("chrM:", chr1, ignore.case = TRUE)| !grepl("chrM:", chr2, ignore.case = TRUE)))){
          return("false_fusion:mitochondrial_genomic")} 
      
        else if (grepl("chrM:", chr1, ignore.case = TRUE) & grepl("chrM:", chr2, ignore.case = TRUE) & grepl("chrM:", chr3, ignore.case = TRUE)){
          return("false_fusion:mitochondrial")}
    
        #check if its a butchered tri-fusion
        if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g3, ":", g1, ":", g2)) |
                str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g2, ":", g3, ":", g1))| 
                str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g1, ":", g3, ":", g2))| 
                str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g2, ":", g1, ":", g3)))) {
            return("wrong_order:tri_fusion")}
    
        #check if reverse order tri-fusion
        if (any(str_detect(subset(Simulated_Fusion_Info_2, fusionType == "tri_fusion")$fusion.gene.id, paste0(g3, ":", g2, ":", g1)))) {
            return("reverse_order:tri_fusion")}
    
    #if blank and none of the above, its just a false fusion
      else if (gene_name1 == gene_name2){
      return("false_fusion:self_misalignment") 
    } else {
          return("false_fusion")}
    }
  #if it's not blank, its already been labelled
  else {return(current_type)}
  }, Annot_Genion_Sim$V1.1, Annot_Genion_Sim$V1.2,  Annot_Genion_Sim$V1.3, Annot_Genion_Sim$fusionType, Annot_Genion_Sim$chr1, Annot_Genion_Sim$chr2 , Annot_Genion_Sim$chr3, Annot_Genion_Sim$V2.1, Annot_Genion_Sim$V2.2)
      
```
Append corresponding depth and sequencing identity labels 
```{r}
# Making Graphs
Annot_LongGF_Sim <- Annot_LongGF_Sim %>%
  mutate(depth = case_when(
    grepl("100GB", Source) ~ "100GB",
    grepl("10GB", Source) ~ "10GB",
    grepl("1GB", Source) ~ "1GB" # Keep original value if no match
  ), Sequence_Identity = case_when(
    grepl("95", Source) ~ "95%",
    grepl("90", Source) ~ "90%",
    grepl("85", Source) ~ "85%" # Keep original value if no match
  ))
Annot_JAFFAL_Sim <- Annot_JAFFAL_Sim%>%
  mutate(depth = case_when(
    grepl("100GB", sample) ~ "100GB",
    grepl("10GB", sample) ~ "10GB",
    grepl("1GB", sample) ~ "1GB"# Keep original value if no match
  ), Sequence_Identity = case_when(
    grepl("95", sample) ~ "95%",
    grepl("90", sample) ~ "90%",
    grepl("85", sample) ~ "85%" # Keep original value if no match
  ))
Annot_JAFFAL_3Gene_Sim <- Annot_JAFFAL_3Gene_Sim%>%
  mutate(depth = case_when(
    grepl("100GB", Source) ~ "100GB",
    grepl("10GB", Source) ~ "10GB",
    grepl("1GB", Source) ~ "1GB"# Keep original value if no match
  ), Sequence_Identity = case_when(
    grepl("95", Source) ~ "95%",
    grepl("90", Source) ~ "90%",
    grepl("85", Source) ~ "85%" # Keep original value if no match
  ))
Annot_Genion_Sim <- Annot_Genion_Sim %>%
  mutate(depth = case_when(
    grepl("100GB", Source) ~ "100GB",
    grepl("10GB", Source) ~ "10GB",
    grepl("1GB", Source) ~ "1GB"  # Keep original value if no match
  ), Sequence_Identity = case_when(
    grepl("95", Source) ~ "95%",
    grepl("90", Source) ~ "90%",
    grepl("85", Source) ~ "85%" # Keep original value if no match
  ))
Annot_FusionSeeker_Sim <- Annot_FusionSeeker_Sim %>%
  mutate(depth = case_when(
    grepl("100GB", Source) ~ "100GB",
    grepl("10GB", Source) ~ "10GB",
    grepl("1GB", Source) ~ "1GB"  # Keep original value if no match
  ), Sequence_Identity = case_when(
    grepl("95", Source) ~ "95%",
    grepl("90", Source) ~ "90%",
    grepl("85", Source) ~ "85%" # Keep original value if no match
  ))

Annot_LongGF_Sim$Algorithm <- "LongGF"
Annot_JAFFAL_Sim$Algorithm <- "JAFFAL"
Annot_JAFFAL_3Gene_Sim$Algorithm <- "JAFFAL"
Annot_Genion_Sim$Algorithm <- "Genion"
Annot_FusionSeeker_Sim$Algorithm <- "FusionSeeker"
```
Label fusions as False Calls, Partial Real and True Recall in Genion and JAFFAL datasets
```{r}
Genion_Fffscores <- Annot_Genion_Sim[c(1:27, 37, 44:46)]%>%
  unique()%>%  mutate(recall_category = case_when(
      grepl("false", fusionType) ~ "False_Call",
      grepl("truncated|reverse|chromosomal_misalignment|wrong", fusionType) ~ "Partial_Recall",
      TRUE ~ "True_Recall"))%>%
      select(Source, V4, V3.x, recall_category, control, depth, Sequence_Identity, fusionType, V7) %>% unique() #remove duplicate rows of true fusions
JAFFAL_scoring <- Annot_JAFFAL_Sim[c(1:25, 43, 50:52)] %>% unique() %>%  
  mutate(recall_category = case_when(
      grepl("false", fusionType) ~ "False_Call",
      grepl("truncated|reverse|chromosomal_misalignment|wrong", fusionType) ~ "Partial_Recall",
      TRUE ~ "True_Recall")) %>%
      select(sample, contig, fusion.genes, spanning.reads, classification, recall_category, control, depth, Sequence_Identity, fusionType) %>% unique()
```
Check if JAFFAL's classification correlates to recall category. 
Make Supplementary Figure 14
```{r}
Supp_Figure14A <- ggplot(filter(JAFFAL_scoring, control =="positive", depth == "100GB", Sequence_Identity == "95%"),
                        aes(y = spanning.reads, x = classification, colour =recall_category ))+
  geom_boxplot(outlier.shape = NA)+
  geom_sina()+  
  theme_minimal()+
  theme(legend.title = element_blank(), 
        legend.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))+
  labs(subtitle = "JAFFAL’s classification of False, Partial and True fusion calls ",
        title = "Sequencing Depth 100Gb , 95% Mean Sequence Identity",
       x = "", y ="supporting reads")

Supp_Figure14B <- ggplot(filter(JAFFAL_scoring, control =="positive", depth == "100GB", Sequence_Identity == "95%", fusionType=="read_through"),
       aes(y = spanning.reads, x = classification, colour =fusionType ))+
  geom_boxplot(outlier.shape = NA)+
  geom_sina()+  
  theme_minimal()+
  theme(legend.title = element_blank(), 
        legend.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))+
  labs(subtitle = "JAFFAL's classication of simulated read-throughs",
       x="",y ="supporting reads")
supp_Figure14C <-ggplot(filter(JAFFAL_scoring, control =="positive", depth == "100GB", Sequence_Identity == "95%", classification=="PotentialReadThrough"),
                        aes(y = spanning.reads, x = classification, colour = recall_category  ))+
  geom_boxplot(outlier.shape = NA)+
  geom_sina()+  
  theme_minimal()+
  theme(legend.title = element_blank(), 
        legend.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))+
  labs(subtitle = "JAFFAL's 'Potential Readthrough' recall classification", x="", y="")+
  ylim(c(0,20))
supp_Figure14 <- supp_Figure14A/(supp_Figure14B + supp_Figure14C)+
  plot_annotation(theme = theme(plot.title = element_text(size = 14, face = "bold")),
                  tag_levels = "A")
  supp_Figure14
ggsave("~/LongReadFusionCallerBenchmark/Figures/supp_Figure14.png", plot = supp_Figure14,  width = 297, height = 210 , unit ="mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/supp_Figure14.pdf", plot = supp_Figure14,  width = 297, height = 210 , unit ="mm")
```
check if Genion's FiN or ffigf score correlate to recall category (True, Partial or False call) or other variables. 
Make Supplementary Figure 15 & 16
```{r}
#supplementary figure 15
ggplot(filter(Genion_Fffscores, control =="positive"))+
  geom_point(aes(x = V4, y= V3.x, colour = recall_category))+
  facet_grid(depth ~ Sequence_Identity)+
  scale_x_log10(breaks = 10^(-5:3))+
  labs(x="FiN score", y="ffigf score")+ 
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))
#supplementary figure 16
ggplot(filter(Genion_Fffscores, control =="positive", fusionType=="read_through"))+
  geom_point(aes(x = V4, y= V3.x, colour = V7))+
  facet_grid(depth ~ Sequence_Identity)+
  scale_x_log10(breaks = 10^(-5:5))+
  labs(x="FiN score", y="ffigf score")+ 
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))
```

Consolidate fusion caller outputs into one dataframe to more easily compare
```{r}
dLGF<- Annot_LongGF_Sim[c(1:24, 34, 41:43)] %>% 
    unique()%>% #remove duplicate values
    select(c(5, 13, 14, 25:28))%>%#select only columns needed for downstream analysis
  rename(read_supp = V3.x)
                        
  
dJF <- Annot_JAFFAL_Sim[c(1:25, 43, 50:52)] %>% 
    unique()%>% #remove duplicate values
  select(c(13, 22, 23, 26:29)) %>% #select only columns needed for downstream analysis
  rename(read_supp = spanning.reads)
dJF3 <- Annot_JAFFAL_3Gene_Sim[c(1:21, 31, 38:40)] %>%
  unique()%>% #remove duplicate values
  select(c(5, 10, 11, 22:25))%>% #extract only columns needed for downstream analysis
  rename(read_supp = Reads)
dGen <- Annot_Genion_Sim[c(1:27, 37, 44:46)]%>%
  unique()%>% #remove duplicate values
  select(c(1, 11, 19, 28:31))%>% #extract only columns needed for downstream analysis
  rename(read_supp = V5,
         fusion.gene.id = V1)
dFS <- Annot_FusionSeeker_Sim[c(1:28, 38, 45:47)]%>%
  unique()%>% 
 select(c(8,15, 16, 29:32))%>%
  rename(read_supp = NumSupp)
combined_data <- rbind(dLGF, dJF, dJF3, dGen, dFS)
combined_data$read_supp <- as.numeric(combined_data$read_supp)

write_tsv(combined_data, paste0(Analysis_output_folder, "combined_data_FusionCalls.tsv"))

combined_data_annotation <- filter(combined_data, read_supp >= 2) %>% unique() %>% mutate(recall_category = case_when(
      grepl("false", fusionType) ~ "False_Call",
      grepl("truncated|reverse|chromosomal_misalignment|wrong", fusionType) ~ "Partial_Recall",
      TRUE ~ "True_Recall"))%>%
  group_by(control, depth, Sequence_Identity, Algorithm, recall_category) %>%
  summarise(RECALL_COUNT = n()) %>%
  summarise(False_Call_Number = sum(RECALL_COUNT[recall_category == "False_Call"], na.rm = TRUE),
            Partial_Call_Number = sum(RECALL_COUNT[recall_category == "Partial_Recall"], na.rm = TRUE),
            True_Call_Number = sum(RECALL_COUNT[recall_category == "True_Recall"], na.rm = TRUE))%>% 
  ungroup() %>% 
  complete(control, depth, Sequence_Identity, Algorithm, fill = list(False_Call_Number = 0, Partial_Call_Number = 0, True_Call_Number = 0))
combined_data_annotation$depth <- factor(combined_data_annotation$depth, levels = c("1GB", "10GB", "100GB"))
combined_data_annotation$Sequence_Identity <- factor(combined_data_annotation$Sequence_Identity, levels = c("85%", "90%", "95%"))

write_tsv(combined_data_annotation, file=paste0(Analysis_output_folder, "combined_data_annotation.tsv"))
```
Check number of false, partial and true calls. Generate Supplementary Figure 2
```{r}
false_call_plot <-ggplot(filter(combined_data_annotation), aes(x = depth, y = Sequence_Identity, fill = False_Call_Number)) +
  geom_tile() +
  geom_text(aes(label = False_Call_Number), color = "black", size = 3) +
  facet_grid( control ~ Algorithm) +  # Separate heatmaps for each algorithm
  scale_fill_gradient(low = "white", high = "red")+  # Adjust color scale
  labs(x = "",
       y = "",
       fill = "False Fusions") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.background = element_rect(color="black", fill="grey85"))

partial_call_plot <- ggplot(filter(combined_data_annotation, control =="positive"), aes(x = depth, y = Sequence_Identity, fill = Partial_Call_Number)) +
  geom_tile() +
  geom_text(aes(label = Partial_Call_Number), color = "black", size = 3) +
  facet_grid(control ~ Algorithm)+
  scale_fill_gradient(low = "white", high = "orange") +  # Adjust color scale
  labs(x = "",
       y = "Mean Sequence Identity",
       fill = "Partially Recalled Fusions") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.background = element_rect(color="black", fill="grey85"))

true_call_plot <-ggplot(filter(combined_data_annotation, control =="positive"), aes(x = depth, y = Sequence_Identity, fill = True_Call_Number)) +
  geom_tile() +
  geom_text(aes(label = True_Call_Number), color = "black", size = 3) +
  facet_grid(control ~ Algorithm) +  # Separate heatmaps for each algorithm
  scale_fill_gradient(low = "white", high = "springgreen4") +  # Adjust color scale
  labs(x = "Sequencing Depth",
       y = "",
       fill = "True Fusions") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.background = element_rect(color="black", fill="grey85"))

Supplementary_Figure2 <- (false_call_plot / partial_call_plot / true_call_plot)+ 
  plot_layout(heights = c(2, 1, 1)) + 
  plot_annotation(theme = theme(plot.title = element_text(size = 14, face = "bold")))

# Display the combined plot
Supplementary_Figure2
ggsave("~/LongReadFusionCallerBenchmark/Figures/Supplementary_Figure2.png", Supplementary_Figure2, width = 210, height = 210, units = "mm")
```

Calculate Recall vs Precision, and F1 score
```{r}
source("~/LongReadFusionCallerBenchmark/Precision_Recall_ReadSupp_stats.R") 
#Calculate Recall vs Precision
stat_summary <- filter(combined_data, read_supp >= 2, control == "positive") %>%  unique() %>%
  mutate(stat_category = case_when(
    grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType) ~ "FALSE_CALL",
    TRUE ~ "TRUE_CALL")) %>%
  
  # First summarisation: Keep fusion_type_count grouped by stat_category
  group_by(control, depth, Sequence_Identity, Algorithm, stat_category) %>%
  summarise(CALL_COUNT = n()) %>%
  summarise(False_Point = sum(CALL_COUNT[stat_category == "FALSE_CALL"], na.rm = TRUE),
            True_Point = sum(CALL_COUNT[stat_category == "TRUE_CALL"], na.rm = TRUE)) %>%
  left_join(Spiked_Counts_wo_SF[c(2:4)])%>%
  # Compute final statistics
  mutate(FDR = False_Point / (False_Point + True_Point),
         Precision = True_Point / (True_Point + False_Point),
         Recall = True_Point / Spiked_Number,  
         F1 = (2 * (Precision * Recall)) / (Precision + Recall),
         Sensitivity = True_Point / (True_Point + (Spiked_Number - True_Point)))

stat_summary$depth <- factor(stat_summary$depth, levels = c("100GB", "10GB", "1GB"))
stat_summary$Sequence_Identity <- factor(stat_summary$Sequence_Identity, levels = c("85%", "90%", "95%"))

write_tsv(stat_summary, file=paste0(Analysis_output_folder, "stat_summary.tsv"))

#Calculate the effect of read support on recall/precision
calculate_read_filt_statistics <- function(data, min_read_supp) {
  data %>%
    filter(read_supp >= min_read_supp, control == "positive")  %>%  
    mutate(stat_category = case_when(
      grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType) ~ "FALSE_CALL",
      TRUE ~ "TRUE_CALL")) %>% 
    group_by(control, depth, Sequence_Identity, Algorithm, stat_category) %>%
    summarise(CALL_COUNT = n()) %>%
    summarise(False_Point = sum(CALL_COUNT[stat_category == "FALSE_CALL"], na.rm = TRUE),
              True_Point = sum(CALL_COUNT[stat_category == "TRUE_CALL"], na.rm = TRUE)) %>%
    left_join(Spiked_Counts_wo_SF[c(2:4)])%>%
    
    mutate(FDR = False_Point / (False_Point + True_Point),
           Precision = True_Point / (True_Point + False_Point),
           Recall = True_Point / Spiked_Number,  
           F1 = (2 * (Precision * Recall)) / (Precision + Recall),
           Sensitivity = True_Point / (True_Point + (Spiked_Number - True_Point)),
           minimum_read_support = min_read_supp)
}

readsupp_filtering <- calculate_read_filt_statistics(combined_data, 2) 
                      
for(number in seq(3, max(sort(unique(combined_data$read_supp))))){
  result <- calculate_read_filt_statistics(combined_data, number)
  readsupp_filtering <- rbind(result, readsupp_filtering) 
}
readsupp_filtering$minimum_read_support <- as.numeric(readsupp_filtering$minimum_read_support)
```
Generate Supplementary Figures 8-10
```{r}
#Supplementary figure 8####
Recall_Precision <- ggplot(filter(stat_summary, control == 'positive'))+
  geom_point(aes(x = Precision, y= Recall, colour = Algorithm))+
  facet_grid(depth ~ Sequence_Identity)+
  labs(x = "Precision", y = "Recall")+
  ylim(c(0,1.00))+
  xlim(c(0,1.00)) +
  theme(plot.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))

ggsave("~/LongReadFusionCallerBenchmark/Figures/Recall-Precision.pdf", plot = Recall_Precision, width = 297, height = 210, units = "mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/Recall-Precision.png", plot = Recall_Precision, width = 297, height = 210, units = "mm")
#Supplementary figure 9####
ReadSupport_Precision <- ggplot(filter(readsupp_filtering, control == "positive"))+
  geom_line(aes(x = minimum_read_support, y = Precision, colour = Algorithm))+
  facet_grid(depth ~ Sequence_Identity)+
  scale_x_log10()+
  labs(x = "Read Support",
       y = "Precision") +
  theme(plot.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))
ggsave("~/LongReadFusionCallerBenchmark/Figures/ReadSupport-Precision.pdf", plot = ReadSupport_Precision, width = 297, height = 210, units = "mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/ReadSupport-Precision.png", plot = ReadSupport_Precision, width = 297, height = 210, units = "mm")
#Supplementary figure 10 ####
ReadSupport_F1Score <- ggplot(filter(readsupp_filtering, control == "positive"))+
  geom_line(aes(x = minimum_read_support, y = F1, colour = Algorithm))+
  facet_grid(depth ~ Sequence_Identity)+
  scale_x_log10()+
  labs(x = "Read Support",
       y = "F1 Score")+
  ylim(c(0,1.00))+
  theme(plot.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12)) 
ggsave("~/LongReadFusionCallerBenchmark/Figures/ReadSupport-F1Score.pdf", plot = ReadSupport_F1Score, width = 297, height = 210, units = "mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/ReadSupport-F1Score.png", plot = ReadSupport_F1Score, width = 297, height = 210, units = "mm")
```
Make Figure 3A 
```{r}
F1_subset<- ggplot(filter(readsupp_filtering, control == "positive", depth == "100GB", Sequence_Identity == "95%"))+
  geom_line(aes(x = minimum_read_support, y = F1, colour = Algorithm))+
  scale_x_log10()+
  labs(x = "Read Support",
       y = "F1")+
  ylim(c(0,1.00)) +
  theme(legend.position = "right",
        legend.text = element_text(size = 12),
    plot.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12))
Precision_subset<- ggplot(filter(readsupp_filtering, control == "positive", depth == "100GB", Sequence_Identity == "95%"))+
  geom_line(aes(x = minimum_read_support, y = Precision, colour = Algorithm))+
  scale_x_log10()+
  labs(x = "", y = "Precision") +
  ylim(c(0,1.00))+
  theme(legend.position = "none",
    plot.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)) +
  labs(title = "Sequencing Depth 100Gb , 95% Mean Sequence Identity")
Recall_subset<- ggplot(filter(readsupp_filtering, control == "positive", depth == "100GB", Sequence_Identity == "95%"))+
  geom_line(aes(x = minimum_read_support, y = Recall, colour = Algorithm))+
  scale_x_log10()+
  labs(x = "Read Support",
       y = "Recall")+
  ylim(c(0,1.00))+
  theme(legend.position = "none",
    plot.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12))

Figure3A <-((Precision_subset /  Recall_subset) | F1_subset) + 
              plot_layout(guides = 'keep')   

```
Check the effect of removing obvious false chimeras from the fusion calls to see if they greatly influence precision. There visually appears to be minimal/negligible influence.
```{r}
manual_filtering <- filter(combined_data, read_supp >=2) %>%  
  mutate(stat_category = case_when(
      grepl("false_fusion:mitochondrial|false_fusion:self_misalignment", fusionType) ~ "obvious_false",
      TRUE ~ "other")) %>% filter(stat_category == "other") %>% select(-'stat_category')%>%  
  mutate(stat_category = case_when(
      grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType) ~ "FALSE_CALL",
      TRUE ~ "TRUE_CALL")) %>%
  
  # First summarisation: Keep fusion_type_count grouped by stat_category
  group_by(control, depth, Sequence_Identity, Algorithm, stat_category) %>%
  summarise(CALL_COUNT = n()) %>%
  summarise(False_Point = sum(CALL_COUNT[stat_category == "FALSE_CALL"], na.rm = TRUE),
    True_Point = sum(CALL_COUNT[stat_category == "TRUE_CALL"], na.rm = TRUE)) %>%
  left_join(Spiked_Counts_wo_SF[c(2:4)])%>%
  # Compute final statistics
  mutate(FDR = False_Point / (False_Point + True_Point),
    Precision = True_Point / (True_Point + False_Point),
    Recall = True_Point / Spiked_Number,  
    F1 = (2 * (Precision * Recall)) / (Precision + Recall),
    Sensitivity = True_Point / (True_Point + (Spiked_Number - True_Point)))

manual_filtering$depth <- factor(manual_filtering$depth, levels = c("100GB", "10GB", "1GB"))
manual_filtering$Sequence_Identity <- factor(manual_filtering$Sequence_Identity, levels = c("85%", "90%", "95%"))

ggplot(filter(manual_filtering, control == 'positive'))+
     geom_point(aes(x = Precision, y= Recall, colour = Algorithm))+
     facet_grid(depth ~ Sequence_Identity)+
  labs(title = "Recall and Precision", subtitle = "heuristic filtering", x = "Precision", y = "Recall")+
  ylim(c(0,1.00))+
  xlim(c(0,1.00))

calculate_read_filt_statistics <- function(data, min_read_supp) {
  data %>%
    filter(read_supp > min_read_supp)  %>%  
  mutate(stat_category = case_when(
      grepl("false_fusion:mitochondrial|false_fusion:self_misalignment", fusionType) ~ "obvious_false",
      TRUE ~ "other")) %>% filter(stat_category == "other") %>% select(-'stat_category') %>%  
  mutate(stat_category = case_when(
      grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType) ~ "FALSE_CALL",
      TRUE ~ "TRUE_CALL")) %>%
  
  # First summarisation: Keep fusion_type_count grouped by stat_category
  group_by(control, depth, Sequence_Identity, Algorithm, stat_category) %>%
  summarise(CALL_COUNT = n()) %>%
  summarise(False_Point = sum(CALL_COUNT[stat_category == "FALSE_CALL"], na.rm = TRUE),
    True_Point = sum(CALL_COUNT[stat_category == "TRUE_CALL"], na.rm = TRUE)) %>%
   left_join(Spiked_Counts_wo_SF[c(2:4)])%>%
  # Compute final statistics
  mutate(FDR = False_Point / (False_Point + True_Point),
    Precision = True_Point / (True_Point + False_Point),
    Recall = True_Point / Spiked_Number,  
    F1 = (2 * (Precision * Recall)) / (Precision + Recall),
    Sensitivity = True_Point / (True_Point + (Spiked_Number - True_Point)))
}

manual_readsupp_filtering <- calculate_read_filt_statistics(combined_data, 2) 
  for(number in unique(sort(combined_data$read_supp))){
  result <- calculate_read_filt_statistics(combined_data, number)
  manual_readsupp_filtering <- rbind(result, manual_readsupp_filtering) 
  }
manual_readsupp_filtering$minimum_read_support <- as.numeric(manual_readsupp_filtering$minimum_read_support)

ggplot(filter(manual_readsupp_filtering, control == "positive"))+
  geom_line(aes(x = minimum_read_support, y = F1, colour = Algorithm))+
  facet_grid(depth ~ Sequence_Identity)+
  scale_x_log10()+
  labs(title = "F1 score in relation to minimum read support", subtitle = "heuristic filtering", 
       x = "Read Support",
       y = "F1 Score")+
    ylim(c(0,1.00)) 

ggplot(filter(manual_readsupp_filtering, control == "positive"))+
  geom_line(aes(x = minimum_read_support, y = Precision, colour = Algorithm))+
  facet_grid(depth ~ Sequence_Identity)+
  scale_x_log10()+
  labs(title = "Precision in relation to minimum read support", subtitle = "heuristic filtering",
       x = "Read Support",
       y = "Precision") 

```
Create a summary of the number of false, partial and true calls in every simulation
```{r}
fusion_profile_stat_summary <- combined_data %>%  filter(read_supp >=2) %>% unique()%>% group_by(control, depth, Sequence_Identity, Algorithm, fusionType) %>%
  summarise(fusion_type_count = n(), .groups = 'drop')%>% 
  ungroup() %>% 
  complete(control, depth, Sequence_Identity, Algorithm, fusionType, fill = list(fusion_type_count = 0)) %>%
  group_by(control, depth, Sequence_Identity, Algorithm) %>%
  mutate(recall_category = case_when(
      grepl("false", fusionType) ~ "False_Call",
      grepl("truncated|reverse|chromosomal_misalignment|wrong", fusionType) ~ "Partial_Recall",
      TRUE ~ "True_Recall"))

fusion_profile_stat_summary$depth <- factor(fusion_profile_stat_summary$depth, levels = c("1GB", "10GB", "100GB"))
fusion_profile_stat_summary$Sequence_Identity <- factor(fusion_profile_stat_summary$Sequence_Identity, levels = c("85%", "90%", "95%"))

FCC <- ggplot(filter(fusion_profile_stat_summary, recall_category == "False_Call"))+ #, control == "positive"
       aes(x = depth, y = Sequence_Identity, fill = fusion_type_count) +
  geom_tile() +
  geom_text(aes(label = fusion_type_count), color = "black", size = 3, alpha =1) +
  facet_nested(control + fusionType ~ Algorithm, labeller = as_labeller(c(c("wrong_order:tri_fusion" = "Wrong Order Tri-fusion", 
                             "false_fusion:self_misalignment" = "Self-Misalignment",
                             "false_fusion:mitochondrial_genomic" = "Mitochondrial:Genomic",
                             "false_fusion:mitochondrial" = "Mitochondrial",
                             "false_fusion" = "False Chimera"), 
                             c("Genion"="Genion", "JAFFAL" = "JAFFAL", "LongGF" = "LongGF", "FusionSeeker" = "FusionSeeker"),
                             c("positive"="Postive", "negative" = "Negative"))) ) +  # Separate heatmaps for each algorithm
  scale_fill_gradient(low = "white", high = "red")+
  scale_alpha_continuous(range = c(0.2, 1)) +  # Adjust transparency based on fusion_type_count
  labs(x = "Depth",
       y = "Mean Sequence Identity",
       fill =  "#Fusions") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.background = element_rect(color="black", fill="lightgrey")) 
ggsave("~/LongReadFusionCallerBenchmark/Figures/False_fusion_category_counts.pdf", plot = FCC, width = 210, height = 297, units = "mm")
```
Make Supplementary Figure 1
```{r}
TRC<- ggplot(filter(fusion_profile_stat_summary, control == "positive", recall_category == "True_Recall"))+ 
       aes(x = depth, y = Sequence_Identity, fill = fusion_type_count)+
  geom_tile()+
  geom_text(aes(label = fusion_type_count), color = "black", size = 3, alpha =1) +
  facet_nested(fusionType ~ Algorithm, labeller = as_labeller(c(c("hybrid"="Inter-\n chromosomal", 
                             "intra_chromosome"="Intra- \n chromosomal",
                             "read_through"="Read-through",
                             "tri_fusion"="Tri-fusion"), 
                             c("Genion"="Genion", "JAFFAL" = "JAFFAL", "LongGF" = "LongGF", "FusionSeeker" = "FusionSeeker"),
                             c("positive"="Postive", "negative" = "Negative"))) ) +  # Separate heatmaps for each algorithm
  scale_fill_gradient(low = "white", high = "springgreen4") +  # Adjust color scale
  scale_alpha_continuous(range = c(0.2, 1)) +  # Adjust transparency based on fusion_type_count
  labs(x = "Depth",
       y = "Mean Sequence Identity",
       fill =  "Fusions") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.background = element_rect(color="black", fill="lightgrey")) 
ggsave("~/LongReadFusionCallerBenchmark/Figures/True_Recall_category.pdf", plot = TRC) #, width = 210, height = 297, units = "mm"
ggsave("~/LongReadFusionCallerBenchmark/Figures/True_Recall_category.png", plot = TRC)
```
Make Supplementary Figure 4
```{r}
PRC <- ggplot(filter(fusion_profile_stat_summary, control == "positive", recall_category == "Partial_Recall"))+ 
       aes(x = depth, y = Sequence_Identity, fill = fusion_type_count) +
  geom_tile() +
  geom_text(aes(label = fusion_type_count), color = "black", size = 3, alpha =1) +
  facet_nested(fusionType ~ Algorithm, labeller = as_labeller(c(c("reverse_order: hybrid"="(R) Inter-\nchromosomal", 
                             "reverse_order: intra_chromosome"="(R)  Intra-\nchromosomal",
                             "reverse_order: read_through"="(R) Read-through",
                             "wrong_order:tri_fusion"="(W) Tri-fusion",
                             "reverse_order:tri_fusion"="(R) Tri-fusion",
                             "reverse_order:truncated_tri_fusion"="(R) Truncated \n Tri-fusion",
                             "truncated_tri_fusion"="Truncated \n Tri-fusion" ,
                             "chromosomal_misalignment:intra_chromosome"="(CM)  Intra-\nchromosomal"), 
                             c("Genion"="Genion", "JAFFAL" = "JAFFAL", "LongGF" = "LongGF", "FusionSeeker" = "FusionSeeker"),
                             c("positive"="Postive", "negative" = "Negative"))) ) +  # Separate heatmaps for each algorithm
  scale_fill_gradient(low = "white", high = "orange") +
  scale_alpha_continuous(range = c(0.2, 1)) +  # Adjust transparency based on fusion_type_count
  labs(x = "Depth",
       y = "Mean Sequence Identity",
       fill =  "#Fusions") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.background = element_rect(color="black", fill="lightgrey")) 
ggsave("~/LongReadFusionCallerBenchmark/Figures/Partial_Recall_category.pdf", plot = PRC, width = 210, height = 300, units = "mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/Partial_Recall_category.png", plot = PRC, width = 210, height = 300, units = "mm")
```


Make Figure 1 part D-G
```{r}
Figure1D <- ggplot(filter(combined_data_annotation, control == "positive", depth == "100GB", Sequence_Identity == "95%")) +
  geom_bar(aes(x = Algorithm, y = True_Call_Number, fill = Algorithm), stat = "identity") +
  theme_minimal() +
  labs(title = "True Fusions",
       subtitle = "95% Sequence Identity 100Gb Depth",
       y = "Simulated Fusions",
       x = "")+
  theme(legend.position = "none")+
  theme(
    plot.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12))+
  ylim(c(0,400))

Figure1E <-ggplot(filter(fusion_profile_stat_summary, control == "positive", recall_category == "True_Recall", depth == "100GB", Sequence_Identity == "95%"))+ 
  
  geom_bar(aes(x = Algorithm, y = fusion_type_count, fill = Algorithm), stat = "identity")+
  facet_wrap(~fusionType, labeller = as_labeller(c(c("hybrid"="Inter-chromosomal", 
                                                     "intra_chromosome"="Intra-chromosomal",
                                                     "read_through"="Read-through",
                                                     "tri_fusion"="Tri-fusion"))))+ 
  theme_minimal() +
  labs(title = "",
       subtitle = "",
       y = "",
       x = "")+
  ylim(c(0,100))+ 
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
  plot.title = element_text(size = 12),
  axis.text = element_text(size = 12),
  axis.title = element_text(size = 12),
  legend.position = "bottom")

Figure1DE <- (Figure1D + Figure1E)+
  plot_annotation(theme = theme(plot.title = element_text(size = 14, face = "bold")))
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure1DE.pdf", plot = Figure1DE)

Figure1F <- ggplot(
  filter(combined_data_annotation, control == "positive", depth == "100GB", Sequence_Identity == "95%")
) +
  geom_bar(aes(x = Algorithm, y = False_Call_Number, fill = Algorithm), stat = "identity") +
  theme_minimal() +
  labs(title = "False Chimeras",
       subtitle = "95% Sequence Identity 100Gb Depth",
       y = "Chimeras",
       x = "")+
  theme(legend.position = "none")+
  theme(
    plot.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12))

Figure1G <-ggplot(filter(fusion_profile_stat_summary, control == "positive", recall_category == "False_Call", depth == "100GB", Sequence_Identity == "95%"))+ 
  
  geom_bar(aes(x = Algorithm, y = fusion_type_count, fill = Algorithm), stat = "identity")+
  facet_wrap(~fusionType, labeller = as_labeller(c(c("wrong_order:tri_fusion" = "Wrong Order Tri-fusion", 
                                                     "false_fusion:self_misalignment" = "Self\nMisalignment",
                                                     "false_fusion:mitochondrial_genomic" = "Mitochondrial\nGenomic",
                                                     "false_fusion:mitochondrial" = "Mitochondrial",
                                                     "false_fusion" = "False Chimera"))))+ 
  theme_minimal() +
  theme(legend.position = "none", axis.text.x=element_blank())+
  theme(
    plot.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12))+
  labs(title = "",
       subtitle = "",
       y = "",
       x = "")

Figure1FG <- (Figure1F + Figure1G)+
  plot_annotation(theme = theme(plot.title = element_text(size = 14, face = "bold")))
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure1FG.pdf", plot = Figure1FG)
```

Make Supplementary Figure 5
```{r}
Figure5A_sup <- ggplot(
  filter(combined_data_annotation, control == "positive", depth == "100GB", Sequence_Identity == "95%")
) +
  geom_bar(aes(x = Algorithm, y = Partial_Call_Number, fill = Algorithm), stat = "identity") +
  theme_minimal() +
  labs(title = "Partially Recalled Fusions",
       subtitle = "95% Sequence Identity 100Gb Depth",
       y = "Simulated Fusions",
       x = "")+
  theme(legend.position = "none")+
  ylim(c(0,400))

Figure5B_sup <-ggplot(filter(fusion_profile_stat_summary, control == "positive", recall_category == "Partial_Recall", depth == "100GB", Sequence_Identity == "95%"))+ 
       
  geom_bar(aes(x = Algorithm, y = fusion_type_count, fill = Algorithm), stat = "identity")+
  facet_wrap(~fusionType, labeller = as_labeller(c(c("reverse_order: hybrid"="(R) Inter- \n Chromosomal", 
                             "reverse_order: intra_chromosome"="(R) Intra-\nChromosomal",
                             "reverse_order: read_through"="(R) Read-through",
                             "wrong_order:tri_fusion"="(W) Tri-Fusion",
                             "reverse_order:tri_fusion"="(R) Tri-Fusion",
                             "reverse_order:truncated_tri_fusion"="(R) Truncated \nTri-Fusion",
                             "truncated_tri_fusion"="Truncated \nTri-Fusion" ,
                             "chromosomal_misalignment:intra_chromosome"="(CM) Intra-\nChromosomal"))))+ 
  theme_minimal() +
  theme(legend.position = "none", axis.text.x=element_blank())+
  labs(title = "",
       subtitle = "",
       y = "",
       x = "")+
  ylim(c(0,100))+ 
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))

Figure5_sup <- (Figure5A_sup + Figure5B_sup)+
  plot_annotation(theme = theme(plot.title = element_text(size = 14, face = "bold")),tag_levels = "A")
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure5_sup.pdf", plot = Figure5_sup, width = 297 , unit ="mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure5_sup.png", plot = Figure5_sup, width = 297 , unit ="mm")
```

Create Supplementary Figure 6
```{r}
combined_data_annotation$depth <- factor(combined_data_annotation$depth, levels = c("1GB", "10GB", "100GB"))
fusion_profile_stat_summary$depth <- factor(fusion_profile_stat_summary$depth, levels = c("1GB", "10GB", "100GB")) 

Figure6A_1_sup <-ggplot(filter(
  combined_data_annotation, control == "positive", Sequence_Identity == "95%"),
  aes(x = depth, y = True_Call_Number, colour = Algorithm, group = interaction(Algorithm, Sequence_Identity))
) +
  geom_point() +
  geom_line()+
  theme_minimal() +
  labs(title = "Sequencing Depth ~ True Fusions",
       subtitle = "95% Sequence Identity",
       y = "Simulated Fusions",
       x = "")+
  ylim(c(0,400))

Figure6A_2_sup <-ggplot(filter(fusion_profile_stat_summary, control == "positive", recall_category == "True_Recall", Sequence_Identity == "95%"),
                    aes(x = depth, y = fusion_type_count, colour = Algorithm, group = interaction(Algorithm, Sequence_Identity)))+ 
  geom_point() +
  geom_line()+
  facet_wrap(~fusionType, labeller = as_labeller(c(c("hybrid"="Inter-chromosomal", 
                             "intra_chromosome"="Intra-chromosomal",
                             "read_through"="Read-through",
                             "tri_fusion"="Tri-fusion"))))+ 
  theme_minimal() +
  theme(legend.position = "none")+
  labs(title = "",
       subtitle = "",
       y = "",
       x = "")+
  ylim(c(0,100))

Figure6B_1_sup <-ggplot(filter(
  combined_data_annotation, control == "positive", Sequence_Identity == "95%"),
  aes(x = depth, y = False_Call_Number, colour = Algorithm, group = interaction(Algorithm, Sequence_Identity))
) +
  geom_point() +
  geom_line()+
  theme_minimal() +
  labs(title = "Sequencing Depth ~ False Fusions",
       subtitle = "95% Sequence Identity",
       y = "Chimeras",
       x = "")

Figure6B_2_sup <-ggplot(filter(fusion_profile_stat_summary, control == "positive", recall_category == "False_Call", Sequence_Identity == "95%"),
                    aes(x = depth, y = fusion_type_count, colour = Algorithm, group = interaction(Algorithm, Sequence_Identity)))+ 
  geom_point() +
  geom_line()+
  facet_wrap(~fusionType, labeller = as_labeller(c(c("wrong_order:tri_fusion" = "Wrong Order Tri-fusion", 
                             "false_fusion:self_misalignment" = "Self-Misalignment",
                             "false_fusion:mitochondrial_genomic" = "Mitochondrial:Genomic",
                             "false_fusion:mitochondrial" = "Mitochondrial",
                             "false_fusion" = "False Chimera"))))+ 
  theme_minimal() +
  theme(legend.position = "none")+
  labs(title = "",
       subtitle = "",
       y = "",
       x = "")

Figure6_sup <- ((Figure6A_1_sup + Figure6A_2_sup)/
            (Figure6B_1_sup + Figure6B_2_sup))+
  plot_annotation(tag_levels = list(c("A.1", "A.2", "B.1", "B.2")),
                  theme = theme(plot.title = element_text(size = 14, face = "bold")))
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure6_sup.pdf", plot = Figure6_sup, width = 270 , height = 270 , unit ="mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure6_sup.png", plot = Figure6_sup, width = 270 , height = 210 , unit ="mm")

```

Make Supplementary Figure 7
```{r}
Figure7A_1_sup <-ggplot(filter(
  combined_data_annotation, control == "positive", depth == "100GB"),
  aes(x = Sequence_Identity, y = True_Call_Number, colour = Algorithm, group = interaction(Algorithm, depth))
) +
  geom_point() +
  geom_line()+
  theme_minimal() +
  labs(title = "Sequence Identity ~ True Fusions",
       subtitle = "100Gb Sequencing Depth",
       y = "Simulated Fusions",
       x = "")+
  ylim(c(0,400))

Figure7A_2_sup <-ggplot(filter(fusion_profile_stat_summary, control == "positive", recall_category == "True_Recall", depth == "100GB"),
                    aes(x = Sequence_Identity, y = fusion_type_count, colour = Algorithm, group = interaction(Algorithm, depth)))+ 
  geom_point() +
  geom_line()+
  facet_wrap(~fusionType, labeller = as_labeller(c(c("hybrid"="Inter-chromosomal", 
                             "intra_chromosome"="Intra-chromosomal",
                             "read_through"="Read-through",
                             "tri_fusion"="Tri-fusion"))))+ 
  theme_minimal() +
  theme(legend.position = "none")+
  labs(title = "",
       subtitle = "",
       y = "",
       x = "")+
  ylim(c(0,100))+ 
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

Figure7B_1_sup <-ggplot(filter(
  combined_data_annotation, control == "positive", depth == "100GB"),
  aes(x = Sequence_Identity, y = False_Call_Number, colour = Algorithm, group = interaction(Algorithm, depth))
) +
  geom_point() +
  geom_line()+
  theme_minimal() +
  labs(title = "Sequence Identity ~ False Chimeras",
       subtitle = "100Gb Sequencing Depth",
       y = "Fusions",
       x = "")

Figure7B_2_sup <-ggplot(filter(fusion_profile_stat_summary, control == "positive", recall_category == "False_Call", depth == "100GB"),
                    aes(x = Sequence_Identity, y = fusion_type_count, colour = Algorithm, group = interaction(Algorithm, depth)))+ 
  geom_point() +
  geom_line()+
  facet_wrap(~fusionType, labeller = as_labeller(c(c("wrong_order:tri_fusion" = "Wrong Order Tri-fusion", 
                             "false_fusion:self_misalignment" = "Self-Misalignment",
                             "false_fusion:mitochondrial_genomic" = "Mitochondrial:Genomic",
                             "false_fusion:mitochondrial" = "Mitochondrial",
                             "false_fusion" = "False Chimera"))))+ 
  theme_minimal() +
  theme(legend.position = "none")+
  labs(title = "",
       subtitle = "",
       y = "",
       x = "")

Figure7_sup <- (Figure7A_1_sup + Figure7A_2_sup)/
               (Figure7B_1_sup + Figure7B_2_sup)+
               plot_annotation(tag_levels = list(c("A.1", "A.2", "B.1", "B.2")),
                  theme = theme(plot.title = element_text(size = 14, face = "bold")))
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure7_sup.pdf", plot = Figure7_sup, width = 270 , height = 270 , unit ="mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure7_sup.png", plot = Figure7_sup, width = 270 , height = 210 , unit ="mm")

```

Make Figure 2
```{r}
stat_summary_2<- stat_summary
stat_summary_2$depth <- factor(stat_summary_2$depth, levels = c("1GB", "10GB", "100GB"))

Figure2A_1 <- ggplot(filter(
  stat_summary_2, control == "positive", Sequence_Identity == "95%"),
  aes(x = depth, y = Precision, colour = Algorithm, group = interaction(Algorithm, Sequence_Identity))) +
  geom_point() +
  geom_line()+
  theme_minimal() +
  labs(title = "Sequencing Depth",
       subtitle = "95% Sequence Identity",
       y = "Precision",
       x = "")+
  theme(
    plot.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12))
Figure2A_2 <-ggplot(filter(
  stat_summary_2, control == "positive", depth == "100GB"),
  aes(x = Sequence_Identity, y = Precision, colour = Algorithm, group = interaction(Algorithm, depth))) +
  geom_point() +
  geom_line()+
  theme_minimal() +
  labs(title = "Sequence Identity",
       subtitle = "100Gb Sequencing Depth",
       y = "",
       x = "")+
  ylim(c(0,1.00))+
  theme(legend.position = "none",
    plot.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12))
Figure2B_1 <-ggplot(filter(
  stat_summary_2, control == "positive", Sequence_Identity == "95%"),
  aes(x = depth, y = Recall, colour = Algorithm, group = interaction(Algorithm, Sequence_Identity))) +
  geom_point() +
  geom_line()+
  theme_minimal() +
  labs(title = "",
       subtitle = "",
       y = "Recall",
       x = "Depth")+
  ylim(c(0,1.00))+
  theme(
    plot.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12))
Figure2B_2 <-ggplot(filter(
  stat_summary_2, control == "positive", depth == "100GB"),
  aes(x = Sequence_Identity, y = Recall, colour = Algorithm, group = interaction(Algorithm, depth))) +
  geom_point() +
  geom_line()+
  theme_minimal() +
  labs(title = "",
       subtitle = "",
       y = "",
       x = "Sequence Identity")+
  ylim(c(0,1.00))+
  theme(legend.position = "none",
    plot.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12))

Figure2 <- (Figure2A_1 + Figure2A_2)/
  (Figure2B_1 + Figure2B_2)+
  plot_annotation(theme = theme(plot.title = element_text(size = 14, face = "bold")))
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure2_paper.png", plot = Figure2, width = 297, height = 210, units = "mm")

```
fusionseeker and some of fusion callers will produce duplicate calls the same fusion multiple times with different read support because of breakpoint differences. We need to consolidate these for ease in comparing how read support for the same fusion will differ by fusion caller tool
```{r}
fusions_in_all_algorithms <- combined_data %>%
  filter(read_supp >=2)%>% unique() %>% #filter for read support greater than 2
  group_by(fusion.gene.id, Sequence_Identity, depth, control, fusionType) %>%
  summarise(overlap = n_distinct(Algorithm)) %>%
  filter(overlap == 4) %>% #filter for fusions in all 4 algorithms
  ungroup() %>%
  left_join(combined_data)%>% unique()%>%
  group_by(fusion.gene.id, Sequence_Identity, depth, control, Algorithm, overlap, fusionType) %>%
  summarize(Total_Read_Supp = sum(read_supp))%>%
  mutate(recall_category = case_when(
      grepl("false", fusionType) ~ "False_Call",
      grepl("truncated|reverse|chromosomal_misalignment|wrong", fusionType) ~ "Partial_Recall",
      TRUE ~ "True_Recall"))

fusions_in_all_algorithms$depth <- factor(fusions_in_all_algorithms$depth, levels = c("1GB", "10GB", "100GB"))
fusions_in_all_algorithms$Sequence_Identity <- factor(fusions_in_all_algorithms$Sequence_Identity, levels = c("95%", "90%", "85%"))
```
Make Figure 3B: comparing read support for the same fusions as called by different algorithms
```{r}
Figure3B <-ggplot(filter(fusions_in_all_algorithms, depth == "100GB", Sequence_Identity == "95%", control == "positive", overlap == 4), aes(x = fusion.gene.id, y = Algorithm, fill = Total_Read_Supp)) +
  geom_tile() +
  geom_text(aes(label = Total_Read_Supp), color = "black", size = 5, alpha =1) +  # Add numbers to the tiles, adjust color for visibility
  labs(x="", y="",
       fill = "Read Support") +
  theme_minimal() +
  facet_nested(~ recall_category , scales = "free_x", space = "free",
               labeller = as_labeller(c("Partial_Recall" = "Partial Recall", "True_Recall" = "True Recall")))+
  scale_fill_gradient2(low = "lemonchiffon", mid="orange" ,high = "orchid", na.value = "grey", midpoint = 25)+
  scale_alpha_continuous(range = c(0.1, 1)) + 
  theme(strip.text.x = element_text(size = 12),
        legend.position = "right",
        legend.text = element_text(size = 12),
        legend.spacing.x = unit(0.1, "cm"),   # reduce horizontal gap
        legend.spacing.y = unit(0.1, "cm"),   # reduce vertical gap inside legend
        legend.margin    = margin(0, 0, 0, 0),# trim around the legend box
        axis.text.x = element_text(angle = 45, hjust = 1), 
        strip.background = element_rect(color="black", fill="grey80"),
        plot.title = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.text.x.bottom = element_text(size = 10),
        axis.title = element_text(size = 12))+ plot_layout(guides = 'keep')

ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure3B_paper.pdf", plot = Figure3B, width = 297, height = 210, units = "mm")  
```
Make Compound Figure 3
```{r}
Figure3<- (Figure3A / Figure3B) +
          plot_annotation(tag_levels = list(c("A", "", "", "B")))
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure3_paper.png", plot = Figure3, width = 210, height = 297, units = "mm")  

```

identify fusions in more than one algorithm
```{r}
#consolidate reads for duplicate fusion calls due to different breakpoints
Total_Read_Supp_summary <- combined_data%>% 
   filter(read_supp >=2)%>% unique() %>% #remove duplicates as a result of annotations
  group_by(fusion.gene.id, Sequence_Identity, depth, control, Algorithm, fusionType)  %>%
  summarise(Total_Read_Supp = sum(read_supp), .groups = "drop")

# Second summarise
Algorithm_summary <- Total_Read_Supp_summary %>%
  group_by(fusion.gene.id, Sequence_Identity, depth, control, fusionType) %>%
  summarise(overlap = n_distinct(Algorithm), .groups = "drop")

# Join to retain Total_Read_Supp
fusions_in_n_algorithms <- left_join(Total_Read_Supp_summary, Algorithm_summary, 
                          by = c("fusion.gene.id", "Sequence_Identity", "depth", "control", "fusionType")) %>%
                          mutate(recall_category = case_when(
                              grepl("false", fusionType) ~ "False_Call",
                              grepl("truncated|reverse|chromosomal_misalignment|wrong", fusionType) ~ "Partial_Recall",
                              TRUE ~ "True_Recall"))
fusions_in_n_algorithms$depth <- factor(fusions_in_n_algorithms$depth, levels = c("1GB", "10GB", "100GB"))
fusions_in_n_algorithms$Sequence_Identity <- factor(fusions_in_n_algorithms$Sequence_Identity, levels = c("95%", "90%", "85%"))
fusions_in_n_algorithms <- fusions_in_n_algorithms %>%
  mutate(fusion.gene.id = fct_reorder(fusion.gene.id, overlap, .desc = TRUE))
```

Supplementary Figure 11
```{r}
supp_Figure11_A<- ggplot(filter(fusions_in_n_algorithms, depth == "100GB", control == "positive", overlap == 4), aes(x = fusion.gene.id, y = Algorithm, fill = Total_Read_Supp)) +
  geom_tile() +
  geom_text(aes(label = Total_Read_Supp), color = "black", size = 4, alpha =1) +  # Add numbers to the tiles, adjust color for visibility
  labs(x="", y="",
       fill = "Read Support") +
  theme_minimal() +
  facet_nested(Sequence_Identity ~ recall_category , 
               scales = "free_x", space = "free",
               labeller = labeller(
    recall_category = as_labeller(c("Partial_Recall" = "Partial Recall", "True_Recall" = "True Recall")),
    Sequence_Identity = label_value  # keep original values unchanged
    ))+
  scale_fill_gradient2(low = "lemonchiffon", mid="orange" ,high = "orchid", na.value = "grey", midpoint = 50)+
  scale_alpha_continuous(range = c(0.1, 1)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        strip.background = element_rect(color="black", fill="grey80"),
        plot.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        panel.spacing.y = unit(0.2, "lines"))  # reduce vertical space
ggsave("~/LongReadFusionCallerBenchmark/Figures/supp_Figure11_A.pdf", plot = supp_Figure11_A,  width = 400, height = 210 , unit ="mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/supp_Figure11_A.png", plot = supp_Figure11_A,  width = 400, height = 210 , unit ="mm")

supp_Figure11_B <-ggplot(filter(fusions_in_n_algorithms, depth == "100GB", control == "positive", overlap >= 2, recall_category == "True_Recall"), aes(x = fusion.gene.id, y = Algorithm, fill = Total_Read_Supp)) +
  geom_tile() +
  geom_text(aes(label = Total_Read_Supp), color = "black", size = 3, alpha =1) +  # Add numbers to the tiles, adjust color for visibility
  labs(x="", y="",
       fill = "Read Support") +
  theme_minimal() +
  facet_nested(Sequence_Identity ~ recall_category , scales = "free_x", space = "free",
               labeller = labeller(
    recall_category = as_labeller(c("Partial_Recall" = "Partial Recall", "True_Recall" = "True Recall")),
    Sequence_Identity = label_value  # keep original values unchanged
    ))+
  scale_fill_gradient2(low = "lemonchiffon", mid="orange" ,high = "orchid", na.value = "grey", midpoint = 50)+
  scale_alpha_continuous(range = c(0.1, 1)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        strip.background = element_rect(color="black", fill="grey80"),
        plot.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        panel.spacing.y = unit(0.2, "lines"))  # reduce vertical space
ggsave("~/LongReadFusionCallerBenchmark/Figures/supp_Figure11_B.pdf", plot = supp_Figure11_B,  width = 400, height = 210 , unit ="mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/supp_Figure11_B.png", plot = supp_Figure11_B,  width = 400, height = 210 , unit ="mm")

supp_Figure11 <- (supp_Figure11_A / supp_Figure11_B)+
  plot_annotation(tag_levels = "A")
ggsave("~/LongReadFusionCallerBenchmark/Figures/supp_Figure11.png", plot = supp_Figure11,  width = 400, height = 420 , unit ="mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/supp_Figure11.pdf", plot = supp_Figure11,  width = 400, height = 420 , unit ="mm")
```
Generate upset plots to see the total number of overlapping fusions between fusion callers
```{r}
Fusion_Simulated_Fusion_overlap_sets <- list(
  JAFFAL = unique(filter(combined_data, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "JAFFAL", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType), read_supp >= 2)$fusion.gene.id),
  Genion = unique(filter(combined_data, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "Genion", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType), read_supp >= 2)$fusion.gene.id),
  LongGF = unique(filter(combined_data, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "LongGF", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType), read_supp >= 2)$fusion.gene.id),
  FusionSeeker = unique(filter(combined_data, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "FusionSeeker", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType), read_supp >= 2)$fusion.gene.id)
)

# Convert the list to a format compatible with UpSetR
Fusion_overlap <- fromList(Fusion_Simulated_Fusion_overlap_sets)
```

Make Figure 1H
```{r}
True_overlap <- ComplexUpset::upset(
  Fusion_overlap, intersect = names(Fusion_Simulated_Fusion_overlap_sets[1:4]), name='',
  width_ratio=0.1, sort_intersections_by='degree',
  queries=list(upset_query(intersect=c("JAFFAL", "Genion", "LongGF", "FusionSeeker"), color='blue')),
  base_annotations = list(
        'Intersection size'=(intersection_size(
          text_colors=c(on_background='black', on_bar='white'),
            mapping=aes(fill='bars_color')
        ) + scale_fill_manual(values=c('bars_color'='springgreen4'), guide='none')
          + ylab('True Recall')
        ))) 
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure1H.pdf", plot = True_overlap, width = 297, height = 210, units = "mm")
```
Create Compound Figure 1:
```{r}
Figure1_paper <- (Figure1D | Figure1E)/
                  (Figure1F | Figure1G)/
                  (True_overlap) +   
  plot_annotation(tag_levels = list(c("D", "E", "F", "G", "H")), 
                  theme = theme(plot.title = element_text(size = 14, face = "bold")                                ))
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure1_paper.pdf", plot = Figure1_paper, width = 300, height = 430, units = "mm", limitsize = FALSE)
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure1_paper.png", plot = Figure1_paper, width = 300, height = 430, units = "mm", limitsize = FALSE)
```

Make Supplementary Figure 3 
```{r}
#Supplementary Figure 3A
Fusion_Simulated_Fusion_overlap_sets <- list(
  JAFFAL = unique(filter(combined_data, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "JAFFAL", grepl("false", fusionType), read_supp >= 2)$fusion.gene.id),
  Genion = unique(filter(combined_data, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "Genion", grepl("false", fusionType), read_supp >= 2)$fusion.gene.id),
  LongGF = unique(filter(combined_data, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "LongGF", grepl("false", fusionType), read_supp >= 2)$fusion.gene.id),
  FusionSeeker = unique(filter(combined_data, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "FusionSeeker", grepl("false", fusionType), read_supp >= 2)$fusion.gene.id))

Fusion_overlap <- fromList(Fusion_Simulated_Fusion_overlap_sets)

# Convert list to binary membership matrix
Fusion_overlap_matrix <- as.data.frame(Fusion_overlap)  # Convert list-based to a dataframe

# Ensure the matrix is in the correct format (0/1 values)
Fusion_overlap_matrix <- as.data.frame(lapply(Fusion_overlap_matrix, as.integer))

# Generate the UpSet plot using ComplexUpset
False_Overlap <- ComplexUpset::upset(
  Fusion_overlap_matrix, names(Fusion_overlap_matrix)[1:4],  # First four sets
  width_ratio=0.1, sort_intersections_by='degree',
  queries=list(upset_query(intersect=c("JAFFAL", "Genion", "LongGF", "FusionSeeker"), color='blue')),
 base_annotations = list(
        'Intersection size'=(intersection_size(
          text_colors=c(on_background='black', on_bar='black'),
            mapping=aes(fill='bars_color')
        ) + scale_fill_manual(values=c('bars_color'='red'), guide='none') 
          + ylab('False Calls')
  )))

#Supplementary Figure 3B
Fusion_Simulated_Fusion_overlap_sets <- list(
  JAFFAL = unique(filter(combined_data, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "JAFFAL", grepl("truncated|reverse|chromosomal_misalignment|wrong", fusionType), read_supp >= 2)$fusion.gene.id),
  Genion = unique(filter(combined_data, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "Genion", grepl("truncated|reverse|chromosomal_misalignment|wrong", fusionType), read_supp >= 2)$fusion.gene.id),
  LongGF = unique(filter(combined_data, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "LongGF", grepl("truncated|reverse|chromosomal_misalignment|wrong", fusionType), read_supp >= 2)$fusion.gene.id),
  FusionSeeker = unique(filter(combined_data, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "FusionSeeker", grepl("truncated|reverse|chromosomal_misalignment|wrong", fusionType), read_supp >= 2)$fusion.gene.id)
)

# Convert the list to a format compatible with UpSetR
Fusion_overlap <- fromList(Fusion_Simulated_Fusion_overlap_sets)
# Generate an upset plot without specifying intersect, which uses the default behavior
PartialCall_Overlap <- ComplexUpset::upset(
  Fusion_overlap, intersect = names(Fusion_Simulated_Fusion_overlap_sets[1:4]),
  width_ratio=0.1, sort_intersections_by='degree',
  queries=list(upset_query(intersect=c("JAFFAL", "Genion", "LongGF", "FusionSeeker"), color='blue')),
  base_annotations = list(
        'Intersection size'=(intersection_size(
          text_colors=c(on_background='black', on_bar='black'),
            mapping=aes(fill='bars_color')
        ) + scale_fill_manual(values=c('bars_color'='orange'), guide='none')
           + ylab('Partial Recall')
        ))) 
supp_Figure3 <-(False_Overlap/PartialCall_Overlap)+   
  plot_annotation(tag_levels = list(c("A", "", "", "B")))
ggsave("~/LongReadFusionCallerBenchmark/Figures/supp_Figure3.pdf", plot = supp_Figure3, width = 297, height =210 , units = "mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/supp_Figure3.png", plot = supp_Figure3, width = 297, height =210 , units = "mm")
supp_Figure3
```

Repeat analysis with removed obvious false chimeras to see if they have any influence. Any differences observed are negligible
```{r}
combined_data_with_heuristic <- filter(combined_data, read_supp >= 2) %>%  
  mutate(stat_category = case_when(
      grepl("false_fusion:mitochondrial|false_fusion:self_misalignment", fusionType) ~ "obvious_false",
      TRUE ~ "other")) %>% filter(stat_category == "other") %>% select(-'stat_category')

Fusion_Simulated_Fusion_overlap_sets <- list(
  JAFFAL = unique(filter(combined_data_with_heuristic, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "JAFFAL", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType), read_supp >= 2)$fusion.gene.id),
  Genion = unique(filter(combined_data_with_heuristic, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "Genion", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType), read_supp >= 2)$fusion.gene.id),
  LongGF = unique(filter(combined_data_with_heuristic, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "LongGF", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType), read_supp >= 2)$fusion.gene.id),
  FusionSeeker = unique(filter(combined_data_with_heuristic, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "FusionSeeker", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType), read_supp >= 2)$fusion.gene.id)
)

# Convert the list to a format compatible with UpSetR
Fusion_overlap <- fromList(Fusion_Simulated_Fusion_overlap_sets)
# Generate an upset plot without specifying intersect, which uses the default behavior
True_overlap <- ComplexUpset::upset(
  Fusion_overlap, intersect = names(Fusion_Simulated_Fusion_overlap_sets[1:4]), name='',
  width_ratio=0.1, sort_intersections_by='degree',
  queries=list(upset_query(intersect=c("JAFFAL", "Genion", "LongGF", "FusionSeeker"), color='blue')),
  base_annotations = list(
        'Intersection size'=(intersection_size(
          text_colors=c(on_background='black', on_bar='white'),
            mapping=aes(fill='bars_color')
        ) + scale_fill_manual(values=c('bars_color'='springgreen4'), guide='none')
          + ylab('True Recall')
        ))) 
ggsave("~/LongReadFusionCallerBenchmark/Figures/Compare_Contrast_True_Calls.pdf", plot = True_overlap, width = 297, height = 210, units = "mm")

Fusion_Simulated_Fusion_overlap_sets <- list(
  JAFFAL = unique(filter(combined_data_with_heuristic, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "JAFFAL", grepl("truncated|reverse|chromosomal_misalignment|wrong", fusionType), read_supp >= 2)$fusion.gene.id),
  Genion = unique(filter(combined_data_with_heuristic, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "Genion", grepl("truncated|reverse|chromosomal_misalignment|wrong", fusionType), read_supp >= 2)$fusion.gene.id),
  LongGF = unique(filter(combined_data_with_heuristic, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "LongGF", grepl("truncated|reverse|chromosomal_misalignment|wrong", fusionType), read_supp >= 2)$fusion.gene.id),
  FusionSeeker = unique(filter(combined_data_with_heuristic, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "FusionSeeker", grepl("truncated|reverse|chromosomal_misalignment|wrong", fusionType), read_supp >= 2)$fusion.gene.id)
)

# Convert the list to a format compatible with UpSetR
Fusion_overlap <- fromList(Fusion_Simulated_Fusion_overlap_sets)
# Generate an upset plot without specifying intersect, which uses the default behavior
PartialCall_Overlap <- ComplexUpset::upset(
  Fusion_overlap, intersect = names(Fusion_Simulated_Fusion_overlap_sets[1:4]),
  width_ratio=0.1, sort_intersections_by='degree',
  queries=list(upset_query(intersect=c("JAFFAL", "Genion", "LongGF", "FusionSeeker"), color='blue')),
  base_annotations = list(
        'Intersection size'=(intersection_size(
          text_colors=c(on_background='black', on_bar='black'),
            mapping=aes(fill='bars_color')
        ) + scale_fill_manual(values=c('bars_color'='orange'), guide='none')
           + ylab('Partial Recall')
        ))) 
ggsave("~/LongReadFusionCallerBenchmark/Figures/Compare_Contrast_Partial_Calls.pdf", plot = PartialCall_Overlap, width = 297, height = 210, units = "mm")

Fusion_Simulated_Fusion_overlap_sets <- list(
  JAFFAL = unique(filter(combined_data_with_heuristic, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "JAFFAL", grepl("false", fusionType), read_supp >= 2)$fusion.gene.id),
  Genion = unique(filter(combined_data_with_heuristic, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "Genion", grepl("false", fusionType), read_supp >= 2)$fusion.gene.id),
  LongGF = unique(filter(combined_data_with_heuristic, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "LongGF", grepl("false", fusionType), read_supp >= 2)$fusion.gene.id),
  FusionSeeker = unique(filter(combined_data_with_heuristic, depth == "100GB", Sequence_Identity == "95%", control == "positive", Algorithm == "FusionSeeker", grepl("false", fusionType), read_supp >= 2)$fusion.gene.id)
)

Fusion_overlap <- fromList(Fusion_Simulated_Fusion_overlap_sets)
# Convert list to binary membership matrix
Fusion_overlap_matrix <- as.data.frame(Fusion_overlap)  # Convert list-based to a dataframe

# Ensure the matrix is in the correct format (0/1 values)
Fusion_overlap_matrix <- as.data.frame(lapply(Fusion_overlap_matrix, as.integer))

# Generate the UpSet plot using ComplexUpset
False_Overlap <- ComplexUpset::upset(
  Fusion_overlap_matrix, names(Fusion_overlap_matrix)[1:4],  # First four sets
  width_ratio=0.1, sort_intersections_by='degree',
  queries=list(upset_query(intersect=c("JAFFAL", "Genion", "LongGF", "FusionSeeker"), color='blue')),
 base_annotations = list(
        'Intersection size'=(intersection_size(
          text_colors=c(on_background='black', on_bar='black'),
            mapping=aes(fill='bars_color')
        ) + scale_fill_manual(values=c('bars_color'='red'), guide='none') 
          + ylab('False Calls')
  )))
ggsave("~/LongReadFusionCallerBenchmark/Figures/Compare_Contrast_False_Calls.pdf", plot = False_Overlap, width = 297, height = 210, units = "mm")

```

Calculate the breakpoint accuracy
```{r}
Simulated_BreakPoint <- Simulated_Fusion_Info_3 
JAFFAL_BreakPoint <- Annot_JAFFAL_Sim[c(1:25, 43, 50:52)] %>% distinct() %>% filter(control == "positive", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType), spanning.reads >= 2) 
  #chrom1, base1, chrom2, base2
SB <- subset(Simulated_BreakPoint, fusion.gene.id == "ENSG00000009830:ENSG00000087258")
J <- subset(JAFFAL_BreakPoint, fusion.gene.id == "ENSG00000009830:ENSG00000087258" & depth == "100GB" & Sequence_Identity == "95%")
gene_names <- str_split("ENSG00000009830:ENSG00000087258", pattern = ":")[[1]] 
      gene1 <- gene_names[1]
      gene2 <- gene_names[2]
     
      SB1 <- as.integer(tail(str_split(subset(SB, alternative_ID == gene1)$exonEnds, ",")[[1]], 1))
      SB2 <- as.integer(head(str_split(subset(SB, alternative_ID == gene2)$exonStarts, ",")[[1]],1))

      JB1 <- as.numeric(J$base1) - SB1
      JB2 <- as.numeric(J$base2) - SB2
 
 list(c(gene1, gene2) , c(SB1, SB2) , c(JB1, JB2))
 
 nrow(distinct(Breakpoint_Accuracy))

Genion_BreakPoint <- Annot_Genion_Sim[c(1:27, 37, 44:46)] %>% distinct() %>% filter(control == "positive", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType)) %>% separate(V8, into = c("Gene1", "Gene2", "Gene3"), sep = ";") %>% 
  separate(Gene1, into = c("Gene1.exonStarts", "Gene1.exonEnds"), sep = "-") %>% separate(Gene2, into = c("Gene2.exonStarts", "Gene2.exonEnds"), sep = "-") %>% separate(Gene3, into = c("Gene3.exonStarts", "Gene3.exonEnds"), sep = "-") %>%
  separate(Gene1.exonStarts, into=c("C1", "Gene1.exonStarts")) %>% separate(Gene2.exonStarts, into=c("C2", "Gene2.exonStarts")) %>% separate(Gene3.exonStarts, into=c("C3", "Gene3.exonStarts")) %>% mutate_at(vars(C1, C2, C3), funs(paste0("chr", .)))# "C1",  "Gene1.2", "C2", "Gene2.1", "Gene2.2", "C3", "Gene3.1" 

JAFFAL_BreakPoint <- Annot_JAFFAL_Sim[c(1:25, 43, 50:52)] %>% distinct() %>% filter(control == "positive", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType), spanning.reads >= 2) #chrom1, base1, chrom2, base2
LongGF_BreakPoint <- Annot_LongGF_Sim[c(1:24, 34, 41:43)] %>% distinct() %>% filter(control == "positive", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType)) #chrom1, breakpoint1, chrom2, base2
FusionSeeker_BreakPoint <- Annot_FusionSeeker_Sim[c(1:28, 38, 45:47)] %>% distinct() %>% filter(control == "positive", !grepl("false|wrong|truncated|reverse|chromosomal_misalignment", fusionType)) #Chrom1, Breakpoint1, Chrom2, Base2

JAFFAL_BreakPoint_Accuracy <-  data.frame(matrix(nrow = 0, ncol = length(c('seq_depth', 'seq_id', 'gene_fusion', 'JB1', 'JB2'))))
colnames(JAFFAL_BreakPoint_Accuracy) <- c('seq_depth', 'seq_id', 'gene_fusion','JB1', 'JB2')
for (seq_depth in c('1GB' , '10GB' , '100GB')) {
  for (seq_id in c('85%' , '90%' , '95%')) {   
    for(gene_fusion in unique(c(JAFFAL_BreakPoint$fusion.gene.id))){
      SB <- subset(Simulated_BreakPoint, fusion.gene.id == gene_fusion)
      
      gene_names <- str_split(gene_fusion, pattern = ":")[[1]] 
      gene1 <- gene_names[1]
      gene2 <- gene_names[2]
     
      SB1 <- as.integer(tail(str_split(subset(SB, alternative_ID == gene1)$exonEnds, ",")[[1]], 1))
      SB2 <- as.integer(head(str_split(subset(SB, alternative_ID == gene2)$exonStarts, ",")[[1]],1))
    
      gene3 <- gene_names[3]
      SB3 <- NA
      SB4 <- NA
      if(is.na(gene3) == FALSE){
        SB3 <- as.integer(str_split(subset(SB, alternative_ID == gene2)$exonEnds, ",")[[1]])
        SB4 <- as.integer(str_split(subset(SB, alternative_ID == gene3)$exonStarts, ",")[[1]])
      }
      
      J  <- subset(JAFFAL_BreakPoint, fusion.gene.id == gene_fusion & depth == seq_depth & Sequence_Identity == seq_id) 
      if(nrow(J) > 0){
          JB1 <- as.numeric(J$base1) - SB1
          JB2 <- as.numeric(J$base2) - SB2
          }else{
            JB1 <- "NA"
            JB2 <- "NA"
            }
      new_row <-  data.frame(seq_depth, seq_id, gene_fusion, JB1, JB2)
        colnames(new_row) <- colnames(JAFFAL_BreakPoint_Accuracy)
         JAFFAL_BreakPoint_Accuracy <- rbind(JAFFAL_BreakPoint_Accuracy, new_row)
    }
  }
}

Genion_Breakpoint_Accuracy <- data.frame(matrix(nrow = 0, ncol = length(c('seq_depth', 'seq_id', 'gene_fusion', 'GB1', 'GB2', 'GB3', 'GB4'))))
colnames(Genion_Breakpoint_Accuracy) <- c('seq_depth', 'seq_id', 'gene_fusion', 'GB1', 'GB2', 'GB3', 'GB4')
for (seq_depth in c('1GB' , '10GB' , '100GB')) {
  for (seq_id in c('85%' , '90%' , '95%')) {   
    for(gene_fusion in unique(c(Genion_BreakPoint$V1))){
      SB <- subset(Simulated_BreakPoint, fusion.gene.id == gene_fusion)
      gene_names <- str_split(gene_fusion, pattern = ":")[[1]] 
      gene1 <- gene_names[1]
      gene2 <- gene_names[2]
     
      SB1 <- as.integer(tail(str_split(subset(SB, alternative_ID == gene1)$exonEnds, ",")[[1]], 1))
      SB2 <- as.integer(head(str_split(subset(SB, alternative_ID == gene2)$exonStarts, ",")[[1]],1))
      
      gene3 <- gene_names[3]
      SB3 <- NA
      SB4 <- NA
      if(is.na(gene3) == FALSE){
        SB3 <- as.integer(str_split(subset(SB, alternative_ID == gene2)$exonEnds, ",")[[1]])
        SB4 <- as.integer(str_split(subset(SB, alternative_ID == gene3)$exonStarts, ",")[[1]])
      } 
      G <- subset(Genion_BreakPoint,  V1 == gene_fusion & depth == seq_depth & Sequence_Identity == seq_id)
        if(nrow(G) > 0){
    
          GB1 <- as.numeric(G$Gene1.exonEnds) - SB1
          GB2 <- as.numeric(G$Gene2.exonStarts) - SB2
          GB3 <- "NA"
          GB4 <- "NA"
          if(is.na(gene3) == FALSE){
                GB3 <- toString(apply(outer(as.numeric(G$Gene2.exonEnds) , SB3, "-"), 1, function(x) x[which.min(abs(x))]))
                GB4 <- toString(apply(outer(as.numeric(G$Gene3.exonStarts) , SB4, "-"), 1, function(x) x[which.min(abs(x))]))
              }
          }else{
            GB1 <- "NA"
            GB2 <- "NA"
            GB3 <- "NA"
            GB4 <- "NA"
          }
     new_row <-  data.frame(seq_depth, seq_id, gene_fusion, GB1, GB2, GB3, GB4)
     colnames(new_row) <- colnames(Genion_Breakpoint_Accuracy)
     Genion_Breakpoint_Accuracy <- rbind(Genion_Breakpoint_Accuracy, new_row)
    }
   } 
  }
      
LongGF_Breakpoint_Accuracy <- data.frame(matrix(nrow = 0, ncol = length(c('seq_depth', 'seq_id', 'gene_fusion', 'LB1', 'LB2'))))
colnames(LongGF_Breakpoint_Accuracy) <- c('seq_depth', 'seq_id', 'gene_fusion', 'LB1', 'LB2')
for (seq_depth in c('1GB' , '10GB' , '100GB')) {
  for (seq_id in c('85%' , '90%' , '95%')) {   
    for(gene_fusion in unique(c(LongGF_BreakPoint$fusion.gene.id))){
      SB <- subset(Simulated_BreakPoint, fusion.gene.id == gene_fusion)
      gene_names <- str_split(gene_fusion, pattern = ":")[[1]] 
      gene1 <- gene_names[1]
      gene2 <- gene_names[2]
     
      SB1 <- as.integer(tail(str_split(subset(SB, alternative_ID == gene1)$exonEnds, ",")[[1]], 1))
      SB2 <- as.integer(head(str_split(subset(SB, alternative_ID == gene2)$exonStarts, ",")[[1]],1))
    
      gene3 <- gene_names[3]
      SB3 <- NA
      SB4 <- NA
      if(is.na(gene3) == FALSE){
        SB3 <- as.integer(str_split(subset(SB, alternative_ID == gene2)$exonEnds, ",")[[1]])
        SB4 <- as.integer(str_split(subset(SB, alternative_ID == gene3)$exonStarts, ",")[[1]])
      } 
      L <- subset(LongGF_BreakPoint, fusion.gene.id== gene_fusion & depth == seq_depth & Sequence_Identity == seq_id)
      if(nrow(L) > 0){
      LB1 <- as.numeric(L$breakpoint1) - SB1
      LB2 <- as.numeric(L$breakpoint2) - SB2
      }else{
        LB1 <- "NA"
        LB2 <- "NA"
        }
      new_row <-  data.frame(seq_depth, seq_id, gene_fusion, LB1, LB2)
      colnames(new_row) <- colnames(LongGF_Breakpoint_Accuracy)
      LongGF_Breakpoint_Accuracy <- rbind(LongGF_Breakpoint_Accuracy, new_row)
  }
 } 
}

FusionSeeker_Breakpoint_Accuracy <- data.frame(matrix(nrow = 0, ncol = length(c('seq_depth', 'seq_id', 'gene_fusion', 'FSB1', 'FSB2'))))
colnames(FusionSeeker_Breakpoint_Accuracy) <- c('seq_depth', 'seq_id', 'gene_fusion', 'FSB1', 'FSB2')
for (seq_depth in c('1GB' , '10GB' , '100GB')) {
  for (seq_id in c('85%' , '90%' , '95%')) {   
    for(gene_fusion in unique(c(FusionSeeker_BreakPoint$fusion.gene.id))){
      SB <- subset(Simulated_BreakPoint, fusion.gene.id == gene_fusion)
      FS <- subset(FusionSeeker_BreakPoint, fusion.gene.id == gene_fusion & depth == seq_depth & Sequence_Identity == seq_id)
      gene_names <- str_split(gene_fusion, pattern = ":")[[1]] 
      gene1 <- gene_names[1]
      gene2 <- gene_names[2]
     
      SB1 <- as.integer(tail(str_split(subset(SB, alternative_ID == gene1)$exonEnds, ",")[[1]], 1))
      SB2 <- as.integer(head(str_split(subset(SB, alternative_ID == gene2)$exonStarts, ",")[[1]],1))
    
      gene3 <- gene_names[3]
      SB3 <- NA
      SB4 <- NA
      if(is.na(gene3) == FALSE){
        SB3 <- as.integer(str_split(subset(SB, alternative_ID == gene2)$exonEnds, ",")[[1]])
        SB4 <- as.integer(str_split(subset(SB, alternative_ID == gene3)$exonStarts, ",")[[1]])
      }
      if(nrow(FS) > 0){
      FSB1 <- as.numeric(FS$Breakpoint1) - SB1
      FSB2 <- as.numeric(FS$Breakpoint2) - SB2
      }else{
        FSB1 <- "NA"
        FSB2 <- "NA"
      }
      new_row <-  data.frame(seq_depth, seq_id, gene_fusion, FSB1, FSB2)
      colnames(new_row) <- colnames(FusionSeeker_Breakpoint_Accuracy)
       FusionSeeker_Breakpoint_Accuracy <- rbind(FusionSeeker_Breakpoint_Accuracy, new_row)
      }
     } 
    }

Breakpoint_Accuracy <- full_join(FusionSeeker_Breakpoint_Accuracy, LongGF_Breakpoint_Accuracy) %>% full_join(Genion_Breakpoint_Accuracy) %>% full_join(JAFFAL_BreakPoint_Accuracy )

Breakpoint_Accuracy[is.na(Breakpoint_Accuracy)] <- "NA"

Breakpoint_Accuracy <- Breakpoint_Accuracy %>% filter(!if_all(FSB1:JB2, ~ .x == "NA")) %>%
  pivot_longer(cols = FSB1:JB2, names_to =  "Tool_Breakpoint", values_to = "Distance_from_Simulated_Breakpoint") %>%
  mutate(Algorithm = case_when(
    grepl("^GB", Tool_Breakpoint) ~ "Genion",
    grepl("^JB", Tool_Breakpoint) ~ "JAFFAL",
    grepl("^LB", Tool_Breakpoint) ~ "LongGF",
    grepl("^FSB", Tool_Breakpoint) ~ "FusionSeeker"))%>%
  mutate(values_vector = strsplit(as.character(Distance_from_Simulated_Breakpoint), ","))%>%
  unnest(values_vector) %>%
 mutate(values_vector = as.numeric(values_vector)) %>%
  mutate(Breakpoint_number = case_when(
    grepl("B1", Tool_Breakpoint) ~ "Breakpoint_1",
    grepl("B2", Tool_Breakpoint) ~ "Breakpoint_2",
    grepl("B3", Tool_Breakpoint) ~ "Breakpoint_3",
    grepl("B4", Tool_Breakpoint) ~ "Breakpoint_4"))

Breakpoint_Accuracy <-filter(Breakpoint_Accuracy, Distance_from_Simulated_Breakpoint != "NA") %>% distinct()
#some breakpoints are before the value and have a negative value, this we transform the values to absolute distances
absolute_Breakpoint_Accuracy <- Breakpoint_Accuracy %>%
  mutate(values_vector = abs(values_vector))
```

Make Figure 4 & Supplementary Figure 13
```{r}
library(ggforce)
absolute_Breakpoint_Accuracy$Distance_from_Simulated_Breakpoint <-
  as.numeric(as.character(absolute_Breakpoint_Accuracy$Distance_from_Simulated_Breakpoint))

#Statistical test: Check if data is normally distributed
breakpoint_stat_data <- filter(absolute_Breakpoint_Accuracy, seq_depth == "100GB", seq_id == "95%")
breakpoint_stat_data$Log1p_values <- log1p(breakpoint_stat_data$values_vector)
by(breakpoint_stat_data$Log1p_values,
   breakpoint_stat_data$Algorithm,
    shapiro.test)
breakpoint1_stat_data <- filter(absolute_Breakpoint_Accuracy, seq_depth == "100GB", seq_id == "95%", Breakpoint_number == "Breakpoint_1")
breakpoint1_stat_data$Log1p_values <- log1p(breakpoint1_stat_data$values_vector)
by(breakpoint1_stat_data$Log1p_values,
   breakpoint1_stat_data$Algorithm,
    shapiro.test)
breakpoint2_stat_data <- filter(absolute_Breakpoint_Accuracy, seq_depth == "100GB", seq_id == "95%", Breakpoint_number == "Breakpoint_2")
breakpoint2_stat_data$Log1p_values <- log1p(breakpoint2_stat_data$values_vector)
by(breakpoint2_stat_data$Log1p_values,
   breakpoint2_stat_data$Algorithm,
    shapiro.test)

ggplot(breakpoint_stat_data,
        aes(sample = Log1p_values)) +
     stat_qq() +
     stat_qq_line(color = "red") +
     facet_grid(Breakpoint_number ~ Algorithm) +
     theme_minimal() +
     labs(title = "Q–Q Plots by Algorithm and Breakpoint",
          x = "Theoretical Quantiles",
          y = "Sample Quantiles")
#Data is not normally distributed, so we will do a Kruskal Wallis Test followed by a pairwise test.
kruskal.test(Log1p_values ~ Algorithm, data = breakpoint_stat_data)
kruskal.test(Log1p_values ~ Algorithm, data = breakpoint1_stat_data)
kruskal.test(Log1p_values ~ Algorithm, data = breakpoint2_stat_data)

#Figure 4 ####
Figure4 <- ggplot(filter(absolute_Breakpoint_Accuracy, seq_depth == "100GB", seq_id == "95%"), 
       aes(y = Distance_from_Simulated_Breakpoint, x = Breakpoint_number, color = Algorithm)) +
  geom_boxplot(outlier.shape = NA) +
  geom_sina(maxwidth = 0.8)+
  theme_minimal()+
  theme(legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        panel.grid.minor = element_blank(),   # remove minor log lines
        panel.grid.major.x = element_blank(),  # optional: remove vertical major lines
        plot.title = element_text(size = 12),
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 12))+ 
  scale_y_continuous(transform = scales::pseudo_log_trans(base=10) , 
                     breaks = c(-1000000000, -100000000, -10000000, -1000000, -100000, -10000, -1000, -100, 
                                -10, -1, 
                                0, 
                                1, 10, 100, 1000, 10000, 100000, 1000000, 10000000, 100000000, 1000000000),
                     labels = c(expression(-10^9), expression(-10^8), expression(-10^7), expression(-10^6), expression(-10^5), expression(-10^4), expression(-10^3), expression(-10^2),
                                expression(-10^1), "-1", 
                                "0", 
                                "1", expression(10^1), 
                                expression(10^2), expression(10^3), expression(10^4), expression(10^5), expression(10^6), expression(10^7), expression(10^8), expression(10^9))
  )+   
  scale_x_discrete(labels = c("Breakpoint 1", "Breakpoint 2"))+
  labs(y = "Distance from simulated breakpoint",
       x= "")+
  stat_compare_means(method = "kruskal.test", aes(group = Algorithm), label.y = 8.5 , show.legend = FALSE)+
  geom_pwc(method = "dunn_test", p.adjust.method = "bonferroni", hide.ns = TRUE, label = "p.adj.signif",
           y.position = c(6.5, 7.5 , 5.5, 6.5 ,7.5 ), show.legend = FALSE)
 
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure4_paper.pdf", plot = Figure4, width = 297, height = 210, units = "mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/Figure4_paper.png", plot = Figure4)
           
supp_Figure13 <-ggplot(filter(absolute_Breakpoint_Accuracy), aes(y = values_vector, x = Breakpoint_number, color = Algorithm)) +
  geom_boxplot(outlier.shape = NA) +
  geom_sina()+
  theme_minimal()+ 
  scale_y_continuous(trans='log1p', 
                     breaks = c(0, 1, 10, 100, 
                                1000, 10000, 100000, 1000000, 
                                10000000, 100000000, 1000000000, 10000000000),
                     labels = c("0", "1", "10", expression(10^2),
             expression(10^3), expression(10^4), expression(10^5), expression(10^6),
             expression(10^7), expression(10^8), expression(10^9), expression(10^10)))+
  scale_x_discrete(labels = c("Breakpoint 1", "Breakpoint 2"))+
  labs(y = "Absolute distance from simulated breakpoint", x= "")+
  facet_grid(seq_depth ~ seq_id)+ 
  theme(legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        panel.grid.minor = element_blank(),   # remove minor log lines
        panel.grid.major.x = element_blank(),  # optional: remove vertical major linesplot.title = element_text(size = 12),
        strip.text.x = element_text(size = 12),
        strip.text.y = element_text(size = 12),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))
ggsave("~/LongReadFusionCallerBenchmark/Figures/supp_Figure13.png", plot = supp_Figure13,  width = 300, height = 420 , unit ="mm")
ggsave("~/LongReadFusionCallerBenchmark/Figures/supp_Figure13.pdf", plot = supp_Figure13,  width = 297, height = 420 , unit ="mm")
```

