---
title: "Cell line algorithm analysis"
author: "Ryley Dorney"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load relevant R libraries for data analysis
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library("viridis")  
library(UpSetR)
library(ComplexUpset)
library(biomaRt)
library(readxl)
```

The 'biomarRt' package is used to help identify corresponding gene ids to gene names and vice versa of our curated list of known gene fusions. 
```{r}
Cell_line_data_directory <- "~/LongReadFusionCallerBenchmark/CellLineData/"
known_fusions <- read.table(paste0(Cell_line_data_directory, "real_gene_fusions.tsv"), header=TRUE)
ensemblv110 <- useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl", version = 110)
gene_info <- getBM(attributes = c("external_gene_name", "ensembl_gene_id"),
                             filters = "external_gene_name",
                             values = unique(c(known_fusions$Gene1, known_fusions$Gene2)),
                             mart = ensemblv110) %>% unique()
known_fusions_biomart <- left_join(known_fusions,  gene_info, by  = c("Gene1"="external_gene_name")) %>% left_join(gene_info, by  = c("Gene2"="external_gene_name"))
write_csv(known_fusions_biomart, file="~/known_fusions_biomart.csv")
```

Read in the manually curated excel spreadsheet of known gene fusions and use biomart to check gene IDs
```{r}
known_fusions_manual_annotation <- read_excel("known_fusions_manual_annotation.xlsx")

gene_info <- getBM(attributes = c("external_gene_name", "ensembl_gene_id"),
                             filters = "ensembl_gene_id",
                             values = unique(c(known_fusions_manual_annotation$ensembl_gene_id.x, known_fusions_manual_annotation$ensembl_gene_id.y)),
                             mart = ensemblv110) %>% unique()

#make annotation of list of known fusions
known_fusions_biomart <-known_fusions_manual_annotation %>% left_join(gene_info, by  = c("ensembl_gene_id.x"="ensembl_gene_id")) %>% left_join(gene_info_2, by  = c("ensembl_gene_id.y"="ensembl_gene_id")) %>% select(c(1:6, 9, 12:13))%>% 
  mutate(external_gene_name.y = coalesce(external_gene_name.y, Gene2), external_gene_name.x = coalesce(external_gene_name.x, Gene1), fusionGeneID = paste0(ensembl_gene_id.x, "::", ensembl_gene_id.y) , fusionGene = paste0(external_gene_name.x, "::", external_gene_name.y))

```

Extract fusion genes called by longgf and output to a tsv file
```{bash}
#change to directory contain LongGF results of cell line data
LongGF_CellLine_Directory="/bioinformatics/ryley/Algorithm_Benchmark/SGNEx_files/LongGF/"
cd ${LongGF_CellLine_Directory}

#find the longgf fusion genes and output to a tsv file
bash ~/LongReadFusionCallerBenchmark/CellLineData/LongGF_extractSumGF.sh
```

Read in Data from fusioncalling output from different algorithms [LongGF, Genion, JAFFAL, FusionSeeker] performed on sequencing data from K562 and MCF7
```{r}
####Read in Data from fusioncalling output from different algorithms on K562 and MCF7
#LongGF
LongGF_CellLine_Directory<-"/bioinformatics/ryley/Algorithm_Benchmark/SGNEx_files/LongGF/"
my_files <- list.files(path = LongGF_CellLine_Directory, pattern = "tsv", full.names=TRUE)
all_tsv <-lapply(my_files, read.table)
names(all_tsv) <- gsub(".tsv","",list.files(path = LongGF_CellLine_Directory, pattern = "tsv",full.names = FALSE),fixed = TRUE)
names <- gsub(".tsv","",list.files(path = LongGF_CellLine_Directory, pattern = "tsv",full.names = FALSE),fixed = TRUE)
for(i in names){
  assign(i, read.table(paste0(LongGF_CellLine_Directory, i, ".tsv")))
}
LongGF_cell_lines <- do.call(rbind, lapply(my_files, function(filename) {
  read.table(filename) %>%
    mutate(Source = basename(filename))
}))

#Genion
Genion_CellLine_Directory <- "/bioinformatics/ryley/Algorithm_Benchmark/SGNEx_files/Genion"
myfiles<- list.files(path = Genion_CellLine_Directory, pattern = "*_genion", full.names = TRUE)
Genion_cell_lines<- do.call(rbind, lapply(myfiles, function(filename) {
  if (file.info(filename)$size > 0) {
    read.table(filename) %>%
      mutate(Source = sub("_genion$", "", basename(filename)))
  } else {
    NULL
  }
}))

#JAFFAL
CellLine_Directory <- "/bioinformatics/ryley/Algorithm_Benchmark/SGNEx_files"
myfiles<-list.files(path = CellLine_Directory, pattern = "jaffa_results.csv", full.names = TRUE, recursive = TRUE)
JAFFAL_cell_lines <-  do.call(rbind, lapply(myfiles, function(filename) {
  read.table(filename, header = TRUE, sep = ",") %>%
    mutate(Source =  sub(".fastq$", "",sample))
}))
#JAFFAL_3Gene 
myfiles<-list.files(path = CellLine_Directory, pattern = ".3gene_summary", full.names = TRUE, recursive = TRUE) 

JAFFAL_cell_lines_3Gene <- do.call(rbind, lapply(myfiles, function(filename) {
  read.table(filename, header = TRUE, sep = "\t") %>%
    mutate(Source = sub(".fastq$", "", basename(dirname(filename))))
}))

#fusionseeker
FusionSeeker_CellLine_Directory <- "/bioinformatics/ryley/Algorithm_Benchmark/SGNEx_files/FusionSeeker"

myfiles<-list.files(path = FusionSeeker_CellLine_Directory, pattern = "confident_genefusion.txt", full.names = TRUE, recursive = TRUE)
FusionSeeker_cell_lines <- do.call(rbind, lapply(myfiles, function(filename){
  read.table(filename, header = TRUE) %>% mutate(Source = sub("^fusionseeker_", "", basename(dirname(filename))))
}))
```

Make a column to indicate which cell line the fusion was found in, and a column to indicate what tool was used, and column to indicate the sequencing depth and library type
```{r}
LongGF_cell_lines <- LongGF_cell_lines %>%
  mutate(Cell_Lines = case_when(
    grepl("K562", Source, ignore.case = TRUE) ~ "K562",
    grepl("MCF7", Source, ignore.case = TRUE) ~ "MCF7"
  ), Algorithm = "LongGF",
  Sequencing_Depth = case_when(
    grepl("1G", Source, ignore.case = TRUE) ~ "1Gb",
    grepl("10G", Source, ignore.case = TRUE) ~ "10Gb",
    grepl("_5G", Source, ignore.case = TRUE) ~ "5Gb",
    grepl("2.5G", Source, ignore.case = TRUE) ~ "2.5Gb",
    grepl("7.5G", Source, ignore.case = TRUE) ~ "7.5Gb",
    .default = "Total"),
  Library = case_when(
    grepl("RNA", Source, ignore.case = TRUE) ~ "direct-RNA",
    grepl("directcDNA", Source, ignore.case = TRUE) ~ "direct-cDNA",
    .default = "PCR-cDNA"))

Genion_cell_lines <- Genion_cell_lines%>%
  mutate(Cell_Lines = case_when(
    grepl("K562", Source, ignore.case = TRUE) ~ "K562",
    grepl("MCF7", Source, ignore.case = TRUE) ~ "MCF7",# Keep original value if no match
  ), Algorithm = "Genion",
  Sequencing_Depth = case_when(
    grepl("1G", Source, ignore.case = TRUE) ~ "1Gb",
    grepl("10G", Source, ignore.case = TRUE) ~ "10Gb",
    grepl("_5G", Source, ignore.case = TRUE) ~ "5Gb",
    grepl("2.5G", Source, ignore.case = TRUE) ~ "2.5Gb",
    grepl("7.5G", Source, ignore.case = TRUE) ~ "7.5Gb",
    .default = "Total"),
  Library = case_when(
    grepl("RNA", Source, ignore.case = TRUE) ~ "direct-RNA",
    grepl("directcDNA", Source, ignore.case = TRUE) ~ "direct-cDNA",
    .default = "PCR-cDNA"))

FusionSeeker_cell_lines <- FusionSeeker_cell_lines %>%
  mutate(Cell_Lines = case_when(
    grepl("K562", Source, ignore.case = TRUE) ~ "K562",
    grepl("MCF7", Source, ignore.case = TRUE) ~ "MCF7",# Keep original value if no match
  ), Algorithm = "FusionSeeker",
  Sequencing_Depth = case_when(
    grepl("1G", Source, ignore.case = TRUE) ~ "1Gb",
    grepl("10G", Source, ignore.case = TRUE) ~ "10Gb",
    grepl("_5G", Source, ignore.case = TRUE) ~ "5Gb",
    grepl("2.5G", Source, ignore.case = TRUE) ~ "2.5Gb",
    grepl("7.5G", Source, ignore.case = TRUE) ~ "7.5Gb",
    .default = "Total"),
  Library = case_when(
    grepl("RNA", Source, ignore.case = TRUE) ~ "direct-RNA",
    grepl("directcDNA", Source, ignore.case = TRUE) ~ "direct-cDNA",
    .default = "PCR-cDNA"))

JAFFAL_cell_lines <- JAFFAL_cell_lines%>%
  mutate(Cell_Lines = case_when(
    grepl("K562", sample, ignore.case = TRUE) ~ "K562",
    grepl("MCF7", sample, ignore.case = TRUE) ~ "MCF7"),# Keep original value if no match
  Sequencing_Depth = case_when(
    grepl("1G", Source, ignore.case = TRUE) ~ "1Gb",
    grepl("10G", Source, ignore.case = TRUE) ~ "10Gb",
    grepl("_5G", Source, ignore.case = TRUE) ~ "5Gb",
    grepl("2.5G", Source, ignore.case = TRUE) ~ "2.5Gb",
    grepl("7.5G", Source, ignore.case = TRUE) ~ "7.5Gb",
    .default = "Total"), Algorithm = "JAFFAL",
  Library = case_when(
    grepl("RNA", Source, ignore.case = TRUE) ~ "direct-RNA",
    grepl("directcDNA", Source, ignore.case = TRUE) ~ "direct-cDNA",
    .default = "PCR-cDNA"))

JAFFAL_cell_lines_3Gene <- JAFFAL_cell_lines_3Gene %>%
  mutate(Cell_Lines = case_when(
    grepl("K562", Source, ignore.case = TRUE) ~ "K562",
    grepl("MCF7", Source, ignore.case = TRUE) ~ "MCF7"# Keep original value if no match
  ), Sequencing_Depth = case_when(
    grepl("1G", Source, ignore.case = TRUE) ~ "1Gb",
    grepl("10G", Source, ignore.case = TRUE) ~ "10Gb",
    grepl("_5G", Source, ignore.case = TRUE) ~ "5Gb",
    grepl("2.5G", Source, ignore.case = TRUE) ~ "2.5Gb",
    grepl("7.5G", Source, ignore.case = TRUE) ~ "7.5Gb",
    .default = "Total"), Algorithm = "JAFFAL",
  Library = case_when(
    grepl("RNA", Source, ignore.case = TRUE) ~ "direct-RNA",
    grepl("directcDNA", Source, ignore.case = TRUE) ~ "direct-cDNA",
    .default = "PCR-cDNA"))
```

Find "Mitochondrial:Genomic" fusions chrM x genomic genes, and "Mitochondrial" chimeras
```{r}
#find "Mitochondrial:Genomic" fusions chrM x genomic genes, and "Mitochondrial" chimeras
mitochondrial_genomic_FusionSeeker_cell_lines <- subset(FusionSeeker_cell_lines, (Chrom1 == "chrM" & Chrom2 != "chrM") | (Chrom1 != "chrM" & Chrom2 == "chrM")) %>% mutate(fusionType = "Mitochondrial:Genomic")

mitochondrial_genomic_LongGF_cell_lines <- filter(LongGF_cell_lines, (grepl("chrM:", V4, ignore.case = TRUE) & !grepl("chrM:", V5, ignore.case = TRUE)) | (!grepl("chrM:", V4, ignore.case = TRUE) & grepl("chrM:", V5, ignore.case = TRUE))) %>% mutate(fusionType = "Mitochondrial:Genomic")

mito_chimeras_FusionSeeker_cell_lines <- subset(FusionSeeker_cell_lines, (Chrom1 == "chrM" & Chrom2 == "chrM")) %>% mutate(fusionType = "Mitochondrial")

mito_chimeras_LongGF_cell_lines <- filter(LongGF_cell_lines, (grepl("chrM:", V4, ignore.case = TRUE) & grepl("chrM:", V5, ignore.case = TRUE))) %>% mutate(fusionType = "Mitochondrial")

mito_chimeras_Genion_cell_lines <- filter(Genion_cell_lines, grepl("M:", V8, ignore.case = TRUE)) #nothing observed

mito_chimeras_JAFFAL_cell_lines <- subset(JAFFAL_cell_lines, (chrom1=="chrM" | chrom2=="chrM")) #nothing observed
#combine into one data frame
mito_chimeras_cell_line <- rbind(mitochondrial_genomic_LongGF_cell_lines[c(3, 6:11)], mito_chimeras_LongGF_cell_lines[c(3, 6:11)])
mito_chimeras_cell_line$NumSupp <-mito_chimeras_cell_line$V3
mito_chimeras_cell_line <-mito_chimeras_cell_line[c(2:8)] %>% rbind(mito_chimeras_FusionSeeker_cell_lines[c(4,10:15)], mitochondrial_genomic_FusionSeeker_cell_lines[c(4,10:15)])

mito_chimeras_cell_line$Sequencing_Depth<- factor(mito_chimeras_cell_line$Sequencing_Depth, levels = c("1Gb", "2.5Gb", "5Gb", "7.5Gb", "10Gb", "Total"))
mito_chimeras_cell_line$Library<- factor(mito_chimeras_cell_line$Library, levels = c("direct-RNA", "direct-cDNA", "PCR-cDNA"))
#plot the mito_chimeras_cell_line data
ggplot(filter(mito_chimeras_cell_line, NumSupp >=2), aes(x = Sequencing_Depth, fill = fusionType)) +
  geom_bar() +
  theme_minimal() +
  labs(title = "False calls by each algorithm", subtitle = "minimum read support of 2", x = "Read Depth", y = "Count")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_grid(Library+Cell_Lines~Algorithm) +
  scale_y_log10()

ggplot(filter(mito_chimeras_cell_line, NumSupp >=2), aes(x = Sequencing_Depth, y = NumSupp)) +
  geom_boxplot(aes(fill = fusionType)) +
  labs(title = "Distribution of Read Support for false chimeras",
       x = "Read Depth",
       y = "Read Support") +
  theme_minimal() +
  facet_grid(Library+Cell_Lines~ Algorithm) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

filter misalignments and fusions between self, e.g. paralogus genes fused 
```{r}
self_alignment_LongGF_cell_lines <- LongGF_cell_lines %>% separate(V2, into = c("V2.1","V2.2"), sep=":", remove = FALSE) %>%  subset(V2.1 == V2.2) %>% rename("NumSupp"="V3")
self_alignment_FusionSeeker_cell_lines <- FusionSeeker_cell_lines %>%
  subset((Gene1 == Gene2) | 
         mapply(grepl, pattern = paste0(Gene1, "_PAR_Y"), x = Gene2) | 
         mapply(grepl, pattern = paste0(Gene2, "_PAR_Y"), x = Gene1))
self_alignment_Genion_cell_lines <- Genion_cell_lines %>% separate(V1, into = c("V1.1","V1.2"), sep="::", remove = FALSE) %>%  subset((V1.1 == V1.2) | mapply(grepl, pattern = paste0(V1.1, "_PAR_Y"), x = V1.2) | 
         mapply(grepl, pattern = paste0(V1.2, "_PAR_Y"), x = V1.1))
self_alignment_JAFFAL_cell_lines <- JAFFAL_cell_lines %>% separate(fusion.genes, into = c("V1.1","V1.2"), sep=":", remove = FALSE) %>% subset(V1.1 == V1.2)


self_alignment_results <- rbind(self_alignment_LongGF_cell_lines[c(5, 8:12)], self_alignment_FusionSeeker_cell_lines[c(4,10:14)]) %>% mutate(fusionType = "Self-Misalignment")
```

combine paralogous gene fusions and mitochondrial chimeras into a dataframe containing obvious false_chimeras
```{r}
false_chimeras <- rbind(self_alignment_results, mito_chimeras_cell_line)
false_chimeras$Sequencing_Depth<- factor(false_chimeras$Sequencing_Depth, levels = c("1Gb", "2.5Gb", "5Gb", "7.5Gb", "10Gb", "Total"))
false_chimeras$Library<- factor(false_chimeras$Library, levels = c("direct-RNA", "direct-cDNA", "PCR-cDNA"))

supp_Figure14 <-ggplot(filter(false_chimeras, NumSupp >=2, Sequencing_Depth == c("1Gb", "2.5Gb", "5Gb", "7.5Gb", "10Gb")), aes(x = Sequencing_Depth, fill = fusionType)) +
  geom_bar() +
  theme_minimal() +
  labs(title = "False calls by each algorithm", subtitle = "minimum read support of 2", x = "Read Depth", y = "Count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))+
  facet_grid(Library+Cell_Lines~Algorithm) +
  scale_y_log10()+
  labs(fill = "Fusion Type") 
ggsave("supp_Figure14.png", plot = supp_Figure14,  width = 210, height =297  , unit ="mm")
```

Filter out the obvious false chimeras from the original dataframe and assign ensembl gene IDs to fusion gene partners
```{r}
#filter out the obvious false chimeras from the data
filtered_LongGF_cell_lines <- anti_join(LongGF_cell_lines, mito_chimeras_LongGF_cell_lines) %>% anti_join(mitochondrial_genomic_LongGF_cell_lines) %>% anti_join(self_alignment_LongGF_cell_lines)
filtered_FusionSeeker_cell_lines <- anti_join(FusionSeeker_cell_lines, mitochondrial_genomic_FusionSeeker_cell_lines) %>% anti_join(mito_chimeras_FusionSeeker_cell_lines) %>% anti_join(self_alignment_FusionSeeker_cell_lines)
```

Seperate fusion gene partners, make new dataframe (F)usion (SEP)arated Algorithm (C)el(l)
```{r}
fsep_LGcl <- filtered_LongGF_cell_lines %>% separate(V2, into = c('Gene1', 'Gene2'), sep = ":", remove = FALSE)
#assign ensembl gene IDs to fusion gene partners
gene_info <- getBM(attributes = c("external_gene_name", "ensembl_gene_id"),
                              filters = "external_gene_name",
                             values = unique(c(fsep_LGcl$Gene1, fsep_LGcl$Gene2)),
                             mart = ensemblv110)
fsep_LGcl <-left_join(fsep_LGcl, gene_info, by  = c("Gene1"="external_gene_name")) %>% left_join(gene_info, by  = c("Gene2"="external_gene_name"))%>% 
  mutate(ensembl_gene_id.y = coalesce(ensembl_gene_id.y, Gene2), ensembl_gene_id.x = coalesce(ensembl_gene_id.x, Gene1)) %>%  unique()

#repeat above for fusionseeker
fsep_FScl <- filtered_FusionSeeker_cell_lines 
gene_info <- getBM(attributes = c("external_gene_name", "ensembl_gene_id", "ensembl_gene_id_version"),
                              filters = "ensembl_gene_id_version",
                             values = unique(c(fsep_FScl$Gene1, fsep_FScl$Gene2)),
                             mart = ensemblv110)
fsep_FScl <- left_join(fsep_FScl, gene_info, by  = c("Gene1"="ensembl_gene_id_version")) %>% left_join(gene_info, by  = c("Gene2"="ensembl_gene_id_version")) %>%  unique() 

#repeat above for JAFFAL
fsep_JAFFALcl <- JAFFAL_cell_lines %>% separate(fusion.genes, into = c('Gene1', 'Gene2'), sep = ":", remove = FALSE)
gene_info <- getBM(attributes = c("external_gene_name", "ensembl_gene_id"),
                              filters = "external_gene_name",
                             values = unique(c(fsep_JAFFALcl$Gene1, fsep_JAFFALcl$Gene2)),
                             mart = ensemblv110)
fsep_JAFFALcl <-left_join(fsep_JAFFALcl, gene_info, by = c("Gene1"="external_gene_name")) %>% 
  left_join(gene_info, by  = c("Gene2"="external_gene_name"))%>% 
  mutate(ensembl_gene_id.y = coalesce(ensembl_gene_id.y, Gene2), ensembl_gene_id.x = coalesce(ensembl_gene_id.x, Gene1)) %>%  unique() 

fsep_JAFFALcl_3Gene <- JAFFAL_cell_lines_3Gene %>% filter(Reads > 1) %>% separate(Fusion, into = c('Gene1', 'Gene2', 'Gene3'), sep = ":", remove = FALSE)
gene_info <- getBM(attributes = c("external_gene_name", "ensembl_gene_id"),
                              filters = "external_gene_name",
                             values = unique(c(fsep_JAFFALcl_3Gene$Gene1, fsep_JAFFALcl_3Gene$Gene2, fsep_JAFFALcl_3Gene$Gene3)),
                             mart = ensemblv110)

fsep_JAFFALcl_3Gene <-left_join(fsep_JAFFALcl_3Gene, gene_info, by  = c("Gene1"="external_gene_name")) %>% left_join(gene_info, by  = c("Gene2"="external_gene_name"))%>% left_join(gene_info, by  = c("Gene3"="external_gene_name"))%>%
  mutate(ensembl_gene_id = coalesce(ensembl_gene_id, Gene3), ensembl_gene_id.y = coalesce(ensembl_gene_id.y, Gene2), ensembl_gene_id.x = coalesce(ensembl_gene_id.x, Gene1)) %>%  unique() 

#separate fusion genes for Genion
fsep_Gencl <- Genion_cell_lines %>% separate(V1, into = c("Gene1", "Gene2", "Gene3"), "::", remove=FALSE)
```

Check which known gene fusions have been re-discovered
```{r}
known_fusions_biomart$discovery <- "known"
fsep_JAFFALcl <- filter(fsep_JAFFALcl, spanning.reads >=2)
fsep_Gencl <- filter(fsep_Gencl, V5 >=2)
fsep_FScl  <- filter(fsep_FScl, NumSupp >= 2)
fsep_LGcl <- filter(fsep_LGcl, V3 >=2)

disc_JAFFALcl <- left_join(fsep_JAFFALcl, known_fusions_biomart[c(1,6,7,10:12)], by = c("ensembl_gene_id.x", "ensembl_gene_id.y", "Cell_Lines"="Cell_Line")) %>%  mutate(discovery = if_else(is.na(discovery), "novel", discovery))
check <- disc_JAFFALcl$fusion.genes[disc_JAFFALcl$discovery=="known"] %>% unique()
disc_JAFFALcl <-disc_JAFFALcl %>%  mutate(discovery = case_when(
    fusion.genes %in% check ~ "known",  # if fusion.genes is in the known list
    is.na(discovery) ~ "novel",         # if discovery is NA, label as novel
    TRUE ~ discovery)) %>%  mutate(fusionGeneID = paste0(ensembl_gene_id.x, "::", ensembl_gene_id.y))
disc_JAFFALcl_3Gene <- fsep_JAFFALcl_3Gene %>%  mutate(discovery ="novel", fusionGeneID = paste0(ensembl_gene_id.x, "::", ensembl_gene_id.y, "::", ensembl_gene_id)) %>% rename(fusion.genes = Fusion, spanning.reads = Reads, classification = Classification)

CellData_summary_JAFFAL_df <- disc_JAFFALcl %>% full_join(disc_JAFFALcl_3Gene) %>%
  group_by(across(all_of(colnames(JAFFAL_cell_lines)))) %>%
  mutate(
    knowncount = case_when(discovery == "known" ~ 1, TRUE ~ 0),
    novelcount = case_when(discovery == "novel" ~ 1, TRUE ~ 0))
CellData_summary_JAFFAL_df <- CellData_summary_JAFFAL_df %>% group_by(Source, Cell_Lines, 
    Sequencing_Depth, Algorithm, Library) %>%
  summarise(knownnumber = sum(knowncount), novelnumber= sum(novelcount))

disc_Genioncl <- left_join(fsep_Gencl, known_fusions_biomart[c(1,6,7,10:12)], by = c("Gene1" = "ensembl_gene_id.x", "Gene2" = "ensembl_gene_id.y", "Cell_Lines"="Cell_Line")) %>%  mutate(discovery = if_else(is.na(discovery), "novel", discovery))
check <- disc_Genioncl$V1[disc_Genioncl$discovery=="known"] %>% unique()
disc_Genioncl <-disc_Genioncl %>%  mutate(discovery = case_when(
    V1 %in% check ~ "known",  # if the fusion is in the known list
    is.na(discovery) ~ "novel",         # if discovery is NA, label as novel
    TRUE ~ discovery))
CellData_summary_Genion_df <- disc_Genioncl %>%
  group_by(across(all_of(colnames(Genion_cell_lines)))) %>%
  mutate(
    knowncount = case_when(discovery == "known" ~ 1, TRUE ~ 0),
    novelcount = case_when(discovery == "novel" ~ 1, TRUE ~ 0))
CellData_summary_Genion_df <- CellData_summary_Genion_df %>% group_by(Source, Cell_Lines, 
    Sequencing_Depth, Algorithm, Library) %>%
  summarise(knownnumber = sum(knowncount), novelnumber= sum(novelcount))

disc_FusionSeekercl <- left_join(fsep_FScl, known_fusions_biomart[c(1,6,7,10:12)], by = c("ensembl_gene_id.x", "ensembl_gene_id.y", "Cell_Lines"="Cell_Line")) %>%  mutate(discovery = if_else(is.na(discovery), "novel", discovery))
check <- disc_FusionSeekercl$fusionGeneID[disc_FusionSeekercl$discovery=="known"] %>% unique()
disc_FusionSeekercl <-disc_FusionSeekercl %>% mutate(fusionGeneID = paste0(ensembl_gene_id.x, "::", ensembl_gene_id.y)) %>%  mutate(discovery = case_when(
    fusionGeneID %in% check ~ "known",  # if fusion.genes is in the known list
    is.na(discovery) ~ "novel",         # if discovery is NA, label as novel
    TRUE ~ discovery))
CellData_summary_FusionSeeker_df <- disc_FusionSeekercl %>%
  group_by(across(all_of(colnames(FusionSeeker_cell_lines)))) %>%
  mutate(
    knowncount = case_when(discovery == "known" ~ 1, TRUE ~ 0),
    novelcount = case_when(discovery == "novel" ~ 1, TRUE ~ 0))
CellData_summary_FusionSeeker_df <- CellData_summary_FusionSeeker_df %>% group_by(Source, Cell_Lines, 
    Sequencing_Depth, Algorithm, Library) %>%
  summarise(knownnumber = sum(knowncount), novelnumber= sum(novelcount))

disc_LongGFcl <- left_join(fsep_LGcl, known_fusions_biomart[c(1,6,7,10:12)], by = c("ensembl_gene_id.x", "ensembl_gene_id.y", "Cell_Lines"="Cell_Line")) %>%  mutate(discovery = if_else(is.na(discovery), "novel", discovery)) %>% mutate(fusionGeneID = paste0(ensembl_gene_id.x, "::", ensembl_gene_id.y))
check <- disc_LongGFcl$fusionGeneID[disc_LongGFcl$discovery=="known"] %>% unique()
disc_LongGFcl <-disc_LongGFcl %>%  mutate(discovery = case_when(
    fusionGeneID %in% check ~ "known",  # if fusion.genes is in the known list
    is.na(discovery) ~ "novel",         # if discovery is NA, label as novel
    TRUE ~ discovery))
CellData_summary_LongGF_df <- disc_LongGFcl %>%
  group_by(across(all_of(colnames(LongGF_cell_lines)))) %>%
  mutate(
    knowncount = case_when(discovery == "known" ~ 1, TRUE ~ 0),
    novelcount = case_when(discovery == "novel" ~ 1, TRUE ~ 0))
CellData_summary_LongGF_df <- CellData_summary_LongGF_df %>% group_by(Source, Cell_Lines, 
    Sequencing_Depth, Algorithm, Library) %>%
  summarise(knownnumber = sum(knowncount), novelnumber= sum(novelcount))

super_CellData_summary <- rbind(CellData_summary_FusionSeeker_df, CellData_summary_Genion_df, CellData_summary_JAFFAL_df, CellData_summary_LongGF_df)
super_CellData_summary$Sequencing_Depth<- factor(super_CellData_summary$Sequencing_Depth, levels = c("1Gb", "2.5Gb", "5Gb", "7.5Gb", "10Gb", "Total"))
super_CellData_summary$Library<- factor(super_CellData_summary$Library, levels = c("direct-RNA", "direct-cDNA", "PCR-cDNA"))
super_CellData_summary %>% filter(Sequencing_Depth != "Total")

```
Make Figure 5: novel vs known fusions
```{r}
#Figure 5 ####
# Plot for K562 with smaller y-axis limit
p_k562 <- ggplot(filter(super_CellData_summary, Sequencing_Depth != "Total", Cell_Lines == "K562")) +
  geom_point(aes(x = novelnumber, y = knownnumber, colour = Algorithm, size = Sequencing_Depth)) +
  facet_grid(Library ~ Cell_Lines)+ 
  theme(legend.position = "none",
        strip.text.y = element_blank(),         # removes row facet labels (Library)
        strip.background.y = element_blank(),    # removes grey background for row labels
        strip.text.x = element_text(size = 12),
        plot.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))+
  scale_x_continuous(trans='log1p', 
                     breaks = c(0, 1, 10, 100, 
                                1000, 10000),
                     labels = c("0", "1", "10", expression(10^2),
                                expression(10^3), expression(10^4)))+
  labs(x = "Novel Fusions", y = "Known Fusions") +
  ylim(0, 6)  # Adjust this limit to your preference

# Plot for MCF7 with larger y-axis limit
p_mcf7 <- ggplot(filter(super_CellData_summary, Sequencing_Depth != "Total", Cell_Lines == "MCF7")) +
  geom_point(aes(x = novelnumber, y = knownnumber, colour = Algorithm, size = Sequencing_Depth)) +
  facet_grid(Library ~ Cell_Lines) +
  labs(x = "Novel Fusions", y = "")+ 
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        strip.text = element_text(size = 12),
        plot.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))+
  scale_x_continuous(trans='log1p', 
                     breaks = c(0, 1, 10, 100, 
                                1000, 10000),
                     labels = c("0", "1", "10", expression(10^2),
                                expression(10^3), expression(10^4)))+
  ylim(0, 55)  # Adjust as needed

#combine the two plots
Figure5 <- p_k562 + p_mcf7
ggsave("Figure5_paper.pdf", plot = Figure5, width = 210, height =200 , units = "mm")
ggsave("Figure5_paper.jpeg", plot = Figure5, width = 210, height =200 , units = "mm")
```

Check how many fusions were found by more than one algorithm
```{r}
Overlap_CellData_JAFFAL <- full_join(disc_JAFFALcl, disc_JAFFALcl_3Gene )%>%
  filter(spanning.reads >=2) %>% rename(read_support = spanning.reads) %>%
  select(c(fusionGeneID, read_support, Source, Cell_Lines, Sequencing_Depth, Algorithm, Library))
Overlap_CellData_Genion <- fsep_Gencl %>% filter(V5 >=2) %>%
  rename(fusionGeneID = V1, read_support = V5)%>% 
  select(c(fusionGeneID, read_support, Source, Cell_Lines, Sequencing_Depth, Algorithm, Library))
Overlap_CellData_FusionSeeker <- fsep_FScl %>%
  mutate(fusionGeneID = paste0(ensembl_gene_id.x, "::", ensembl_gene_id.y))  %>%
  filter(NumSupp >= 2) %>%
  rename(read_support = NumSupp)%>%
  select(c(fusionGeneID, read_support, Source, Cell_Lines, Sequencing_Depth, Algorithm, Library))
Overlap_CellData_LongGF <- fsep_LGcl %>%
  mutate(fusionGeneID = paste0(ensembl_gene_id.x, "::", ensembl_gene_id.y)) %>%
  filter(V3 >=2)  %>%
  rename(read_support = V3)%>% select(c(fusionGeneID, read_support, Source, Cell_Lines, Sequencing_Depth, Algorithm, Library))

head(rbind(Overlap_CellData_JAFFAL, Overlap_CellData_Genion, Overlap_CellData_FusionSeeker, Overlap_CellData_LongGF))
library(dplyr)
library(tidyr)
library(UpSetR)
library(ComplexUpset)
fusion_data <- rbind(Overlap_CellData_JAFFAL, Overlap_CellData_Genion, Overlap_CellData_FusionSeeker, Overlap_CellData_LongGF)

Overlap_CellData_algorithms <- fusion_data %>% 
  group_by(fusionGeneID, Cell_Lines, Sequencing_Depth, Library) %>%
  summarise(overlap = n_distinct(Algorithm), .groups = "drop") %>%
  left_join(fusion_data, by = c("fusionGeneID", "Cell_Lines", "Sequencing_Depth", "Library"))%>%
  group_by(fusionGeneID, Cell_Lines, Sequencing_Depth, Library, Algorithm, overlap) %>%
  summarise(Total_Read_Supp = sum(read_support), .groups = "drop")
Overlap_CellData_algorithms$Sequencing_Depth<- factor(Overlap_CellData_algorithms$Sequencing_Depth, levels = c("1Gb", "2.5Gb", "5Gb", "7.5Gb", "10Gb", "Total"))
Overlap_CellData_algorithms$Library<- factor(Overlap_CellData_algorithms$Library, levels = c("direct-RNA", "direct-cDNA", "PCR-cDNA"))

Overlap_CellData_discovery <- left_join(Overlap_CellData_algorithms, known_fusions_biomart[c(1,6,7,10:12)], by = c("fusionGeneID", "Cell_Lines"="Cell_Line")) %>%  mutate(discovery = if_else(is.na(discovery), "novel", discovery))

# Step 1: Summarize Read_Supp per algorithm
OverlapknownOnly <- Overlap_CellData_discovery %>%
  filter(discovery != "novel") %>%
  group_by(fusionGene, Cell_Lines, Sequencing_Depth, Library, Algorithm) %>%
  summarise(Read_Supp = sum(Total_Read_Supp), .groups = "drop")

# Step 2: Compute overlap (distinct algorithms per fusion call)
Overlap_CellData_info <- OverlapknownOnly %>%
  group_by(fusionGene, Cell_Lines, Sequencing_Depth, Library) %>%
  summarise(overlap = n_distinct(Algorithm), .groups = "drop")

# Step 3: Join back to retain Algorithm column
OverlapknownOnly <- OverlapknownOnly %>%
  left_join(Overlap_CellData_info, by = c("fusionGene", "Cell_Lines", "Sequencing_Depth", "Library")) %>%
  mutate(fusionGene = fct_reorder(fusionGene, overlap, .desc = TRUE))

ggplot(filter(OverlapknownOnly, Sequencing_Depth == "Total"), 
       aes(x = fusionGene, y = Algorithm, fill = Read_Supp)) +
  geom_tile() +
  geom_text(aes(label = Read_Supp), color = "black", size = 3, alpha =1) +  # Add numbers to the tiles, adjust color for visibility
  labs(x="", y="", fill = "Read Support") +
  theme_minimal() +
  facet_grid(Library ~ Cell_Lines, scales = "free_x", space = "free") +
  scale_fill_gradient2(low = "lemonchiffon", mid="orange" ,high = "orchid", na.value = "grey", midpoint = 1000)+
  scale_alpha_continuous(range = c(0.1, 1)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.background = element_rect(color="black", fill="grey80"))

#Supplementary figure 12####
supp_Figure12 <-  ggplot(filter(Overlap_CellData_algorithms, Sequencing_Depth == "Total", overlap > 3), 
       aes(x = fusionGeneID, y = Algorithm, fill = Total_Read_Supp)) +
  geom_tile() +
  geom_text(aes(label = Total_Read_Supp), color = "black", size = 4, alpha =1) +  # Add numbers to the tiles, adjust color for visibility
  labs(x="", y="", fill = "") +
  theme_minimal() +
  facet_grid(Library ~ Cell_Lines, scales = "free_x", space = "free") +
  scale_fill_gradient2(low = "lemonchiffon", mid="orange" ,high = "orchid", na.value = "grey", midpoint = 30)+
  scale_alpha_continuous(range = c(0.1, 1)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        strip.background = element_rect(color="black", fill="grey80"),
        plot.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))
ggsave("supp_Figure12.png", plot = supp_Figure12,  width = 297, height = 210 , unit ="mm")
ggsave("supp_Figure12.pdf", plot = supp_Figure12,  width = 297, height = 210 , unit ="mm")

```

```{r}
Overlap_CellData_total <- Overlap_CellData_discovery %>% filter(Sequencing_Depth == "Total") %>%
  separate()

ggplot(filter(Overlap_CellData_total, str_count(fusionGeneID, "::") == 2), 
       aes(x = fusionGeneID, y = Algorithm, fill = Total_Read_Supp)) +
  geom_tile() +
  geom_text(aes(label = Total_Read_Supp), color = "black", size = 3, alpha =1) +  # Add numbers to the tiles, adjust color for visibility
  labs(x="", y="", fill = "") +
  theme_minimal() +
  facet_grid(Library ~ Cell_Lines, scales = "free_x", space = "free") +
   scale_fill_gradient2(low = "lemonchiffon", mid="orange" ,high = "orchid", na.value = "grey", , midpoint = 30)+
  scale_alpha_continuous(range = c(0.1, 1)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.background = element_rect(color="black", fill="grey80"))

MCF7_total_directcDNA <- Overlap_CellData_total %>% filter(Cell_Lines == "MCF7", Library == "direct-cDNA") %>% select(c(discovery, Algorithm, fusionGeneID ))
```

```{r}
shared_fusions_cell <- Overlap_CellData_algorithms %>% filter(Sequencing_Depth == "Total")%>%
  separate(fusionGeneID, into = c("Gene1","Gene2","Gene3"), sep="::", remove = FALSE)

gene_names_shared_fusions_cell <- getBM(attributes = c("external_gene_name", "ensembl_gene_id"),
                              filters = "ensembl_gene_id",
                             values = unique(c(shared_fusions_cell$Gene1, shared_fusions_cell$Gene2, shared_fusions_cell$Gene3)),
                             mart = ensemblv110) %>% unique()

shared_fusions_cell <- left_join(shared_fusions_cell, gene_names_shared_fusions_cell, by = c("Gene1" = "ensembl_gene_id")) %>%
  left_join(gene_names_shared_fusions_cell, by = c("Gene2" = "ensembl_gene_id")) %>%
  left_join(gene_names_shared_fusions_cell, by = c("Gene3" = "ensembl_gene_id")) %>%
          mutate(ensembl_gene1 = ifelse(is.na(external_gene_name.x) |nchar(external_gene_name.x) == 0, Gene1, external_gene_name.x), 
                 ensembl_gene2 = ifelse(is.na(external_gene_name.y) |nchar(external_gene_name.y) == 0, Gene2, external_gene_name.y), 
                 ensembl_gene3 = ifelse(is.na(external_gene_name) | nchar(external_gene_name) == 0, Gene3, external_gene_name)) %>%
  unite("fusionGene", c(ensembl_gene1, ensembl_gene2, ensembl_gene3), sep="::", remove= FALSE, na.rm = TRUE) 

shared_fusions_cell_upset <- shared_fusions_cell %>%
  distinct(Cell_Lines, Library, Algorithm, fusionGene) %>%
  mutate(present = 1) %>%
  pivot_wider(
    names_from = Algorithm,
    values_from = present,
    values_fill = 0
  )

shared_fusions_cell_upset  %>%
  filter(Cell_Lines == "K562", Library == "direct-cDNA") %>%
  upset(
    intersect = c("FusionSeeker", "LongGF", "JAFFAL", "Genion"),
    name = "Fusions"
  )
shared_fusions_cell_upset  %>%
  filter(Cell_Lines == "K562", Library == "direct-cDNA") %>%
  upset(
    intersect = c("FusionSeeker", "LongGF", "JAFFAL", "Genion"),
    name = "Fusions"
  )
```


```{r}
#Depmap known fusions
DepMap_knownfusions <- read.csv("~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/DepMap_OmicsFusionFiltered.csv", 
           header = TRUE)
Model_Depmap <- (read.csv("~/LibraryBenchmarkAnalysis/LibraryBenchmarkAnalysis_RProject/Model.csv", 
           header = TRUE))
DepMap_knownfusions <- left_join(DepMap_knownfusions, Model_Depmap) %>% select(Gene1, Gene2, StrippedCellLineName) %>% filter((StrippedCellLineName == "K562"| StrippedCellLineName =="MCF7"))

DepMap_K562 <- filter(DepMap_knownfusions, StrippedCellLineName == "K562")%>%
  extract(Gene1, into = c("Gene1_name", "Gene1_ensembl"), 
          regex = "([^ ]+) \\(([^)]+)\\)", remove = FALSE) %>%
  extract(Gene2, into = c("Gene2_name", "Gene2_ensembl"), 
          regex = "([^ ]+) \\(([^)]+)\\)", remove = FALSE) %>%
  mutate(
    Gene1_ensembl = na_if(Gene1_ensembl, "."),
    Gene2_ensembl = na_if(Gene2_ensembl, "."))

DepMap_MCF7 <- filter(DepMap_knownfusions, StrippedCellLineName == "MCF7")%>%
  extract(Gene1, into = c("Gene1_name", "Gene1_ensembl"), 
          regex = "([^ ]+) \\(([^)]+)\\)", remove = FALSE) %>%
  extract(Gene2, into = c("Gene2_name", "Gene2_ensembl"), 
          regex = "([^ ]+) \\(([^)]+)\\)", remove = FALSE) %>%
  mutate(
    Gene1_ensembl = na_if(Gene1_ensembl, "."),
    Gene2_ensembl = na_if(Gene2_ensembl, "."))

ensembl_current <-  useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl")
GeneIDs<-unique(rbind(getBM(attributes = c("external_gene_name", "ensembl_gene_id", "ensembl_gene_id_version", "external_synonym", "hgnc_symbol"),
                        filters = c("external_gene_name"),
                        values = unique(c(DepMap_MCF7$Gene1_name, DepMap_MCF7$Gene2_name, DepMap_K562$Gene1_name, DepMap_K562$Gene2_name)),
                        mart = ensembl_current), 
               getBM(attributes = c("external_gene_name", "ensembl_gene_id", "ensembl_gene_id_version", "external_synonym", "hgnc_symbol"),
                        filters = c("external_synonym"),
                        values = unique(c(DepMap_MCF7$Gene1_name, DepMap_MCF7$Gene2_name, DepMap_K562$Gene1_name, DepMap_K562$Gene2_name)),
                        mart = ensembl_current), 
               getBM(attributes = c("external_gene_name", "ensembl_gene_id", "ensembl_gene_id_version", "external_synonym", "hgnc_symbol"),
                        filters = c("hgnc_symbol"),
                        values = unique(c(DepMap_MCF7$Gene1_name, DepMap_MCF7$Gene2_name, DepMap_K562$Gene1_name, DepMap_K562$Gene2_name)),
                        mart = ensembl_current)))

GeneIDs_long <- GeneIDs %>%
  pivot_longer(cols = c(external_gene_name, external_synonym), 
               names_to = "name_type", values_to = "gene_symbol") %>%
  distinct()

DepMap_K562_withID <- DepMap_K562 %>%
  left_join(GeneIDs_long, by = c("Gene1_name" = "gene_symbol")) %>%
  left_join(GeneIDs_long, by = c("Gene2_name" = "gene_symbol"), suffix = c("_Gene1", "_Gene2"))

DepMap_MCF7_withID <- DepMap_MCF7 %>%
  left_join(GeneIDs_long, by = c("Gene1_name" = "gene_symbol")) %>%
  left_join(GeneIDs_long, by = c("Gene2_name" = "gene_symbol"), suffix = c("_Gene1", "_Gene2"))

DepMap_specificCells <- rbind(DepMap_K562_withID, DepMap_MCF7_withID)
DepMap_specificCells$discovery <- "known"

disc_JAFFALcl <- left_join(fsep_JAFFALcl, DepMap_specificCells[c(2,5,7,8,12,16)], by = c(
    "Gene1" = "Gene1_name",
    "Gene2" = "Gene2_name",
    "ensembl_gene_id.x" = "ensembl_gene_id_Gene1",
    "ensembl_gene_id.y" = "ensembl_gene_id_Gene2",
    "Cell_Lines" = "StrippedCellLineName"
  )) %>%
  mutate(discovery = if_else(is.na(discovery), "novel", discovery))
check <- disc_JAFFALcl$fusion.genes[disc_JAFFALcl$discovery=="known"] %>% unique()
disc_JAFFALcl <-disc_JAFFALcl %>%  mutate(discovery = case_when(
    fusion.genes %in% check ~ "known",  # if fusion.genes is in the known list
    is.na(discovery) ~ "novel",         # if discovery is NA, label as novel
    TRUE ~ discovery)) %>%  mutate(fusionGeneID = paste0(ensembl_gene_id.x, "::", ensembl_gene_id.y))
disc_JAFFALcl_3Gene <- fsep_JAFFALcl_3Gene %>%  mutate(discovery ="novel", fusionGeneID = paste0(ensembl_gene_id.x, "::", ensembl_gene_id.y, "::", ensembl_gene_id)) %>% rename(fusion.genes = Fusion, spanning.reads = Reads, classification = Classification)

CellData_summary_JAFFAL_df <- disc_JAFFALcl %>% full_join(disc_JAFFALcl_3Gene) %>%
  group_by(across(all_of(colnames(JAFFAL_cell_lines)))) %>%
  mutate(
    knowncount = case_when(discovery == "known" ~ 1, TRUE ~ 0),
    novelcount = case_when(discovery == "novel" ~ 1, TRUE ~ 0))
CellData_summary_JAFFAL_df <- CellData_summary_JAFFAL_df %>% group_by(Source, Cell_Lines, 
    Sequencing_Depth, Algorithm, Library) %>%
  summarise(knownnumber = sum(knowncount), novelnumber= sum(novelcount))

disc_Genioncl <- left_join(fsep_Gencl, DepMap_specificCells[c(2,3,7,8,12,16)], by = c("Gene1" = "ensembl_gene_id.x", "Gene2" = "ensembl_gene_id.y", "Cell_Lines"="Cell_Line")) %>%  mutate(discovery = if_else(is.na(discovery), "novel", discovery))
check <- disc_Genioncl$V1[disc_Genioncl$discovery=="known"] %>% unique()
disc_Genioncl <-disc_Genioncl %>%  mutate(discovery = case_when(
    V1 %in% check ~ "known",  # if the fusion is in the known list
    is.na(discovery) ~ "novel",         # if discovery is NA, label as novel
    TRUE ~ discovery))
CellData_summary_Genion_df <- disc_Genioncl %>%
  group_by(across(all_of(colnames(Genion_cell_lines)))) %>%
  mutate(
    knowncount = case_when(discovery == "known" ~ 1, TRUE ~ 0),
    novelcount = case_when(discovery == "novel" ~ 1, TRUE ~ 0))
CellData_summary_Genion_df <- CellData_summary_Genion_df %>% group_by(Source, Cell_Lines, 
    Sequencing_Depth, Algorithm, Library) %>%
  summarise(knownnumber = sum(knowncount), novelnumber= sum(novelcount))

disc_FusionSeekercl <- left_join(fsep_FScl, DepMap_specificCells[c(2,3,7,8,12,16)], by = c("ensembl_gene_id.x", "ensembl_gene_id.y", "Cell_Lines"="Cell_Line")) %>%  mutate(discovery = if_else(is.na(discovery), "novel", discovery))
check <- disc_FusionSeekercl$fusionGeneID[disc_FusionSeekercl$discovery=="known"] %>% unique()
disc_FusionSeekercl <-disc_FusionSeekercl %>% mutate(fusionGeneID = paste0(ensembl_gene_id.x, "::", ensembl_gene_id.y)) %>%  mutate(discovery = case_when(
    fusionGeneID %in% check ~ "known",  # if fusion.genes is in the known list
    is.na(discovery) ~ "novel",         # if discovery is NA, label as novel
    TRUE ~ discovery))
CellData_summary_FusionSeeker_df <- disc_FusionSeekercl %>%
  group_by(across(all_of(colnames(FusionSeeker_cell_lines)))) %>%
  mutate(
    knowncount = case_when(discovery == "known" ~ 1, TRUE ~ 0),
    novelcount = case_when(discovery == "novel" ~ 1, TRUE ~ 0))
CellData_summary_FusionSeeker_df <- CellData_summary_FusionSeeker_df %>% group_by(Source, Cell_Lines, 
    Sequencing_Depth, Algorithm, Library) %>%
  summarise(knownnumber = sum(knowncount), novelnumber= sum(novelcount))

disc_LongGFcl <- left_join(fsep_LGcl, DepMap_specificCells[c(2,3,7,8,12,16)], by = c("ensembl_gene_id.x", "ensembl_gene_id.y", "Cell_Lines"="Cell_Line")) %>%  mutate(discovery = if_else(is.na(discovery), "novel", discovery)) %>% mutate(fusionGeneID = paste0(ensembl_gene_id.x, "::", ensembl_gene_id.y))
check <- disc_LongGFcl$fusionGeneID[disc_LongGFcl$discovery=="known"] %>% unique()
disc_LongGFcl <-disc_LongGFcl %>%  mutate(discovery = case_when(
    fusionGeneID %in% check ~ "known",  # if fusion.genes is in the known list
    is.na(discovery) ~ "novel",         # if discovery is NA, label as novel
    TRUE ~ discovery))
CellData_summary_LongGF_df <- disc_LongGFcl %>%
  group_by(across(all_of(colnames(LongGF_cell_lines)))) %>%
  mutate(
    knowncount = case_when(discovery == "known" ~ 1, TRUE ~ 0),
    novelcount = case_when(discovery == "novel" ~ 1, TRUE ~ 0))
CellData_summary_LongGF_df <- CellData_summary_LongGF_df %>% group_by(Source, Cell_Lines, 
    Sequencing_Depth, Algorithm, Library) %>%
  summarise(knownnumber = sum(knowncount), novelnumber= sum(novelcount))

super_CellData_summary <- rbind(CellData_summary_FusionSeeker_df, CellData_summary_Genion_df, CellData_summary_JAFFAL_df, CellData_summary_LongGF_df)
# Convert Sequencing_Depth into an ordered factor
super_CellData_summary$Sequencing_Depth<- factor(super_CellData_summary$Sequencing_Depth, levels = c("1Gb", "2.5Gb", "5Gb", "7.5Gb", "10Gb", "Total"))
# Convert Library type into an ordered factor
super_CellData_summary$Library<- factor(super_CellData_summary$Library, levels = c("direct-RNA", "direct-cDNA", "PCR-cDNA"))
# Quick check: filter out "Total" depth entries
super_CellData_summary %>% filter(Sequencing_Depth != "Total")
 
Cell_Line_novelvsknown <- ggplot(filter(super_CellData_summary, Sequencing_Depth != "Total"))+
  geom_point(aes(x = novelnumber , y =knownnumber, colour = Algorithm, shape = Sequencing_Depth, ))+
  facet_grid(Library~Cell_Lines, scales = "free_y")+
  labs(x="Novel Fusions", y = "Known Fusions") 
ggsave("Cell_Line_novelvsknown.pdf", plot = Cell_Line_novelvsknown, width = 297, height = 210, units = "mm")
```

