---
title: "SGNex algorithm analysis"
author: "Ryley Dorney"
date: "`r Sys.Date()`"
output: html_document
---
Load relevant R libraries for data analysis
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyr)
library(naniar)
library(readr)
library(dplyr)
library(ggplot2)
library(ggh4x)
library(biomaRt)
library(UpSetR)
library(ComplexUpset)
library(patchwork)
library(ggforce)
library(ggpubr)
library("viridis")  
library(readxl)
```

The 'biomarRt' package is used to help identify corresponding gene ids to gene names and vice versa of our curated list of known gene fusions. 
```{r}
source("~/LongReadFusionCallerBenchmark/CellLineData/Known_K562_MCF7.R")
```

Extract fusion genes called by longgf and output to a tsv file
```{bash}
SGNEX_DIR=/bioinformatics/ryley/Gencode44/SGNex_Fusionv44/

#find the longgf fusion genes and output to a tsv file
bash ~/LongReadFusionCallerBenchmark/CellLineData/LongGF_extractSumGF.sh
```

Read in Data from fusioncalling output from different algorithms [LongGF, Genion, JAFFAL, FusionSeeker, CTAT-LR and GFSeeler] performed on sequencing data from K562 and MCF7
```{r}
####Read in Data from fusioncalling output from different algorithms on K562 and MCF7
source("~/LongReadFusionCallerBenchmark/CellLineData/PreprocessSGNexFusions.R")
```

```{r}
############################
# set up dataframes to check
############################
CTATLR_sensecheck <- CTATLR_SGNex
JAFFAL_sensecheck <- JAFFAL_SGNex
Genion_sensecheck <- Genion_SGNex
FusionSeeker_sensecheck <- FusionSeeker_SGNex
LongGF_sensecheck <- LongGF_SGNex
GFSeeker_sensecheck <- GFSeeker_SGNex

#######################################
# Run check atypical and false fusions
#######################################
source("~/LongReadFusionCallerBenchmark/CellLineData/CheckAtypicalFusions.R")

CTATLR_SGNex_msa_annot <- CTATLR_sensemito_annotated
JAFFAL_SGNex_msa_annot <- JAFFAL_sensemito_annotated
Genion_SGNex_msa_annot <- Genion_sensemito_annotated
FusionSeeker_SGNex_msa_annot <- FusionSeeker_sensemito_annotated
LongGF_SGNex_msa_annot <- LongGF_sensemito_annotated
GFSeeker_SGNex_msa_annot <- GFSeeker_sensemito_annotated

#Annotate other types of fusions
source("~/LongReadFusionCallerBenchmark/CellLineData/Annotate_fusiontype_SGNex.R")


```

Seperate fusion gene partners, make new dataframe (F)usion (SEP)arated Algorithm (C)el(l)
```{r}
Annot_CTATLR_SGNex 
Annot_JAFFAL_SGNex 
Annot_JAFFAL_SGNex_3Gene
Annot_Genion_SGNex 
Annot_FusionSeeker_SGNex 
Annot_LongGF_SGNex 
Annot_GFSeeker_SGNex 

source("~/LongReadFusionCallerBenchmark/CellLineData/SGNex_assign_EnsemblID.R")
```

Check which known gene fusions have been re-discovered
```{r}
CTATLR_SGNex_Annot <- Annot_CTATLR_SGNex  
JAFFAL_SGNex_Annot <- Annot_JAFFAL_SGNex
Genion_SGNex_Annot <- Annot_Genion_SGNex 
FusionSeeker_SGNex_Annot <-Annot_FusionSeeker_SGNex 
GFSeeker_SGNex_Annot <- Annot_GFSeeker_SGNex 
LongGF_SGNex_Annot <-Annot_LongGF_SGNex 

source("~/LongReadFusionCallerBenchmark/CellLineData/SGNex_DiscoveryNRecall.R")
```

Check how many fusions were found by more than one algorithm
```{r}
source("~/LongReadFusionCallerBenchmark/CellLineData/SGNex_Overlap_Fusions.R")
```



